### 2020-10-19 FirstDraftRMarkdown.Rmd
Add files via upload

---
title: "Analyzing the Effects of the NBA COVID Bubble on the NBA Playoffs"
subtitle: "A Case Study for Home-Court Advantage"
author: |
  | Michael Price
  | michael.price@uconn.edu
  | Department of Statistics, University of Connecticut
  | Dr. Jun Yan
  | jun.yan@uconn.edu
  | Department of Statistics, University of Connecticut
date: "10/19/2020"
output: pdf_document
---
#Abstract: 
Due to the COVID-19 pandemic the 2020 NBA playoffs were played inside of a bubble in Disney World. This meant that there would be no fans in attendance and no traveling for teams which should result in removing home-court advantage from the games. This idea was a hot topic of discussion as analysts and fans were concerned over the possible effects it may have on the outcome of playoff games. Home court advantage has historically played a huge role in NBA playoff series outcomes. Thus, to study the effects of the bubble(and home-court advantage) the 2020 season can be compared to the 2017, 2018, and 2019 playoffs. While many factors contribute to the outcome of games, the bottom line is that points scored is the deciding factor of who wins games so scoring will be the primary focus of this study. The measures of interest are team scoring totals and team shooting percentage on two pointers, three pointers, and free throws. Comparing these measures for home teams and away teams in 2020 vs. 2017-2019 shows how the 2020 playoffs favored away teams more than normal.

\newpage
# I. Introduction  
|   Home court advantage is often discussed in sports circles as a contributing factor to the outcome of games. It is well known that the home team typically benefits from some competitive edge from playing at their home court, resulting in a better chance of winning. Due to this fact, the NBA playing the 2020 playoffs in a bubble brought a great deal of concern for fans, teams, journalists, and others. Specific details about anticipated effects are discussed in Steve Aschburner’s article “How Orlando's Neutral Site Will Impact Teams, Referees and Games”(2020). Aschburner explains that due to the unprecedented circumstances nobody was really sure exactly how the NBA playoffs would play out. Aschburner also voiced some concern about the potential effects of removing home-court advantage. The NBA did make attempts to recreate the effects by putting the "home” team logo on the court and allowing the "home" team to play crowd noise and music, but most people doubted these attempts would recreate a true playoff atmosphere. During the 2020 NBA playoffs, home teams only won about 48.2% of the games. This was indeed lower than normal, which is better represented using the 2017-2019 NBA playoff data where home teams won 61.32% of games. This shift in the home team winning percentage surely indicates the need for more thorough research and investigation.

|   So, what happened? Did the home teams fail to perform up to normal standards without the help of home-court advantage? Were away teams able to rise to the occasion and perform better not having to deal with the headache of going on the road? This paper seeks to answer just that, using scoring totals and scoring efficiency as indicators of team performance. This study will also deepen understanding of how home-court advantage affects home and away teams on a nightly basis in the NBA.

|  This study is quite different from previous home-court advantage studies. By using the neutral site games of 2020 we will get to compare home and away performance to a control. Typically, studies just compare home vs away performance. These studies don’t separate the effects of home-court advantage into the specific effect on the home team and the specific effect on the away team. They’re focused mainly on the fact home teams typically outperform away teams.There are some similar studies that use college basketball, which does play some neutral court games during their season, but these studies still mainly focus on whether or not home teams outperform away teams and not how home and away teams are impacted individually. Examples of these studies will be reviewed in greater detail in Section II Literature Review.

|  This study will compare home team performance in 2020 at a neutral site with no fans vs. 2017-19 playoffs with fans. Likewise, away team performance in 2020 at a neutral site with no fans vs. 2017-19 playoffs with fans. By comparing home teams in 2020 to home teams in 2017-19 and away teams in 2020 to away teams in 2017-19, I can add a new perspective to the field of research. This will allow for a more accurate understanding of the effects of home-court advantage on home and away teams in the NBA. We won’t just see that home-court advantage helps home teams outperform away teams, but this study will actually dissect the effects of home-court advantage on home teams and away teams performance individually. I want to add how home-court advantage actually comes to fruition to build on the existing knowledge that there is a home-court advantage effect.

|  9 statistical tests can be run to understand the differences in 2020 vs. previous years. First, whether or not the difference between home win percentage in 2020 and 2017-19 is statistically significant. To confirm the difference we see in the winning percentage is truly significant. Then we can look for differences in home scoring in 2020 vs 2017-2019. Similarly, we can do the same test, but for differences in away scoring in 2020 vs 2017-2019. Also, differences in team shooting(for two pointers, three pointers, and free throws) from 2020 vs 2017-2019 for both home and away teams. The results from these simple tests will bring a new understanding of how home-court advantage is impacting games by altering the performance of the home and away teams.

# II. Literature Review  
|	There is tons of published literature on the effects of home-court advantage. Jason Kotecki(2014) studies the impact of home-court advantage on winning in the NBA. Kotecki finds that home-court advantage is found when using performance-based statistics. Kotecki specifically cites field goal percentage, free throw percentage, and points. All of which he proves to be significantly impacted by home-court. Kotecki concludes that home teams will play better at home in front of their home fans. This is done by comparing home performance vs. away performance in games. This is similar to many other studies in this field. Given the data and circumstances, this has been the best way to study the effects in the NBA.

|	College basketball, as noted before, does provide some opportunity to study home-court vs. neutral court. One study conducted by David A. Harville and Michael H. Smith studied the effect of home court advantage using the 1991-1992 college basketball season. This study design is more similar to my own. Unlike the NBA, it is not uncommon to have a few games played at neutral sites during the college basketball season. This allowed them to construct two samples, one of home teams and one of neutral teams. They formulate their study as a regression equation predicting the expected difference in score for home teams. They set up their study to find if the home teams won games by more points when they had home court advantage vs. when playing at neutral court. This study succeeds in proving that home-court advantage exists. However, it doesn’t allow them to make inferences about how home-court advantage specifically impacts home and away performance. Like other research in the field its focus is on the fact that home teams outperform away teams.

```{r setup, include=FALSE}
#Data Collection
game_log_2020<-read.csv("Desktop/Thesis/2020 game data.csv", header=T, sep = ",")
game_log_2017_2019<-read.csv("Desktop/Thesis/2017 2019 game data.csv", header=T)

#Data Cleaning
game_log_2020$W.L<-ifelse(game_log_2020$W.L=="W", 1, 0)
game_log_2017_2019$W.L<-ifelse(game_log_2017_2019$W.L=="W", 1, 0)

game_log_2020_home<-game_log_2020[game_log_2020$Home.Away=="FALSE",]
game_log_2020_away<-game_log_2020[game_log_2020$Home.Away=="TRUE",]
game_log_2017_2019_home<-game_log_2017_2019[game_log_2017_2019$Home.Away=="FALSE", ]
game_log_2017_2019_away<-game_log_2017_2019[game_log_2017_2019$Home.Away=="TRUE", ]

years<-2010:2020
winpct<-c(.6707, .667, .679, .635, .562, .593, .6744, .5696, .707, .561, .482)
```

# III. Data Collection and Methods  
|  The 2020 bubble provides a new and exciting opportunity to study home-court advantage for the NBA. Unlike college basketball, aside from a few exhibition/preseason games, the NBA always has a home and away team. So, for the first time in NBA history the bubble allows NBA home and away performance to be compared vs a control/neutral field. To construct this study data was collected from the official NBA website, NBA.com. The NBA bubble featured 8 seeding games then a standard playoff format. The focus of this study was the play during the playoff games since it followed the standard playoff format and can easily be compared back to other playoffs. So, this study was designed to compare the 2020 playoffs to the three previous playoffs collectively. It is important to make sure that any observed differences are truly from a home-court advantage. So, in order to control for the changing play style of the NBA, we will limit the study to 2020 vs 2017-2019. The reason being the faster pace play and more common use of the three-point shot in modern basketball. If we used data from say 10 years ago, or earlier, observed differences may not be from effects of the NBA bubble, but rather from the effects of drastic changes in the style of play between the seasons. However, basketball evolves relatively slowly so we can reasonably assume 2017-2019 are at least very close in pace and playing style to 2020.

  The main variables of interest are whether or not the home team won, scoring totals for home and away teams, and shooting percentages for home and away teams on two pointers, three pointers, and free throws. While many other measures could be used for measuring the outcome of the game and team performance, scoring seemed to be the most important. The winner of a game is determined by who scores more points. So, naturally when studying the outcome of winning, changes in scoring and scoring efficiency are important to understand.

  Home court advantage is the basic idea that the home team is more likely to win. So laying a foundation of typical home-court advantage is crucial. Before focusing on the 2017 to 2020 playoffs we can take a quick look at home team win percentages since 2010. Notice on the graph below(Figure 1), the 10 years before 2020, the home team winning percentage ranged from around 0.56 to 0.7 and never dipped below .5. The 2020 bubble, however, broke this historic pattern dipping down below .5. Foreshadowing, the confirmation of the expectation that the effect of home-court advantage was removed in the 2020 playoffs.

##Figure 1
```{r, echo=FALSE}
plot(years, winpct, ylim = c(0,1), title("NBA Home Playoff Win % Since 2010"), col = ifelse(winpct==.494, "Red", "Black"), pch=16)
text(years, winpct, labels = winpct, cex = 0.5, pos = 1)
text(2010, 0.51, labels = "0.5", cex = 0.6, col = "Green")
abline(h=0.5, col = "Green")
```

  Moving on to the main focus of the study, comparing 2020 to 2017-2019. First, a brief look at the distribution of scoring depicted below in Figure 2. The green histograms indicate home and red histograms indicate away scoring. The top row is scoring for 2020 and the bottom is scoring for 2017-2019. All histograms seem to be fairly normally distributed which is important for statistical analysis. It’s hard to see any clear differences between years, but it’s important to look at and understand the distribution of measures before moving onto the analysis.

##Figure 2
```{r, echo=FALSE}
opar<-par(no.readonly = TRUE)
par(mfrow=c(2,2))
hist(game_log_2020_home$PTS, xlab = "Points", col = "green",  xlim = c(60, 160), main = "Home 2020")
hist(game_log_2020_away$PTS,  xlab = "Points", col = "red",  xlim = c(60, 160), main = "Away 2020")
hist(game_log_2017_2019_home$PTS, xlab = "Points", col="green",  xlim = c(60, 160), main = "Home 2017-2019" )
hist(game_log_2017_2019_away$PTS, xlab = "Points", col = "red",  xlim = c(60, 160), main = "Home 2017-2019", breaks = 8 )
par(opar)
```

  Figure 3 shows home shooting for two pointers, three pointers and free throws for 2020(top) vs. 2017-19(bottom). The histograms appear to be fairly similarly distributed between 2020 and 2017-19. Likewise, Figure 4, shows the same data except for away teams. Just like Figure 3, Figure 4 histograms don’t show any obvious differences. The normality of the distributions for both figures 3 and 4 are important for meeting the required assumptions for statistical analysis.

##Figure 3
```{r, echo=FALSE}
opar<-par(no.readonly = TRUE)
par(mfrow=c(2,3))
hist(game_log_2020_home$X2P., xlab = "Shooting Percentage", col= "green", main = "2020 Home 2 pointers", xlim = c(20, 70))
hist(game_log_2020_home$X3P., xlab = "Shooting Percentage", col= "green", main = "2020 Home 3 pointers", xlim = c(10, 70))
hist(game_log_2020_home$FT., xlab = "Shooting Percentage", col= "green", main = "2020 Home Free Throws", xlim = c( 50, 100))
hist(game_log_2017_2019_home$X2P., xlab = "Shooting Percentage", col= "green", main = "2017-2019 2 pointers", xlim = c(20, 70))
hist(game_log_2017_2019_home$X3P., xlab = "Shooting Percentage", col= "green", main = "2017-2019 3 pointers", xlim = c(10, 70))
hist(game_log_2017_2019_home$FT., xlab = "Shooting Percentage", col= "green", main = "2017-2019 Free Throws", xlim = c( 50, 100))
par(opar)
```

##Figure 4
```{r, echo=FALSE}
opar<-par(no.readonly = TRUE)
par(mfrow=c(2,3))
hist(game_log_2020_away$X2P., xlab = "Shooting Percentage", col= "red", main = "2020 Away 2 pointers", xlim= c(30, 80))
hist(game_log_2020_away$X3P., xlab = "Shooting Percentage", col= "red", main = "2020 Away 3 pointers", xlim = c(10, 60))
hist(game_log_2020_away$FT., xlab = "Shooting Percentage", col= "red", main = "2020 Away Free Throws", xlim = c(40, 100))
hist(game_log_2017_2019_away$X2P., xlab = "Shooting Percentage", col= "red", main = "2017-19 Away 2 pointers", xlim= c(30, 80))
hist(game_log_2017_2019_away$X3P., xlab = "Shooting Percentage", col= "red", main = "2017-19 Away 3 pointers", xlim = c(10, 60))
hist(game_log_2017_2019_away$FT., xlab = "Shooting Percentage", col= "red", main = "2017-19 Away Free Throws", xlim = c(40, 100))
par(opar)
```

  Given the approximate normality of data, 9 different z-tests were performed in order to test the effects of the COVID bubble on the 2020 NBA playoffs. The 9 questions answered were:  
* Is the home team winning percentage in 2020 different than that it was in 2017-2019?  
* Is the average home team scoring different in 2020 than it was over 2017-2019?  
* Is the average away team scoring different in 2020 than it was over 2017-2019?  
* Are home teams making two pointers at the same rate in 2020 as 2017-2019?  
* Are home teams making three pointers at the same rate in 2020 as 2017-2019?  
* Are home teams making free throws at the same rate in 2020 as 2017-2019?  
* Are away teams making two pointers at the same rate in 2020 as 2017-2019?  
* Are away teams making three pointers at the same rate in 2020 as 2017-2019?  
* Are away teams making free throws at the same rate in 2020 as 2017-2019?

# IV. Analysis and Results

Measure | 2020 | 2017-19 | 95% CI | Test Statistic | P-value 
--------|------|---------|--------|----------------|--------
Home Team Win % | .482 | 0.613 | (-0.255, -0.0075) | 2.091 | 0.0365
Home Team Scoring | 110.06 | 108.103 | (-1.253, 5.168) | 1.195 | 0.2321
Away Team Scoring | 109.06 | 104.025 | (2.083, 7.988) | 3.3424 | 0.00083
Home 2 Pt% | 0.523 | 0.515 | (-0.011, 0.025) | 0.783 | 0.4335
Home 3 Pt% | 0.363 | 0.357 | (-0.014, 0.027) | 0.563 | 0.5733
Home FT% | 0.793 | 0.774 | (-0.001, 0.041) | 1.817 | 0.069
Away 2 Pt% |  0.536 | 0.504 | (0.0147, 0.0503) | 3.587 | 0.00033
Away 3 Pt% | 0.357 | 0.346 | (-0.01, 0.031) | 0.983 | 0.3256
Away FT% | 0.783 | 0.777 | (-0.017, 0.027) | 0.531 | 0.6601

*Note: P-values still significant after Bonferroni correction(multiplying p-value 9 for each test conducted) are denoted by highlighter yellow. Those significant before, but not after Bonferroni correction are denoted by peach.*  
	Starting from the top of the table, we see a statistically significant decrease in home win percentage in 2020 from 2017-19, with p-value of 0.0365. This confirms our belief that home-court advantage was lost in the 2020 NBA playoffs. However, after accounting for multiple tests using Bonferroni correction the p-value is no longer significant. So, we may only cautiously say there is evidence that home-court advantage was not a factor in 2020.

|	Home team performance did not seem to be negatively impacted by losing home-court advantage like expected. Home scoring, two point, three point shooting and  free throw shooting all show no significant difference, on average, between 2020 vs. 2017-19. Home teams do not appear to have any evidence suggesting they played at a lower level in 2020 than they did in previous years when they had home-court advantage.

	Away teams saw more of an impact than home teams. For starters, they saw a significant increase in mean points per game by about 5 points (95% CI: 2.083, 7.988). This is important to note because like we said before it’s scoring points that allows teams to win basketball games, so this is evidence that away teams performed better in the bubble. Likewise, the away team two point shooting efficiency increased by an average of about 0.03 (95% CI: 0.0147, 0.0503). However, unlike two point shooting, away teams did not see a statistically significant difference in three point and free throw shooting. Unlike home teams, away teams do have evidence of change in performance in the bubble. The away teams seemed to perform better than they would under normal conditions as a visiting team.

# V. Discussion 
|	We definitely see an impact of playing the 2020 NBA playoffs in the bubble, there were some minor surprises in the results of the study, but overall it seemed that away teams fared better in the bubble than previous years. Starting from the dip in home winning percentage to below .482 it’s clear something was different. Although the results were not significant after a Bonferonni correction it’s still important to consider and understand that home teams seemed to struggle to win compared to normal conditions.  
|	Moving on to total scoring. As previously stated, this is perhaps the most important measure of a basketball game given that it is what ultimately decides the winner. So these conclusions are particularly interesting and valuable. While the home teams in 2020 did see a slight uptick in average scoring it was not large enough to be statistically significant. However, away team average scoring did increase by a statistically significant amount. This goes hand in hand with our intuition and conclusion about the home winning percentage decreasing. If away teams are scoring significantly more and home teams are not, then we expect to see away teams winning a larger amount of games. This may give a little more reason to believe the conclusion that there was a significant decrease in home winning percentage, despite failing to be significant after Bonferroni correction.  Also, these conclusions give reason to believe that home-court advantage stems mainly from adverse effects on the visiting team.  
|  So away teams are scoring more points, where are those points coming from? The initial thought should be shooting percentage. It seems pretty obvious if you shoot better from the field you are going to score more points. Here we found that away teams are shooting significantly better from two-point shots. However, we know the three point shooting and free-throw shooting were not significantly different. Also, it is important to remember that there’s other factors besides scoring that may affect outcome of games, but were not considered in this study. Another future study may want to look into other potential factors to help understand and expand on how away teams improved.  
|  Surprisingly, it appeared for the most part that home team performance was unaffected by playing in the bubble. They saw some slight insignificant increases in total scoring, two point shooting, three point shooting and free throw shooting. The home teams did not appear to perform any differently, in terms of scoring and scoring efficiency, while playing on a neutral court.  
|  Overall, it appears that away teams were able to close the gap with home teams with home-court advantage removed. Away teams significantly increased their scoring, partly due to the fact of increased two point shooting efficiency. However, there are other potential sources of added scoring worth researching. For example, turnovers, steals, assist, rebounds, and many more game statistics. There are plenty of other possibilities besides just shooting efficiency to pick through looking for more possible sources of added points for away teams. Ultimately, home teams were not largely impacted in terms of their overall performance. It seems their failure to win in 2020, is more due to better away performance than any decrease in their own performance. In fact, the only significant change for home teams found in this study was actually an improvement.

# VI. Conclusion
|	Separating the effects of the home-court advantage into home effects and away effects allowed for some interesting new insights. Previously, we knew that on average home teams outperformed away teams. It was less clear whether it was from positive effects on the home team or negative effects on the road team or perhaps a bit of both. The biggest takeaway from this study is the main source of home court-advantage is the negative effects playing on the road away teams face. Home court advantage may better be categorized as an away team disadvantage. In 2020 there wasn’t any evidence of regression for home team performance, based on the performance measures used, despite being stripped of home-court advantage. Yet, home teams lost about 13% more of games in the 2020 playoffs than the typical average. And this was because of the improvement of away teams. No longer having to face the struggle of traveling, pressure from opposing fans, or playing on an unfamiliar court, teams saw an improvement in their play. They found a way to put more points on the board and ultimately win more. At least some of that improvement came from significantly higher two point efficiency. Future studies may want to use the 2020 NBA bubble and compare vs previous years using other performance measures. This will further help explain what is lost in the performance of away teams when they travel to opposing arenas. This study is only the beginning of possibilities for studies using the 2020 NBA bubble as a case study for home-court advantage.

##References
Aschburner, Steve.(2020) “How Orlando's Neutral Site Will Impact Teams, Referees and Games.” NBA.com, National Basketball Association,        www.nba.com/article/2020/06/22/season-restart-home-court-advantage-no-more.

Harville, D., & Smith, M. (1994). The Home-Court Advantage: How Large Is It, and Does It Vary from Team to Team? The American Statistician, 48(1), 22-28. doi:10.2307/2685080

Kotecki, Jason '14 (2014) "Estimating the Effect of Home Court Advantage on Wins in the  NBA," The Park Place Economist: Vol. 22 Available at: http://digitalcommons.iwu.edu/parkplace/vol22/iss1/13

### 2020-10-19 FirstDraftRMarkdown.Rmd
Add files via upload

@@
 Starting from the dip in home winning percentage to below .482 it’s clear something was different.
-Although the results were not significant after a Bonferonni correction it’s still important to consider and understand that home teams seemed to struggle to win compared to normal conditions.
+Although the results were not significant after a Bonferroni correction it’s still important to consider and understand that home teams seemed to struggle to win compared to normal conditions.
 |	Moving on to total scoring.

### 2020-10-26 Manuscript/NBABubbleAnalysis.Rmd
Add files via upload

---
title: "Analyzing the Effects of the NBA COVID Bubble on the NBA Playoffs"
subtitle: "A Case Study for Home-Court Advantage"
author: |
  | Michael Price
  | michael.price@uconn.edu
  | Department of Statistics, University of Connecticut
  | Dr. Jun Yan
  | jun.yan@uconn.edu
  | Department of Statistics, University of Connecticut

date: "`r format(Sys.time(), '%d, %B, %Y')`"
documentclass: article
papersize: letter
fontsize: 12pt
bibliography: citations.bib
biblio-style: asa
output: pdf_document
---
#Abstract: 
The 2020 NBA playoffs were played inside of a bubble in Disney World because of
the COVID-19 pandemic. This meant that there would be no fans in attendance, 
games played on neutral courts and no traveling for teams which in theory 
removes home-court advantage from the games. This idea was a hot topic of 
discussion as analysts and fans debated the possible effects it may have on the
outcome of games. Home-court advantage has historically played a huge role in 
NBA playoff series outcomes. Thus, to study the effects of the bubble, and 
home-court advantage, the 2020 season is compared to the 2017, 2018, and 2019
playoffs. While many factors contribute to the outcome of games, points scored
is the deciding factor of who wins games so scoring will be the primary focus 
of this study. The specific measures of interest are team scoring totals and 
team shooting percentage on two pointers, three pointers, and free throws. 
Comparing these measures for home teams and away teams in 2020 vs. 2017-2019
shows how the 2020 playoffs favored away teams more than normal, particularly
with two point shooting and total scoring.

```{r setup, echo = FALSE, message =  FALSE, warning = FALSE}
library(BSDA)

knitr::opts_chunk$set(fig.width = 5, fig.height = 4, dpi = 300,
out.width = "90%", fig.align = "center")
```
\newpage
# I. Introduction  
Home-court advantage is often discussed in sports circles as a contributing 
factor to the outcome of games. It is well-known that the home team typically
benefits from some competitive edge from playing at their home-court, resulting
in a better chance of winning. Thus, the NBA playing the 2020 playoffs in a 
bubble brought a great deal of concern for fans, teams, journalists, and others. 
Steve Aschburner [@Aschburner] discusses details about anticipated effects in 
his article “How Orlando's Neutral Site Will Impact Teams, Referees and Games”. 
Aschburner shares concerns from former players, coaches and other experts about
the potential effects of removing home-court advantage. Aschburner notes that the
NBA did make attempts to recreate the effects by putting the "home” team logo on
the court and allowing the "home" team to play crowd noise and music, but most 
people doubted these small attempts would recreate a true playoff atmosphere. 
During the 2020 NBA playoffs, home teams only won about 48.2% of the games. This
is lower than normal, which Aschburner claims usually floats around 60%. This 
shift in the home team winning percentage surely indicates the opportunity for
thorough investigation.

So, what happened? Did the home teams fail to perform up to normal standards 
without the help of home-court advantage? Were away teams able to rise to 
the occasion and perform better not having to deal with the headache of going
on the road? This paper seeks to answer that using scoring totals and shooting
percentages as indicators of team performance. This will deepen understanding
of how home-court advantage affects home and away teams in the NBA.

This study is quite different from earlier NBA home-court advantage studies. 
By using the neutral site games of 2020 we will get to compare home and away
performance to a control. Typically, studies just compare home vs away 
performance. These studies don’t separate the effects of home-court advantage
into the specific effect on the home team and the specific effect on the away
team. They show that home teams outperform away teams, but not if this is a 
result of home teams overperforming or away teams underperforming because of 
home-court advantage. Some of these studies are reviewed in greater detail in
Section II Literature Review.

This study will compare home team performance in 2020 at a neutral site with 
no fans vs. 2017-19 playoffs with fans. Likewise, away team performance in 2020
at a neutral site with no fans vs. 2017-19 playoffs with fans. By comparing home
teams in 2020 to home teams in 2017-19 and away teams in 2020 to away teams in 
2017-19, I can add a new perspective to the field of research. This will allow for
a more accurate understanding of the effects of home-court advantage on home and
away teams in the NBA. We won’t just see that home-court advantage helps home 
teams outperform away teams, but this study will actually clearly separate the 
effects of home-court advantage on home teams and away teams performance individually.

9 statistical tests are run to understand the differences in 2020 vs. earlier 
years. First, whether or not the difference between home win percentage in 2020
and 2017-19 is statistically significant. To confirm the difference we see in the
winning percentage is truly significant. Then we can look for differences in home
scoring in 2020 vs 2017-2019. Similarly, we can do the same test, but for 
differences in away scoring in 2020 vs 2017-2019. Also, differences in team 
shooting(for two pointers, three pointers, and free throws) from 2020 vs 2017-2019
for both home and away teams. The results from these simple tests will bring a new
understanding of how home-court advantage is impacting games by altering the performance
of the home and away teams.

# II. Literature Review  
There is tons of published literature on the effects of home-court advantage. 
[@Kotecki] studies the impact of home-court advantage on winning in the NBA. 
Kotecki proves the existence of home-court advantage using performance-based
statistics, specifically field goal percentage, free throw percentage, and 
points. All of which he proves significantly improved from home-court advantage, 
indicating home-court helps teams play better. This conclusion comes from 
comparing home performance vs. away performance in games.

[@Carron1992] cite 4 main game location factors for home and away teams.
First, the crowd factor, which is the impact of fans cheering. Second, learning
factors, which they categorize as the advantage from home teams from playing at
a familiar venue. Third, travel factors, the idea that away teams may face 
fatigue and jet lag from traveling. Last, rule factors, which says that home teams
may benefit from some advantages in rules and officiating. They acknowledge that
these factors would all be removed if games were played at a neutral site even 
if one team was designated as “home team”. This study was reviewed a decade later
by [@Carron2005]. The 2005 review goes over the new findings from studies 
about the significance of these 4 game location factors. Since 1992 they found 
that results on these 4 factors are mixed. However, there is some evidence 
supporting crowd and travel factors impact games in the NBA. There is less 
evidence suggesting learning and rule factors impact the NBA. One interesting 
finding cited by [@Carron2005] is that in the absence of crowds result 
in overall performance increases.

[@Cao] study the effects of pressure on performance in the NBA. They use free
throws as their measure of interest, and research home and away factors in the
study. They expected that home fans could distract and put pressure on opposing
players to make free throws, which would negatively impact away team shooting. 
Likewise, the home team would see a positive effect on free throw shooting. 
However, they find insignificant evidence that home status has a substantial 
impact on missing from the free throw line.

[@Greer] focuses his study on the influence of spectator booing on home-court
advantage in basketball. The three methods of performance used in this study 
were scoring, violations and turnovers. This study was conducted using the men’s
basketball programs at two large universities. The study finds that social 
support, like booing, is an important contributor to home-court advantage. Greer
explains, whether the influence is greater on visiting team performance or
referee calls is less clear. However, the data does seem to lean slightly in
favor of affecting visiting team performance.

[@Harris] use two point shots, three-point shots and free throws as measures of
interest to study home-court advantage. They find that two point shots are the
strongest predictor of home-court advantage. They suggest that evidence supports
that home teams should try to shoot more two point shots and force their opponent
to take more two point shot attempts. This strategy will maximize the benefits
of home-court advantage and give them the best chance to win.

College basketball, as noted before, does offer some opportunity to study 
home-court  vs. neutral court. One study conducted by [@Harville] studied the
effect of home-court advantage using the 1991-1992 college basketball season. 
Unlike the NBA, it is not uncommon to have a few games played at neutral sites
during the college basketball season. This allowed them to construct two samples,
one of home teams and one of neutral teams. They formulate their study as a 
regression equation predicting the expected difference in score for home teams. 
They set up their study to find if the home teams won games by more points when 
they had home-court advantage vs. when playing at neutral court. This study 
succeeds in proving that home-court advantage exists.

```{r data, include=FALSE}
#Data Collection
game_log_2020 <- read.csv(file="Game_Log_2020.csv", header=TRUE)
game_log_2017_2019 <- read.csv(file="2017-2019_game_data.csv", header=TRUE)

#Data Cleaning
game_log_2020$W.L <- ifelse(game_log_2020$W.L=="W", 1, 0)
game_log_2017_2019$W.L <- ifelse(game_log_2017_2019$W.L=="W", 1, 0)

game_log_2020_home <- game_log_2020[game_log_2020$Home.Away=="FALSE",]
game_log_2020_away <- game_log_2020[game_log_2020$Home.Away=="TRUE",]
game_log_2017_2019_home <- game_log_2017_2019[game_log_2017_2019$Home.Away=="FALSE", ]
game_log_2017_2019_away <- game_log_2017_2019[game_log_2017_2019$Home.Away=="TRUE", ]

years <- 2010:2020
winpct <- c(.6707, .667, .679, .635, .562, .593, .6744, .5696, .707, .561, .482)
```

# III. Data Collection and Methods  
The 2020 bubble provides a new and exciting opportunity to study home-court
advantage for the NBA. Unlike college basketball, aside from a few 
exhibition/preseason games, the NBA always has a home and away team. So, for
the first time in NBA history the bubble allows NBA home and away performance
to be compared vs a control/neutral field. The NBA bubble, as a neutral court,
removed all 4 possible game location factors impacting home-court advantage 
hypothesized by [@Carron2005]. The NBA bubble featured 8 seeding games then a
standard playoff format. The focus of this study was the play during the playoff
games since it followed the standard playoff format and can easily be compared
back to other playoffs. For this study, the 2020 playoffs were compared against
the three previous playoffs collectively. It is important to make sure that any
observed differences are truly from a home-court advantage. So, in order to control
for the changing play style of the NBA, we will limit the study to 2020 vs 2017-2019.
The reason being the faster pace play and more common use of the three-point shot 
in modern basketball. If we used data from say 10 years ago, or earlier, observed
differences may not be from effects of the NBA bubble, but rather from the effects
of drastic changes in the style of play between the seasons. However, basketball
evolves slow enough that we can reasonably assume 2017-2019 are at least very 
close in pace and playing style to 2020.

Data was collected from the official NBA website. The main variables of interest
are whether or not the home team won, scoring totals for home and away teams, and
shooting percentages for home and away teams on two pointers, three pointers, and
free throws. These variables were very popular and frequently used in the related
literature discussed earlier. While many other measures could be used for measuring
the outcome of the game and team performance, scoring seemed to be the most 
important. The winner of a game is determined by who scores more points. So, 
naturally when studying the outcome of winning, changes in scoring and scoring 
efficiency are important to understand.

Home-court advantage is the basic idea that the home team is more likely to win.
So laying a foundation of typical home-court advantage is crucial. Before 
focusing on the 2017 to 2020 playoffs we can take a quick look at home team win
percentages since 2010. Notice on the graph below in Figure 1, the
10 years before 2020, the home team winning percentage ranged from around 0.56
to 0.7 and never dipped below .5. The 2020 bubble, however, broke this historic
pattern dipping down below .5. Foreshadowing, the confirmation of the expectation
that the effect of home-court advantage was removed in the 2020 playoffs.

```{r Fig1, echo=FALSE, fig.cap = "Winning percentage of NBA home teams in the playoffs since 2010, the green line denotes .500. ", fig.align='center'}
plot(years, winpct, ylim = c(0,1), title("NBA Home Playoff Win % Since 2010"), col = ifelse(winpct==.482, "Red", "Black"), pch=16)
text(years, winpct, labels = winpct, cex = 0.5, pos = 1)
text(2010, 0.51, labels = "0.5", cex = 0.6, col = "Green")
abline(h=0.5, col = "Green")
```

Moving on to the main focus of the study, comparing 2020 to 2017-2019. First, 
a brief look at the distribution of scoring depicted below in Figure 2.
The green histograms indicate home and red histograms indicate away scoring. The
top row is scoring for 2020 and the bottom is scoring for 2017-2019. All histograms
seem to be fairly normally distributed which is important for statistical analysis.
It’s hard to see any clear differences between years, but it’s important to look at
and understand the distribution of measures before moving onto the analysis.

```{r Figtwo, echo=FALSE, fig.cap="Shows the home(green) and away (red) scoring distirbutions for 2020(top) and 2017-2019(bottom)", fig.align='center'}
opar<-par(no.readonly = TRUE)
par(mfrow=c(2,2))
hist(game_log_2020_home$PTS, xlab = "Points", col = "green",
     xlim = c(60, 160), main = "Home 2020")

hist(game_log_2020_away$PTS,  xlab = "Points", col = "red",
     xlim = c(60, 160), main = "Away 2020")

hist(game_log_2017_2019_home$PTS, xlab = "Points", col="green",
     xlim = c(60, 160), main = "Home 2017-2019" )

hist(game_log_2017_2019_away$PTS, xlab = "Points", col = "red",
     xlim = c(60, 160), main = "Home 2017-2019", breaks = 8 )

par(opar)
```

Figure 3 shows home shooting for two pointers, three pointers and
free throws for 2020(top) vs. 2017-19(bottom). The histograms appear to be fairly
similarly distributed between 2020 and 2017-19. Likewise, Figure 4, shows the same
data except for away teams. Just like Figure 3, Figure 4 histograms don’t show any
obvious differences. The normality of the distributions for both figures 3 and 4
are important for meeting the required assumptions for statistical analysis.

(ref:cap-Figthree) Home shooting percentages for two pointers, three pointers and free throws for 2020(top) vs. 2017-19 (bottom)
```{r Figthree, echo=FALSE, fig.cap="(ref:cap-Figthree)"}
opar<-par(no.readonly = TRUE)
par(mfrow=c(2,3))

hist(game_log_2020_home$X2P., xlab = "Shooting Percentage", col= "green",
     main = "2020 Home 2 pointers", xlim = c(20, 70))

hist(game_log_2020_home$X3P., xlab = "Shooting Percentage", col= "green",
     main = "2020 Home 3 pointers", xlim = c(10, 70))

hist(game_log_2020_home$FT., xlab = "Shooting Percentage", col= "green",
     main = "2020 Home Free Throws", xlim = c( 50, 100))

hist(game_log_2017_2019_home$X2P., xlab = "Shooting Percentage", col= "green",
     main = "2017-2019 2 pointers", xlim = c(20, 70))

hist(game_log_2017_2019_home$X3P., xlab = "Shooting Percentage", col= "green",
     main = "2017-2019 3 pointers", xlim = c(10, 70))

hist(game_log_2017_2019_home$FT., xlab = "Shooting Percentage", col= "green",
     main = "2017-2019 Free Throws", xlim = c( 50, 100))

par(opar)
```

```{r Fig4, echo=FALSE, fig.cap="Away shooting percentages for two pointers, three pointers and free throws for 2020(top) vs. 2017-19 (bottom)", fig.align='center'}
opar<-par(no.readonly = TRUE)
par(mfrow=c(2,3))
hist(game_log_2020_away$X2P., xlab = "Shooting Percentage", col= "red", 
     main = "2020 Away 2 pointers", xlim= c(30, 80))

hist(game_log_2020_away$X3P., xlab = "Shooting Percentage", col= "red", 
     main = "2020 Away 3 pointers", xlim = c(10, 60))

hist(game_log_2020_away$FT., xlab = "Shooting Percentage", col= "red", 
     main = "2020 Away Free Throws", xlim = c(40, 100))

hist(game_log_2017_2019_away$X2P., xlab = "Shooting Percentage", col= "red",
     main = "2017-19 Away 2 pointers", xlim= c(30, 80))

hist(game_log_2017_2019_away$X3P., xlab = "Shooting Percentage", col= "red",
     main = "2017-19 Away 3 pointers", xlim = c(10, 60))

hist(game_log_2017_2019_away$FT., xlab = "Shooting Percentage", col= "red",
     main = "2017-19 Away Free Throws", xlim = c(40, 100))
par(opar)
```

  Given the approximate normality of data, 9 different z-tests were performed in order to test the effects of the COVID bubble on the 2020 NBA playoffs. The 9 questions answered were:  
* Is the home team winning percentage in 2020 different than that it was in 2017-2019?  
* Is the average home team scoring different in 2020 than it was over 2017-2019?  
* Is the average away team scoring different in 2020 than it was over 2017-2019?  
* Are home teams making two pointers at the same rate in 2020 as 2017-2019?  
* Are home teams making three pointers at the same rate in 2020 as 2017-2019?  
* Are home teams making free throws at the same rate in 2020 as 2017-2019?  
* Are away teams making two pointers at the same rate in 2020 as 2017-2019?  
* Are away teams making three pointers at the same rate in 2020 as 2017-2019?  
* Are away teams making free throws at the same rate in 2020 as 2017-2019?

# IV. Analysis and Results
```{r, echo=FALSE}
#Analysis
one <- prop.test(x=c(sum(game_log_2020_home$W.L), sum(game_log_2017_2019_home$W.L)), 
                 n=c(nrow(game_log_2020_home), nrow(game_log_2017_2019_home)), correct=FALSE)
rowone <- c(as.numeric(one$estimate[1]), as.numeric(one$estimate[2]), one$conf.int,
          sqrt(as.numeric(one$statistic)), one$p.value)

two <- z.test(x=game_log_2020_home$PTS, y=game_log_2017_2019_home$PTS, sigma.x = 
                sd(game_log_2020_home$PTS), sigma.y = sd(game_log_2017_2019_home$PTS))
rowtwo <- c(as.numeric(two$estimate[1]), as.numeric(two$estimate[2]), 
            two$conf.int, two$statistic, two$p.value)

three <- z.test(x=game_log_2020_away$PTS, y=game_log_2017_2019_away$PTS, sigma.x = 
                  sd(game_log_2020_away$PTS), sigma.y = sd(game_log_2017_2019_away$PTS))
rowthree <- c(as.numeric(three$estimate[1]), as.numeric(three$estimate[2]), 
              three$conf.int, three$statistic, three$p.value)

four <- prop.test(x=c(sum(game_log_2020_home$X2PM), sum(game_log_2017_2019_home$X2PM)), 
                  n=c(sum(game_log_2020_home$X2PA), sum(game_log_2017_2019_home$X2PA)))
rowfour <- c(as.numeric(four$estimate[1]), as.numeric(four$estimate[2]), 
             four$conf.int, sqrt(as.numeric(four$statistic)), four$p.value)

five <- prop.test(x=c(sum(game_log_2020_home$X3PM), sum(game_log_2017_2019_home$X3PM)), 
                  n=c(sum(game_log_2020_home$X3PA), sum(game_log_2017_2019_home$X3PA)))
rowfive <- c(as.numeric(five$estimate[1]), as.numeric(five$estimate[2]), 
             five$conf.int,  sqrt(as.numeric(five$statistic)), five$p.value)

six <- prop.test(x=c(sum(game_log_2020_home$FTM), sum(game_log_2017_2019_home$FTM)), 
                 n=c(sum(game_log_2020_home$FTA), sum(game_log_2017_2019_home$FTA)))
rowsix <- c(as.numeric(six$estimate[1]), as.numeric(six$estimate[2]), six$conf.int,
           sqrt(as.numeric(six$statistic)), six$p.value)

seven <- prop.test(x=c(sum(game_log_2020_away$X2PM), sum(game_log_2017_2019_away$X2PM)), 
                   n=c(sum(game_log_2020_away$X2PA), sum(game_log_2017_2019_away$X2PA)))
rowseven <- c(as.numeric(seven$estimate[1]), as.numeric(seven$estimate[2]), 
              seven$conf.int,  sqrt(as.numeric(seven$statistic)), seven$p.value)

eight <- prop.test(x=c(sum(game_log_2020_away$X3PM), sum(game_log_2017_2019_away$X3PM)), 
                   n=c(sum(game_log_2020_away$X3PA), sum(game_log_2017_2019_away$X3PA)))
roweight <- c(as.numeric(eight$estimate[1]), as.numeric(eight$estimate[2]), 
              eight$conf.int,  sqrt(as.numeric(eight$statistic)), eight$p.value)

nine <- prop.test(x=c(sum(game_log_2020_away$FTM), sum(game_log_2017_2019_away$FTM)), 
                  n=c(sum(game_log_2020_away$FTA), sum(game_log_2017_2019_away$FTA)))
rownine <- c(as.numeric(nine$estimate[1]), as.numeric(nine$estimate[2]), 
             nine$conf.int,  sqrt(as.numeric(nine$statistic)), nine$p.value)

titles <- c(2020, "2017-19", "95% CI Low", "95% CI Up", "Test Statistic", "P-value")
rows <- c("Home Team Win %", "Home Team Scoring", "Away Team Scoring", "Home Team 2P%", "Home Team 3P%", "Home Team FT%", "Away Team 2P%", "Away Team  3P%", "Away Team FT%")
results <- data.frame(rowone, rowtwo, rowthree, rowfour, rowfive, rowsix,
                    rowseven, roweight, rownine, stringsAsFactors = FALSE)
resultsfinal <- data.frame(t(results))
resultsfinal <- round(resultsfinal, 4)
colnames(resultsfinal) <- titles
row.names(resultsfinal) <- rows

```

```{r table, echo = FALSE}
knitr::kable(resultsfinal, caption = 'The results from the 9 statistical tests', align = "cccccc")
```

Starting from the top of the Table 1, we see a statistically significant change
in home win percentage in 2020 from 2017-19, with p-value of 0.0365. The 95% 
confidence interval of (-0.255, -0.0075) confirms our belief that home-court 
advantage was lost in the 2020 NBA playoffs. However, after accounting for 
multiple tests using Bonferroni correction the p-value is no longer significant.
So, we may only cautiously say there is evidence that home-court advantage was 
not a factor in 2020.

Home team performance did not seem to be negatively impacted by losing home-court
advantage like expected. Home scoring, two point, three point shooting and free 
throw shooting all show no significant difference, on average, between 2020 vs. 
2017-19. Home teams do not appear to have any evidence suggesting they played at 
a lower level in 2020 than they did in previous years when they had home-court advantage.

Away teams saw more of an impact than home teams. For starters, they saw a 
significant increase in mean points per game by about 5 points (95% CI: 2.083, 7.988)
Likewise, the away team two point shooting efficiency increased by an average of
about 0.03 (95% CI: 0.0147, 0.0503). However, unlike two point shooting, away 
teams did not see a statistically significant difference in three point and free
throw shooting. Overall, away teams have evidence of change in performance in 
the bubble. The away teams seemed to perform better than they would under normal
conditions as a visiting team.

# V. Discussion 
Generally it seemed that away teams fared better in the 2020 NBA playoff bubble
than previous years on the road. Starting from the dip in home winning percentage
to below .482 it’s clear something was different. Although the difference was not
significant after a Bonferroni correction it’s still important to consider and
understand that home teams seemed to struggle to win compared to normal conditions.
Compared to [@Kotecki]who finds home teams consistently have a significantly
better record than away teams boasting about a 60.5% win percentage in his sample,
the 48.2% home winning percentage of 2020 home teams is a shift from normal.
It’s clear that home teams did not benefit from the usual advantages provided by
being the home team.

While the home teams in 2020 did see a slight uptick in average scoring it was
not large enough to be statistically significant. However, away team average 
scoring did increase by a statistically significant amount. This goes hand in 
hand with our intuition and conclusion about the home winning percentage 
decreasing. If away teams are scoring significantly more and home teams are not,
then we expect to see away teams winning a larger amount of games. This may give
more reason to believe the conclusion that there was a significant decrease in
home winning percentage in 2020, despite failing to be significant after Bonferroni
correction. Only away team scoring being significantly impacted by playing on
a neutral court indicates that home-court advantage stems mainly from adverse
effects on the visiting team.

An interesting finding is all shooting and scoring numbers for both home and
away teams did make at least small increases. Although these increases weren’t
all significant these increases are exactly what is reported in [@Carron2005]
when they explain how evidence suggests that teams perform better with the absence
of fans. This is important because it coincides with our conclusion that home-court
advantage mostly plays into games by negatively impacting away teams. If fans cause
overall performance to drop, then home court advantage must come from a bigger drop
in away performance than the drop in home performance. This is why away teams were 
able to close the gap with home teams with home-court advantage removed.

Separating the effects of the home-court advantage into home effects and away
effects allowed for some interesting new insights. Previously, we knew that on
average home teams outperformed away teams. It was less clear whether it was 
from positive effects on the home team or negative effects on the road team or
perhaps a bit of both. The biggest takeaway from this study is the main source
of home-court-advantage is the negative effects playing on the road away teams 
face. In 2020 there wasn’t any evidence of regression for home team performance,
based on the performance measures used, despite being stripped of home-court
advantage. Yet, home teams lost about 12% more of games in the 2020 playoffs 
than the typical average. This was because of the improvement of away teams. 
No longer having to face the struggle of traveling, pressure from opposing fans,
or playing on an unfamiliar court, teams saw an improvement in their play and an
increase in winning. The improvement of away teams confirms a proposition from 
[@Greer)] who believed that the positive social impact of crowds benefiting 
home teams may be a result of inhibiting away teams.

At least some of that improvement from away teams came from significantly higher 
two point efficiency. This corresponds with the conclusion from [@Harris],
where they found home teams are best suited to capitalize on advantages from two point
shots. Normally, by shooting more 2 themselves and forcing away teams to shoot more 2
they home team benefits from increasing home-court advantage. However, with away teams
significantly improving two point shooting in the bubble this strategy was no longer 
viable and home-court advantage disappeared.

Future studies may want to use the 2020 NBA bubble and compare vs previous years
using other performance measures. For example, turnovers, steals, assist, rebounds,
and many more game statistics. There are plenty of other possibilities besides
just shooting efficiency to pick through looking for more possible sources of
added points for away teams.This will further help explain what is lost in the
performance of away teams when they travel to opposing arenas. This study is only
the beginning of possibilities for studies using the 2020 NBA bubble as a case study 
for home-court advantage.

Unfortunately, this study is limited by a one time sample. It is unknown if there
will ever be another NBA playoffs played in this fashion. It seems unlikely that 
these conditions will ever be repeated. That being said, it does not seem like a 
follow-up study using the same measures with a different sample will ever be
possible. If possible, that type of study could help strengthen the conclusions
made in this paper.

#References

### 2020-10-26 Simple Code for Graphs and Analysis.R
Delete Simple Code for Graphs and Analysis.R

@@ -1,63 +1,0 @@
-#Data Collection
-game_log_2020<-read.csv("Desktop/Thesis/2020 game data.csv", header=T, sep = ",")
-game_log_2017_2019<-read.csv("Desktop/Thesis/2017 2019 game data.csv", header=T)
-
-#Data Cleaning
-game_log_2020$W.L<-ifelse(game_log_2020$W.L=="W", 1, 0)
-game_log_2017_2019$W.L<-ifelse(game_log_2017_2019$W.L=="W", 1, 0)
-
-game_log_2020_home<-game_log_2020[game_log_2020$Home.Away=="FALSE",]
-game_log_2020_away<-game_log_2020[game_log_2020$Home.Away=="TRUE",]
-game_log_2017_2019_home<-game_log_2017_2019[game_log_2017_2019$Home.Away=="FALSE",]
-game_log_2017_2019_away<-game_log_2017_2019[game_log_2017_2019$Home.Away=="TRUE",]
-
-#Home Winning Percentage Over Time
-years<-2010:2020
-winpct<-c(.6707, .667, .679, .635, .562, .593, .6744, .5696, .707, .561, .482)
-
-plot(years, winpct, ylim = c(0,1), title("NBA Home Playoff Win % Since 2010"), col = ifelse(winpct==.494, "Red", "Black"), pch=16)
-text(years, winpct, labels = winpct, cex = 0.5, pos = 1)
-text(2010, 0.51, labels = "0.5", cex = 0.6, col = "Green")
-abline(h=0.5, col = "Green")
-
-#histograms
-opar<-par(no.readonly = TRUE)
-par(mfrow=c(2,2))
-hist(game_log_2020_home$PTS, xlab = "Points", col = "green",  xlim = c(60, 160), main = "Home 2020")
-hist(game_log_2020_away$PTS,  xlab = "Points", col = "red",  xlim = c(60, 160), main = "Away 2020")
-hist(game_log_2017_2019_home$PTS, xlab = "Points", col="green",  xlim = c(60, 160), main = "Home 2017-2019" )
-hist(game_log_2017_2019_away$PTS, xlab = "Points", col = "red",  xlim = c(60, 160), main = "Home 2017-2019", breaks = 8 )
-par(opar)
-
-opar<-par(no.readonly = TRUE)
-par(mfrow=c(2,3))
[…]

### 2020-10-27 Manuscript/NBABubbleAnalysis.Rmd
Update NBABubbleAnalysis.Rmd

@@
 focusing on the 2017 to 2020 playoffs we can take a quick look at home team win
-percentages since 2010. Notice on the graph below in Figure 1, the
-10 years before 2020, the home team winning percentage ranged from around 0.56
-to 0.7 and never dipped below .5. The 2020 bubble, however, broke this historic
-pattern dipping down below .5. Foreshadowing, the confirmation of the expectation
-that the effect of home-court advantage was removed in the 2020 playoffs.
+percentages since 2010. Notice in Figure 1, the10 years before 2020, the home team
+winning percentage ranged from around 0.56 to 0.7 and never dipped below .5. The 
+2020 bubble, however, broke this historicpattern dipping down below .5. Foreshadowing, 
+the confirmation of the expectationthat the effect of home-court advantage was removed
+in the 2020 playoffs.
 
@@
 Moving on to the main focus of the study, comparing 2020 to 2017-2019. First, 
-a brief look at the distribution of scoring depicted below in Figure 2.
+a brief look at the distribution of scoring depicted in Figure 2.
 The green histograms indicate home and red histograms indicate away scoring. The
@@
 
-  Given the approximate normality of data, 9 different z-tests were performed in order to test the effects of the COVID bubble on the 2020 NBA playoffs. The 9 questions answered were:  
+  Given the approximate normality of data, 9 different z-tests were performed in 
+  order to test the effects of the COVID bubble on the 2020 NBA playoffs. The 9 questions were:  
 * Is the home team winning percentage in 2020 different than that it was in 2017-2019?  
@@
 titles <- c(2020, "2017-19", "95% CI Low", "95% CI Up", "Test Statistic", "P-value")
-rows <- c("Home Team Win %", "Home Team Scoring", "Away Team Scoring", "Home Team 2P%", "Home Team 3P%", "Home Team FT%", "Away Team 2P%", "Away Team  3P%", "Away Team FT%")
+rows <- c("Home Team Win %", "Home Team Scoring", "Away Team Scoring", "Home Team 2P%", "Home Team 3P%",
+          "Home Team FT%", "Away Team 2P%", "Away Team  3P%", "Away Team FT%")
 results <- data.frame(rowone, rowtwo, rowthree, rowfour, rowfive, rowsix,

### 2020-10-27 Manuscript/NBABubbleAnalysis.Rmd
Update NBABubbleAnalysis.Rmd

@@
 ---
-#Abstract: 
+Abstract: 
 The 2020 NBA playoffs were played inside of a bubble in Disney World because of
@@
 
-titles <- c(2020, "2017-19", "95% CI Low", "95% CI Up", "Test Statistic", "P-value")
+titles <- c(2020, "2017-19", "95% CI LL", "95% CI UL", "Test Stat.", "P-value")
 rows <- c("Home Team Win %", "Home Team Scoring", "Away Team Scoring", "Home Team 2P%", "Home Team 3P%",

### 2020-10-27 Manuscript/NBABubbleAnalysis.Rmd
Update NBABubbleAnalysis.Rmd

@@
 
-# III. Data Collection and Methods  
+# III. Data and Methods  
 The 2020 bubble provides a new and exciting opportunity to study home-court
@@
 
-(ref:cap-Figthree) Home shooting percentages for two pointers, three pointers and free throws for 2020(top) vs. 2017-19 (bottom)
-```{r Figthree, echo=FALSE, fig.cap="(ref:cap-Figthree)"}
+```{r Figthree, echo=FALSE, fig.cap="Home shooting percentages for two pointers, three pointers and free throws for 2020(top) vs. 2017-19 (bottom)"}
 opar<-par(no.readonly = TRUE)

### 2020-11-02 Manuscript/NBABubbleAnalysis.Rmd
Update NBABubbleAnalysis.Rmd

@@
 papersize: letter
-fontsize: 12pt
+fontsize: 11pt
 bibliography: citations.bib
 biblio-style: asa
-output: pdf_document
+output: bookdown::pdf_document2
+header-includes:
+   - \usepackage{float}
+toc: False
 ---
@@
 library(BSDA)
-
-knitr::opts_chunk$set(fig.width = 5, fig.height = 4, dpi = 300,
-out.width = "90%", fig.align = "center")
+library(ggplot2)
+knitr::opts_chunk$set(fig.width = 4, fig.height = 3, dpi = 300,
+out.width = "90%", fig.align = "center", dev = "pdf", fig.pos = "H")
 ```
@@
 bubble brought a great deal of concern for fans, teams, journalists, and others. 
-Steve Aschburner [@Aschburner] discusses details about anticipated effects in 
-his article “How Orlando's Neutral Site Will Impact Teams, Referees and Games”. 
+Steve Aschburner discusses details about anticipated effects in his article 
+“How Orlando's Neutral Site Will Impact Teams, Referees and Games” @Aschburner. 
 Aschburner shares concerns from former players, coaches and other experts about
@@
 There is tons of published literature on the effects of home-court advantage.
-[@Kotecki] studies the impact of home-court advantage on winning in the NBA. 
-Kotecki proves the existence of home-court advantage using performance-based
+Many of which mainly focus on the effects of home-court advantage using shooting
+in the NBA. @Kotecki studies the impact of home-court advantage on winning in the
+NBA. Kotecki proves the existence of home-court advantage using performance-based
 statistics, specifically field goal percentage, free throw percentage, and 
@@
 indicating home-court helps teams play better. This conclusion comes from 
-comparing home performance vs. away performance in games.
+comparing home performance vs. away performance in games. @Cao studied the effects
+of pressure on performance in the NBA. They use free throws as their measure of 
+interest, and research home and away factors in the study. They expected that home
+fans could distract and put pressure on opposing players to make free throws,
+which would negatively impact away team shooting. Likewise, the home team
+would see a positive effect on free throw shooting. However, they find insignificant
+evidence that home status has a substantial impact on missing from the free throw 
+line. @Harris use two point shots, three-point shots and free throws as measures
+of interest to study home-court advantage. They find that two point shots are
+the strongest predictor of home-court advantage. They suggest that evidence 
+supports that home teams should try to shoot more two point shots and force
+their opponent to take more two point shot attempts. This strategy will maximize
+the benefits of home-court advantage and give them the best chance to win.
 
-[@Carron1992] cite 4 main game location factors for home and away teams.
-First, the crowd factor, which is the impact of fans cheering. Second, learning
-factors, which they categorize as the advantage from home teams from playing at
-a familiar venue. Third, travel factors, the idea that away teams may face 
-fatigue and jet lag from traveling. Last, rule factors, which says that home teams
-may benefit from some advantages in rules and officiating. They acknowledge that
-these factors would all be removed if games were played at a neutral site even 
-if one team was designated as “home team”. This study was reviewed a decade later
-by [@Carron2005]. The 2005 review goes over the new findings from studies 
-about the significance of these 4 game location factors. Since 1992 they found 
-that results on these 4 factors are mixed. However, there is some evidence 
-supporting crowd and travel factors impact games in the NBA. There is less 
-evidence suggesting learning and rule factors impact the NBA. One interesting 
-finding cited by [@Carron2005] is that in the absence of crowds result 
-in overall performance increases.
-
-[@Cao] study the effects of pressure on performance in the NBA. They use free
-throws as their measure of interest, and research home and away factors in the
-study. They expected that home fans could distract and put pressure on opposing
-players to make free throws, which would negatively impact away team shooting. 
-Likewise, the home team would see a positive effect on free throw shooting. 
-However, they find insignificant evidence that home status has a substantial 
-impact on missing from the free throw line.
-
-[@Greer] focuses his study on the influence of spectator booing on home-court
-advantage in basketball. The three methods of performance used in this study 
-were scoring, violations and turnovers. This study was conducted using the men’s
-basketball programs at two large universities. The study finds that social 
-support, like booing, is an important contributor to home-court advantage. Greer
-explains, whether the influence is greater on visiting team performance or
+Some studies focus less on shooting and more on scoring differences and other metrics.
+For example, @Greer focuses his study on the influence of spectator booing on 
+Home-court advantage in basketball. The three methods of performance used
+in this study were scoring, violations and turnovers. This study was conducted 
+using the men’s basketball programs at two large universities. The study finds that
+social support, like booing, is an important contributor to home-court advantage. 
+Greer explains, whether the influence is greater on visiting team performance or
 referee calls is less clear. However, the data does seem to lean slightly in
-favor of affecting visiting team performance.
-
-[@Harris] use two point shots, three-point shots and free throws as measures of
-interest to study home-court advantage. They find that two point shots are the
-strongest predictor of home-court advantage. They suggest that evidence supports
-that home teams should try to shoot more two point shots and force their opponent
-to take more two point shot attempts. This strategy will maximize the benefits
-of home-court advantage and give them the best chance to win.
-
-College basketball, as noted before, does offer some opportunity to study 
-home-court  vs. neutral court. One study conducted by [@Harville] studied the
+favor of affecting visiting team performance. One study conducted by @Harville studied the
 effect of home-court advantage using the 1991-1992 college basketball season. 
@@
 succeeds in proving that home-court advantage exists.
+
+There are also studies focused on the factors contributing to home-court advantage. 
+@Carron1992 cite 4 main game location factors for home and away teams.
+First, the crowd factor, which is the impact of fans cheering. Second, learning
+factors, which they categorize as the advantage from home teams from playing at
+a familiar venue. Third, travel factors, the idea that away teams may face 
+fatigue and jet lag from traveling. Last, rule factors, which says that home teams
+may benefit from some advantages in rules and officiating. They acknowledge that
+these factors would all be removed if games were played at a neutral site even 
+if one team was designated as “home team”. This study was reviewed a decade later
+by @Carron2005. The 2005 review goes over the new findings from studies 
+about the significance of these 4 game location factors. Since 1992 they found 
+that results on these 4 factors are mixed. However, there is some evidence 
+supporting crowd and travel factors impact games in the NBA. There is less 
+evidence suggesting learning and rule factors impact the NBA. One interesting 
+finding cited by @Carron2005 is that in the absence of crowds result 
+in overall performance increases.
 
@@
 
-game_log_2020_home <- game_log_2020[game_log_2020$Home.Away=="FALSE",]
-game_log_2020_away <- game_log_2020[game_log_2020$Home.Away=="TRUE",]
-game_log_2017_2019_home <- game_log_2017_2019[game_log_2017_2019$Home.Away=="FALSE", ]
-game_log_2017_2019_away <- game_log_2017_2019[game_log_2017_2019$Home.Away=="TRUE", ]
+game_log_2020$Home.Away <- ifelse(game_log_2020$Home.Away == FALSE, "Home", "Away")
+game_log_2017_2019$Home.Away <- ifelse(game_log_2017_2019$Home.Away == FALSE, "Home", "Away")
+
+game_log_2020_home <- game_log_2020[game_log_2020$Home.Away=="Home",]
+game_log_2020_away <- game_log_2020[game_log_2020$Home.Away=="Away",]
+game_log_2017_2019_home <- game_log_2017_2019[game_log_2017_2019$Home.Away=="Home", ]
+game_log_2017_2019_away <- game_log_2017_2019[game_log_2017_2019$Home.Away=="Away", ]
+
+#combineddata
+combined <- rbind(game_log_2020, game_log_2017_2019)
+combined$Year <- as.factor(combined$Year)
+combined <-  na.omit(combined)
 
@@
 removed all 4 possible game location factors impacting home-court advantage 
-hypothesized by [@Carron2005]. The NBA bubble featured 8 seeding games then a
+hypothesized by @Carron1992. The NBA bubble featured 8 seeding games then a
 standard playoff format. The focus of this study was the play during the playoff
@@
 focusing on the 2017 to 2020 playoffs we can take a quick look at home team win
-percentages since 2010. Notice in Figure 1, the10 years before 2020, the home team
+percentages since 2010. Notice in Figure \@ref(fig:Fig1), the10 years before 2020, the home team
 winning percentage ranged from around 0.56 to 0.7 and never dipped below .5. The 
@@
 
-```{r Fig1, echo=FALSE, fig.cap = "Winning percentage of NBA home teams in the playoffs since 2010, the green line denotes .500. ", fig.align='center'}
+(ref:cap-Fig1) "Winning percentage of NBA home teams in the playoffs since 2010, the green line denotes .500"
+```{r Fig1, echo=FALSE, fig.cap = " (ref:cap-Fig1)", fig.align='center'}
 plot(years, winpct, ylim = c(0,1), title("NBA Home Playoff Win % Since 2010"), col = ifelse(winpct==.482, "Red", "Black"), pch=16)
@@
 Moving on to the main focus of the study, comparing 2020 to 2017-2019. First, 
-a brief look at the distribution of scoring depicted in Figure 2.
+a brief look at the distribution of scoring depicted in Figure \@ref(fig:Fig2).
 The green histograms indicate home and red histograms indicate away scoring. The
@@
 
-```{r Figtwo, echo=FALSE, fig.cap="Shows the home(green) and away (red) scoring distirbutions for 2020(top) and 2017-2019(bottom)", fig.align='center'}
-opar<-par(no.readonly = TRUE)
-par(mfrow=c(2,2))
-hist(game_log_2020_home$PTS, xlab = "Points", col = "green",
-     xlim = c(60, 160), main = "Home 2020")
-
-hist(game_log_2020_away$PTS,  xlab = "Points", col = "red",
-     xlim = c(60, 160), main = "Away 2020")
-
-hist(game_log_2017_2019_home$PTS, xlab = "Points", col="green",
-     xlim = c(60, 160), main = "Home 2017-2019" )
-
-hist(game_log_2017_2019_away$PTS, xlab = "Points", col = "red",
-     xlim = c(60, 160), main = "Home 2017-2019", breaks = 8 )
-
-par(opar)
+(ref:cap-Fig2) Shows the home(blue) and away (red) scoring distirbutions for 2020(bottom) and 2017-2019(top)"
+```{r Fig2, echo=FALSE, fig.cap="(ref:cap-Fig2)", fig.align='center'}
+ggplot(combined) + 
+  geom_histogram(aes(x = PTS, y = ..density.., fill = Home.Away), bins = 10, binwidth = 5)+
+  facet_grid(Year ~ Home.Away)
 ```
 
-Figure 3 shows home shooting for two pointers, three pointers and
+Figure \@ref(fig:Fig3) shows home shooting for two pointers, three pointers and
 free throws for 2020(top) vs. 2017-19(bottom). The histograms appear to be fairly
-similarly distributed between 2020 and 2017-19. Likewise, Figure 4, shows the same
-data except for away teams. Just like Figure 3, Figure 4 histograms don’t show any
-obvious differences. The normality of the distributions for both figures 3 and 4
-are important for meeting the required assumptions for statistical analysis.
+similarly distributed between 2020 and 2017-19. Likewise, Figure \@ref(fig:Fig4),
+shows the same data except for away teams. Just like Figure \@ref(fig:Fig3),
+Figure \@ref(fig:Fig4) histograms don’t show anyobvious differences. The normality
+of the distributions in both Figure \@ref(fig:Fig3) and Figure \@ref(fig:Fig4) are
+important for meeting the required assumptions for statistical analysis.
 
-```{r Figthree, echo=FALSE, fig.cap="Home shooting percentages for two pointers, three pointers and free throws for 2020(top) vs. 2017-19 (bottom)"}
-opar<-par(no.readonly = TRUE)
-par(mfrow=c(2,3))
+(ref:cap-Fig3) Home shooting percentages for two pointers, three pointers and free throws for 2020(top) vs. 2017-19 (bottom)
+```{r Fig3, echo=FALSE, fig.cap="(ref:cap-Fig3)"}
+hometwo2020 <- data.frame(Year=game_log_2020_home$Year, Type=rep("Two Pointers", 83),
+                          Percentage=game_log_2020_home$X2P.) 
+homethree2020 <- data.frame(Year=game_log_2020_home$Year, Type=rep("Three Pointers", 83),
+                            Percentage=game_log_2020_home$X3P.) 
+homeFT2020 <- data.frame(Year=game_log_2020_home$Year, Type=rep("Free Throws", 83),
+                         Percentage=game_log_2020_home$FT.) 
+hometwo2017_2019 <- data.frame(Year=game_log_2017_2019_home$Year, Type=rep("Two Pointers", 243),
+                               Percentage=game_log_2017_2019_home$X2P.) 
+homethree2017_2019 <- data.frame(Year=game_log_2017_2019_home$Year, Type=rep("Three Pointers", 243),
+                                 Percentage=game_log_2017_2019_home$X3P.) 
+homeFT2017_2019 <- data.frame(Year=game_log_2017_2019_home$Year, Type=rep("Free Throws", 243),
+                              Percentage=game_log_2017_2019_home$FT.) 
+homegraph<-rbind(hometwo2020, homethree2020, homeFT2020, hometwo2017_2019, homethree2017_2019,
+                 homeFT2017_2019)
 
-hist(game_log_2020_home$X2P., xlab = "Shooting Percentage", col= "green",
-     main = "2020 Home 2 pointers", xlim = c(20, 70))
-
-hist(game_log_2020_home$X3P., xlab = "Shooting Percentage", col= "green",
-     main = "2020 Home 3 pointers", xlim = c(10, 70))
-
-hist(game_log_2020_home$FT., xlab = "Shooting Percentage", col= "green",
-     main = "2020 Home Free Throws", xlim = c( 50, 100))
-
-hist(game_log_2017_2019_home$X2P., xlab = "Shooting Percentage", col= "green",
-     main = "2017-2019 2 pointers", xlim = c(20, 70))
-
-hist(game_log_2017_2019_home$X3P., xlab = "Shooting Percentage", col= "green",
-     main = "2017-2019 3 pointers", xlim = c(10, 70))
-
-hist(game_log_2017_2019_home$FT., xlab = "Shooting Percentage", col= "green",
-     main = "2017-2019 Free Throws", xlim = c( 50, 100))
-
-par(opar)
+ggplot(homegraph) + 
+  geom_histogram(aes(x = Percentage, y = ..density..), fill = "#00BFC4", bins = 10, binwidth = 5)+
+  facet_grid(Year ~ Type)
 ```
 
-```{r Fig4, echo=FALSE, fig.cap="Away shooting percentages for two pointers, three pointers and free throws for 2020(top) vs. 2017-19 (bottom)", fig.align='center'}
-opar<-par(no.readonly = TRUE)
-par(mfrow=c(2,3))
-hist(game_log_2020_away$X2P., xlab = "Shooting Percentage", col= "red", 
-     main = "2020 Away 2 pointers", xlim= c(30, 80))
+(ref:cap-Fig4) Away shooting percentages for two pointers, three pointers and free throws for 2020(top) vs. 2017-19 (bottom)"
+```{r Fig4, echo=FALSE, fig.cap="(ref:cap-Fig4)", fig.align='center'}
+hometwo2020 <- data.frame(Year=game_log_2020_home$Year, Type=rep("Two Pointers", 83),
+                          Percentage=game_log_2020_home$X2P.) 
+homethree2020 <- data.frame(Year=game_log_2020_home$Year, Type=rep("Three Pointers", 83),
+                            Percentage=game_log_2020_home$X3P.) 
+homeFT2020 <- data.frame(Year=game_log_2020_home$Year, Type=rep("Free Throws", 83),
+                         Percentage=game_log_2020_home$FT.) 
+hometwo2017_2019 <- data.frame(Year=game_log_2017_2019_home$Year, Type=rep("Two Pointers", 243),
+                               Percentage=game_log_2017_2019_home$X2P.) 
+homethree2017_2019 <- data.frame(Year=game_log_2017_2019_home$Year, Type=rep("Three Pointers", 243),
+                                 Percentage=game_log_2017_2019_home$X3P.) 
+homeFT2017_2019 <- data.frame(Year=game_log_2017_2019_home$Year, Type=rep("Free Throws", 243),
+                              Percentage=game_log_2017_2019_home$FT.) 
+homegraph<-rbind(hometwo2020, homethree2020, homeFT2020, hometwo2017_2019, homethree2017_2019,
+                 homeFT2017_2019)
 
-hist(game_log_2020_away$X3P., xlab = "Shooting Percentage", col= "red", 
-     main = "2020 Away 3 pointers", xlim = c(10, 60))
-
-hist(game_log_2020_away$FT., xlab = "Shooting Percentage", col= "red", 
-     main = "2020 Away Free Throws", xlim = c(40, 100))
-
-hist(game_log_2017_2019_away$X2P., xlab = "Shooting Percentage", col= "red",
-     main = "2017-19 Away 2 pointers", xlim= c(30, 80))
-
-hist(game_log_2017_2019_away$X3P., xlab = "Shooting Percentage", col= "red",
-     main = "2017-19 Away 3 pointers", xlim = c(10, 60))
-
-hist(game_log_2017_2019_away$FT., xlab = "Shooting Percentage", col= "red",
-     main = "2017-19 Away Free Throws", xlim = c(40, 100))
-par(opar)
+ggplot(homegraph) + 
+  geom_histogram(aes(x = Percentage, y = ..density..), fill = "#00BFC4", bins = 10, binwidth = 5)+
+  facet_grid(Year ~ Type)
 ```
@@
 titles <- c(2020, "2017-19", "95% CI LL", "95% CI UL", "Test Stat.", "P-value")
-rows <- c("Home Team Win %", "Home Team Scoring", "Away Team Scoring", "Home Team 2P%", "Home Team 3P%",
-          "Home Team FT%", "Away Team 2P%", "Away Team  3P%", "Away Team FT%")
+rows <- c("Home Win %", "Home Scoring", "Away Scoring", "Home 2P%", "Home 3P%",
+          "Home FT%", "Away 2P%", "Away 3P%", "Away FT%")
 results <- data.frame(rowone, rowtwo, rowthree, rowfour, rowfive, rowsix,
@@
 resultsfinal <- data.frame(t(results))
-resultsfinal <- round(resultsfinal, 4)
+resultsfinal1 <- round(resultsfinal[1:5], 3)
+resultsfinal2 <- round(resultsfinal[6], 4)
+resultsfinal <-cbind(resultsfinal1, resultsfinal2)
 colnames(resultsfinal) <- titles
 row.names(resultsfinal) <- rows
-
 ```
@@
 
-Starting from the top of the Table 1, we see a statistically significant change
+Starting from the top of the Table \@ref(tab:table), we see a statistically significant change
 in home win percentage in 2020 from 2017-19, with p-value of 0.0365. The 95% 
@@
 understand that home teams seemed to struggle to win compared to normal conditions.
-Compared to [@Kotecki]who finds home teams consistently have a significantly
+Compared to @Kotecki who finds home teams consistently have a significantly
 better record than away teams boasting about a 60.5% win percentage in his sample,
@@
 away teams did make at least small increases. Although these increases weren’t
-all significant these increases are exactly what is reported in [@Carron2005]
+all significant these increases are exactly what is reported in @Carron2005
 when they explain how evidence suggests that teams perform better with the absence
@@
 increase in winning. The improvement of away teams confirms a proposition from 
-[@Greer)] who believed that the positive social impact of crowds benefiting 
+@Greer who believed that the positive social impact of crowds benefiting 
 home teams may be a result of inhibiting away teams.
@@
 At least some of that improvement from away teams came from significantly higher 
-two point efficiency. This corresponds with the conclusion from [@Harris],
+two point efficiency. This corresponds with the conclusion from @Harris,
 where they found home teams are best suited to capitalize on advantages from two point

### 2020-11-02 Manuscript/NBABubbleAnalysis.Rmd
Update NBABubbleAnalysis.Rmd

@@
 in home win percentage in 2020 from 2017-19, with p-value of 0.0365. The 95% 
-confidence interval of (-0.255, -0.0075) confirms our belief that home-court 
+confidence interval of (-0.255, -0.008) confirms our belief that home-court 
 advantage was lost in the 2020 NBA playoffs. However, after accounting for 
@@
 Likewise, the away team two point shooting efficiency increased by an average of
-about 0.03 (95% CI: 0.0147, 0.0503). However, unlike two point shooting, away 
+about 0.03 (95% CI: 0.015, 0.050). However, unlike two point shooting, away 
 teams did not see a statistically significant difference in three point and free

### 2020-11-02 Manuscript/NBABubbleAnalysis.Rmd
Update NBABubbleAnalysis.Rmd

@@
 where they found home teams are best suited to capitalize on advantages from two point
-shots. Normally, by shooting more 2 themselves and forcing away teams to shoot more 2
-they home team benefits from increasing home-court advantage. However, with away teams
-significantly improving two point shooting in the bubble this strategy was no longer 
-viable and home-court advantage disappeared.
+shots. Normally, by shooting more two pointers themselves and forcing away teams to 
+shoot more two pointers the home team benefits from increasing effects of home-court advantage. 
+However, with away teams significantly improving two point shooting in the bubble this strategy 
+was no longer viable and home-court advantage disappeared.
 

### 2020-11-02 Manuscript/NBABubbleAnalysis.Rmd
Update NBABubbleAnalysis.Rmd

@@
 ```{r Fig4, echo=FALSE, fig.cap="(ref:cap-Fig4)", fig.align='center'}
-hometwo2020 <- data.frame(Year=game_log_2020_home$Year, Type=rep("Two Pointers", 83),
-                          Percentage=game_log_2020_home$X2P.) 
-homethree2020 <- data.frame(Year=game_log_2020_home$Year, Type=rep("Three Pointers", 83),
-                            Percentage=game_log_2020_home$X3P.) 
-homeFT2020 <- data.frame(Year=game_log_2020_home$Year, Type=rep("Free Throws", 83),
-                         Percentage=game_log_2020_home$FT.) 
-hometwo2017_2019 <- data.frame(Year=game_log_2017_2019_home$Year, Type=rep("Two Pointers", 243),
-                               Percentage=game_log_2017_2019_home$X2P.) 
-homethree2017_2019 <- data.frame(Year=game_log_2017_2019_home$Year, Type=rep("Three Pointers", 243),
-                                 Percentage=game_log_2017_2019_home$X3P.) 
-homeFT2017_2019 <- data.frame(Year=game_log_2017_2019_home$Year, Type=rep("Free Throws", 243),
-                              Percentage=game_log_2017_2019_home$FT.) 
-homegraph<-rbind(hometwo2020, homethree2020, homeFT2020, hometwo2017_2019, homethree2017_2019,
-                 homeFT2017_2019)
+awaytwo2020 <- data.frame(Year=game_log_2020_away$Year, Type=rep("Two Pointers", 83),
+                          Percentage=game_log_2020_away$X2P.) 
+awaythree2020 <- data.frame(Year=game_log_2020_away$Year, Type=rep("Three Pointers", 83),
+                            Percentage=game_log_2020_away$X3P.) 
+awayFT2020 <- data.frame(Year=game_log_2020_away$Year, Type=rep("Free Throws", 83),
+                         Percentage=game_log_2020_away$FT.) 
+awaytwo2017_2019 <- data.frame(Year=game_log_2017_2019_away$Year, Type=rep("Two Pointers", 243),
+                               Percentage=game_log_2017_2019_away$X2P.) 
+awaythree2017_2019 <- data.frame(Year=game_log_2017_2019_away$Year, Type=rep("Three Pointers", 243),
+                                 Percentage=game_log_2017_2019_away$X3P.) 
+awayFT2017_2019 <- data.frame(Year=game_log_2017_2019_away$Year, Type=rep("Free Throws", 243),
+                              Percentage=game_log_2017_2019_away$FT.) 
+awaygraph<-rbind(awaytwo2020, awaythree2020, awayFT2020, awaytwo2017_2019, awaythree2017_2019,
+                 awayFT2017_2019)
 
-ggplot(homegraph) + 
-  geom_histogram(aes(x = Percentage, y = ..density..), fill = "#00BFC4", bins = 10, binwidth = 5)+
+ggplot(awaygraph) + 
+  geom_histogram(aes(x = Percentage, y = ..density..), fill = "#F8766D", bins = 10, binwidth = 5)+
   facet_grid(Year ~ Type)
@@
 shots. Normally, by shooting more two pointers themselves and forcing away teams to 
-shoot more two pointers the home team benefits from increasing effects of home-court advantage. 
-However, with away teams significantly improving two point shooting in the bubble this strategy 
-was no longer viable and home-court advantage disappeared.
+shoot more two pointers the home team benefits from increasing effects of
+home-court advantage. However, with away teams significantly improving two point
+shooting in the bubble this strategy was no longer viable and home-court
+advantage disappeared.
 

### 2020-11-03 Manuscript/NBABubbleAnalysis.Rmd
Update NBABubbleAnalysis.Rmd

@@
 There is tons of published literature on the effects of home-court advantage.
-Many of which mainly focus on the effects of home-court advantage using shooting
-in the NBA. @Kotecki studies the impact of home-court advantage on winning in the
+Many NBA home-court advantage studies analyze the effects by studying shooting
+percentages. @Kotecki examines the impact of home-court advantage on winning in the
 NBA. Kotecki proves the existence of home-court advantage using performance-based

### 2020-11-03 Manuscript/NBABubbleAnalysis.Rmd
Update NBABubbleAnalysis.Rmd

@@
 
-Unfortunately, this study is limited by a one time sample. It is unknown if there
-will ever be another NBA playoffs played in this fashion. It seems unlikely that 
-these conditions will ever be repeated. That being said, it does not seem like a 
-follow-up study using the same measures with a different sample will ever be
-possible. If possible, that type of study could help strengthen the conclusions
+Unfortunately, this study is limited by a one time sample. It seems unlikely that 
+these conditions will ever be repeated. That being said, it will not be possible
+to have a follow-up study using the same measures with a different sample. If
+possible, that type of study could help strengthen the conclusions
 made in this paper.

### 2020-11-03 Manuscript/NBABubbleAnalysis.Rmd
Update NBABubbleAnalysis.Rmd

@@
                 sd(game_log_2020_home$PTS), sigma.y = sd(game_log_2017_2019_home$PTS))
-rowtwo <- c(as.numeric(two$estimate[1]), as.numeric(two$estimate[2]), 
+rowtwo <- c(as.numeric(two$estimate[1])/100, as.numeric(two$estimate[2])/100, 
             two$conf.int, two$statistic, two$p.value)
@@
                   sd(game_log_2020_away$PTS), sigma.y = sd(game_log_2017_2019_away$PTS))
-rowthree <- c(as.numeric(three$estimate[1]), as.numeric(three$estimate[2]), 
+rowthree <- c(as.numeric(three$estimate[1])/100, as.numeric(three$estimate[2])/100, 
               three$conf.int, three$statistic, three$p.value)

### 2020-11-03 Manuscript/NBABubbleAnalysis.Rmd
Update NBABubbleAnalysis.Rmd

@@
 focusing on the 2017 to 2020 playoffs we can take a quick look at home team win
-percentages since 2010. Notice in Figure \@ref(fig:Fig1), the10 years before 2020, the home team
+percentages since 2010. Notice in Figure \@ref(fig:Fig1), the 10 years before 2020, the home team
 winning percentage ranged from around 0.56 to 0.7 and never dipped below .5. The 
-2020 bubble, however, broke this historicpattern dipping down below .5. Foreshadowing, 
-the confirmation of the expectationthat the effect of home-court advantage was removed
+2020 bubble broke this historic pattern dipping down below .5. Foreshadowing, 
+the confirmation of the expectation that the effect of home-court advantage was removed
 in the 2020 playoffs.
@@
 ```{r Fig1, echo=FALSE, fig.cap = " (ref:cap-Fig1)", fig.align='center'}
-plot(years, winpct, ylim = c(0,1), title("NBA Home Playoff Win % Since 2010"), col = ifelse(winpct==.482, "Red", "Black"), pch=16)
+plot(years, winpct, ylim = c(0,1), title("NBA Home Playoff Win % Since 2010"),
+     col = ifelse(winpct==.482, "Red", "Black"), pch=16)
 text(years, winpct, labels = winpct, cex = 0.5, pos = 1)
@@
 shows the same data except for away teams. Just like Figure \@ref(fig:Fig3),
-Figure \@ref(fig:Fig4) histograms don’t show anyobvious differences. The normality
+Figure \@ref(fig:Fig4) histograms don’t show any obvious differences. The normality
 of the distributions in both Figure \@ref(fig:Fig3) and Figure \@ref(fig:Fig4) are

### 2020-11-03 Manuscript/NBABubbleAnalysis.Rmd
Update NBABubbleAnalysis.Rmd

@@
 just shooting efficiency to pick through looking for more possible sources of
-added points for away teams.This will further help explain what is lost in the
+added points for away teams. This will further help explain what is lost in the
 performance of away teams when they travel to opposing arenas. This study is only

### 2020-11-03 Manuscript/NBABubbleAnalysis.Rmd
Update NBABubbleAnalysis.Rmd

@@
 #Data Collection
-game_log_2020 <- read.csv(file="Game_Log_2020.csv", header=TRUE)
-game_log_2017_2019 <- read.csv(file="2017-2019_game_data.csv", header=TRUE)
+game_log_2020 <- read.csv(file="../data/Game_Log_2020.csv", header=TRUE)
+game_log_2017_2019 <- read.csv(file="../data/2017-2019_game_data.csv", header=TRUE)
 

### 2020-11-03 Manuscript/NBABubbleAnalysis.Rmd
minor fix

@@
    - \usepackage{float}
+   - \usepackage[margin=1in]{geometry}
 toc: False
@@
 \newpage
-# I. Introduction  
+# Introduction  
 Home-court advantage is often discussed in sports circles as a contributing 
@@
 
-#References
+# References

### 2020-11-04 Manuscript/NBABubbleAnalysis.Rmd
abstract

--- 2020-11-03 Manuscript/NBABubbleAnalysis.Rmd
@@
 ---
-title: "Analyzing the Effects of the NBA COVID Bubble on the NBA Playoffs"
+title: "The Effects of the NBA COVID Bubble on the NBA Playoffs"
 subtitle: "A Case Study for Home-Court Advantage"
 author: |
-  | Michael Price
+  | Michael Price and Jun Yan
   | michael.price@uconn.edu
   | Department of Statistics, University of Connecticut
-  | Dr. Jun Yan
-  | jun.yan@uconn.edu
-  | Department of Statistics, University of Connecticut
-
 date: "`r format(Sys.time(), '%d, %B, %Y')`"
@@
 the COVID-19 pandemic. This meant that there would be no fans in attendance, 
-games played on neutral courts and no traveling for teams which in theory 
-removes home-court advantage from the games. This idea was a hot topic of 
+games played on neutral courts and no traveling for teams, which in theory 
+removes home-court advantage from the games. This setting has attracted a lot of
 discussion as analysts and fans debated the possible effects it may have on the
 outcome of games. Home-court advantage has historically played a huge role in 
-NBA playoff series outcomes. Thus, to study the effects of the bubble, and 
-home-court advantage, the 2020 season is compared to the 2017, 2018, and 2019
-playoffs. While many factors contribute to the outcome of games, points scored
-is the deciding factor of who wins games so scoring will be the primary focus 
+NBA playoff series outcomes. The 2020 playoff provided a unique
+opportunituy to study the effects of the bubble, and home-court
+advantage by comparing the 2020 season with the seasons in the past
+While many factors contribute to the outcome of games, points scored
+is the deciding factor of who wins games so scoring is the primary focus 
 of this study. The specific measures of interest are team scoring totals and 
@@
 Comparing these measures for home teams and away teams in 2020 vs. 2017-2019
-shows how the 2020 playoffs favored away teams more than normal, particularly
+shows that the 2020 playoffs favored away teams more than normal, particularly
 with two point shooting and total scoring.
@@
 ```
-\newpage
-# I. Introduction  
+
+# Introduction  
 Home-court advantage is often discussed in sports circles as a contributing 

### 2020-11-04 Manuscript/NBABubbleAnalysis.Rmd
Update NBABubbleAnalysis.Rmd

@@
 header-includes:
-   - \usepackage{float}
    - \usepackage[margin=1in]{geometry}
@@
 library(ggplot2)
-knitr::opts_chunk$set(fig.width = 4, fig.height = 3, dpi = 300,
-out.width = "90%", fig.align = "center", dev = "pdf", fig.pos = "H")
+knitr::opts_chunk$set(fig.width = 5, fig.height = 4, dpi = 300,
+out.width = "90%", fig.align = "center", dev = "pdf")
 ```
 
-# Introduction  
+# Introduction {#sec:intro}
 Home-court advantage is often discussed in sports circles as a contributing 
@@
 home-court advantage. Some of these studies are reviewed in greater detail in
-Section II Literature Review.
+Section \@ref(sec:litrev) Literature Review.
 
@@
 
-# II. Literature Review  
+# Literature Review {#sec:litrev}
 There is tons of published literature on the effects of home-court advantage.
@@
 
-# III. Data and Methods  
+# Data {#sec:data}
 The 2020 bubble provides a new and exciting opportunity to study home-court
@@
 
-# IV. Analysis and Results
+# Methods {#sec:methods}
+
+# Analysis and Results {#sec:results}
 ```{r, echo=FALSE}
@@
 
-# V. Discussion 
+# Discussion {#sec:disc}
 Generally it seemed that away teams fared better in the 2020 NBA playoff bubble
@@
 
-# References
+# References {#sec:ref}

### 2020-11-04 Manuscript/NBABubbleAnalysis.Rmd
Update NBABubbleAnalysis.Rmd

@@
 ```{r Fig1, echo=FALSE, fig.cap = " (ref:cap-Fig1)", fig.align='center'}
-plot(years, winpct, ylim = c(0,1), title("NBA Home Playoff Win % Since 2010"),
-     col = ifelse(winpct==.482, "Red", "Black"), pch=16)
-text(years, winpct, labels = winpct, cex = 0.5, pos = 1)
-text(2010, 0.51, labels = "0.5", cex = 0.6, col = "Green")
-abline(h=0.5, col = "Green")
+winyear <- cbind(winpct, years)
+winyear <- as.data.frame(winyear)
+highlight <- data.frame(winpct = .482, years = 2020)
+ggplot(winyear, aes(x=years, y=winpct))+ 
+      geom_point()+
+      geom_point(data=highlight, 
+             aes(x=years, y=winpct), 
+             color='red')+
+      scale_x_continuous(name = "Year", breaks = years, labels = years)+
+      scale_y_continuous(limits = c(0.4, 0.8))+
+      geom_hline(aes(yintercept = 0.5), color = "green")+
+      geom_text(aes(label=winpct), hjust=0, vjust=-1)
 ```

### 2020-11-08 Manuscript/NBABubbleAnalysis.Rmd
Update NBABubbleAnalysis.Rmd

@@
 # Data {#sec:data}
-The 2020 bubble provides a new and exciting opportunity to study home-court
-advantage for the NBA. Unlike college basketball, aside from a few 
-exhibition/preseason games, the NBA always has a home and away team. So, for
-the first time in NBA history the bubble allows NBA home and away performance
-to be compared vs a control/neutral field. The NBA bubble, as a neutral court,
-removed all 4 possible game location factors impacting home-court advantage 
-hypothesized by @Carron1992. The NBA bubble featured 8 seeding games then a
-standard playoff format. The focus of this study was the play during the playoff
-games since it followed the standard playoff format and can easily be compared
-back to other playoffs. For this study, the 2020 playoffs were compared against
-the three previous playoffs collectively. It is important to make sure that any
-observed differences are truly from a home-court advantage. So, in order to control
-for the changing play style of the NBA, we will limit the study to 2020 vs 2017-2019.
-The reason being the faster pace play and more common use of the three-point shot 
-in modern basketball. If we used data from say 10 years ago, or earlier, observed
-differences may not be from effects of the NBA bubble, but rather from the effects
-of drastic changes in the style of play between the seasons. However, basketball
-evolves slow enough that we can reasonably assume 2017-2019 are at least very 
-close in pace and playing style to 2020.
-
 Data was collected from the official NBA website. The main variables of interest
@@
 # Methods {#sec:methods}
+The 2020 bubble provides a new and exciting opportunity to study home-court
+advantage for the NBA. Unlike college basketball, aside from a few 
+exhibition/preseason games, the NBA always has a home and away team. So, for
+the first time in NBA history the bubble allows NBA home and away performance
+to be compared vs a control/neutral field. The NBA bubble, as a neutral court,
+removed all 4 possible game location factors impacting home-court advantage 
+hypothesized by @Carron1992. The NBA bubble featured 8 seeding games then a
+standard playoff format. The focus of this study was the play during the playoff
+games since it followed the standard playoff format and can easily be compared
+back to other playoffs. For this study, the 2020 playoffs were compared against
+the three previous playoffs collectively. It is important to make sure that any
+observed differences are truly from a home-court advantage. So, in order to control
+for the changing play style of the NBA, we will limit the study to 2020 vs 2017-2019.
+The reason being the faster pace play and more common use of the three-point shot 
+in modern basketball. If we used data from say 10 years ago, or earlier, observed
+differences may not be from effects of the NBA bubble, but rather from the effects
+of drastic changes in the style of play between the seasons. However, basketball
+evolves slow enough that we can reasonably assume 2017-2019 are at least very 
+close in pace and playing style to 2020.
+
+Comparisons between 2020 and 2017-19 home and away teams were made on home team
+winning percentage, total team scoring and two-point, three-point and free throw
+shooting. Comparing the differences in these metrics for home and away teams in
+2020 vs previous years will provide valuable insights to the understanding of 
+home-court advantage. We can see how going on the road may negatively impact 
+away performance and how playing at home may positively impact home performance.
+If there are differences in scoring for home or away teams the differences can
+be used to show how home-court advantage affects overall performance of home
+and away teams. While testing for differences in shooting will provide added
+context for how home-court advantage specifically affects performance. 
+Shooting percentages aren’t the only possible metrics affected by home-court
+advantage, but they are the most obvious and certainly an important one. In
+order to compare these differences simple two-sample z-tests were conducted.
+Test for differences in means was conducted for comparing scoring totals. 
+Tests for differences in proportions were conducted for comparing winning
+percentage and the three different shooting percentages.
 

### 2020-11-08 Manuscript/NBABubbleAnalysis.Rmd
Update NBABubbleAnalysis.Rmd

@@
 order to compare these differences simple two-sample z-tests were conducted.
-Test for differences in means was conducted for comparing scoring totals. 
-Tests for differences in proportions were conducted for comparing winning
-percentage and the three different shooting percentages.
+Test for differences in means was conducted for comparing scoring totals, shown in 
+Equation \@ref(eq:zmean). Tests for differences in proportions were conducted for
+comparing winning percentage and the three different shooting percentages, shown
+in Equation \@ref(eq:zprop).
+
+\begin{align}
+    z = \frac{(\bar{X}_1 - \bar{X}_2)-0}{\sqrt{\frac{s_1^2}{n_1}+\frac{s_2^2}{n_2}}}.
+    (\#eq:zmean)
+\end{align}
+
+\begin{align}
+    z = \frac{(\hat{p}_1 - \hat{p}_2)-0}{\sqrt{\hat{p}(1-\hat{p})(\frac{1}{n_1}+\frac{1}{n_2})}}.
+    (\#eq:zprop)
+\end{align}
 

### 2020-11-08 Manuscript/NBABubbleAnalysis.Rmd
Update NBABubbleAnalysis.Rmd

@@
   geom_histogram(aes(x = PTS, y = ..density.., fill = Home.Away), bins = 10, binwidth = 5)+
-  facet_grid(Year ~ Home.Away)
+  facet_grid(Year ~ Home.Away)+
+  theme(legend.position = "none")
 ```

### 2020-11-08 Manuscript/NBABubbleAnalysis.Rmd
Update NBABubbleAnalysis.Rmd

@@
 years <- 2010:2020
-winpct <- c(.6707, .667, .679, .635, .562, .593, .6744, .5696, .707, .561, .482)
+winpct <- c(.671, .667, .679, .635, .562, .593, .674, .570, .707, .561, .482)
 ```
@@
       geom_hline(aes(yintercept = 0.5), color = "green")+
-      geom_text(aes(label=winpct), hjust=0, vjust=-1)
+      geom_text(aes(label=winpct), hjust=0, vjust=-1, size=2)
 ```

### 2020-11-09 Manuscript/NBABubbleAnalysis.Rmd
Update NBABubbleAnalysis.Rmd

@@
 library(ggplot2)
-knitr::opts_chunk$set(fig.width = 5, fig.height = 4, dpi = 300,
+knitr::opts_chunk$set(fig.width = 6, fig.height = 4, dpi = 300,
 out.width = "90%", fig.align = "center", dev = "pdf")
@@
 ```{r table, echo = FALSE}
-knitr::kable(resultsfinal, caption = 'The results from the 9 statistical tests', align = "cccccc")
+knitr::kable(resultsfinal, caption = 'The results from the 9 statistical tests', align = "cccccc", booktabs = TRUE)
 ```

### 2020-11-10 Manuscript/NBABubbleAnalysis.Rmd
Update NBABubbleAnalysis.Rmd

@@
 
-(ref:cap-Fig2) Shows the home(blue) and away (red) scoring distirbutions for 2020(bottom) and 2017-2019(top)"
+(ref:cap-Fig2) Shows the home(blue) and away (red) scoring distributions for 2020(bottom) and 2017-2019(top)"
 ```{r Fig2, echo=FALSE, fig.cap="(ref:cap-Fig2)", fig.align='center'}
@@
 comparing winning percentage and the three different shooting percentages, shown
-in Equation \@ref(eq:zprop).
+in Equation \@ref(eq:zprop). All 9 tests are conducted under the assumption and null
+hypothesis that there should be no difference in statistics from 2020 compared
+to those from 2017-2019.
 

### 2020-11-10 Manuscript/NBABubbleAnalysis.Rmd
Update NBABubbleAnalysis.Rmd

@@
 For example, @Greer focuses his study on the influence of spectator booing on 
-Home-court advantage in basketball. The three methods of performance used
+home-court advantage in basketball. The three methods of performance used
 in this study were scoring, violations and turnovers. This study was conducted 
@@
 ```{r table, echo = FALSE}
-knitr::kable(resultsfinal, caption = 'The results from the 9 statistical tests', align = "cccccc", booktabs = TRUE)
+knitr::kable(resultsfinal, caption = 'The results from the 9 tests', align = "cccccc", booktabs = TRUE)
 ```

### 2020-11-10 Manuscript/NBABubbleAnalysis.Rmd
Update NBABubbleAnalysis.Rmd

@@
 
-(ref:cap-Fig1) "Winning percentage of NBA home teams in the playoffs since 2010, the green line denotes .500"
+(ref:cap-Fig1) Winning percentage of NBA home teams in the playoffs since 2010, the green line denotes .500
 ```{r Fig1, echo=FALSE, fig.cap = " (ref:cap-Fig1)", fig.align='center'}
@@
 Figure \@ref(fig:Fig3) shows home shooting for two pointers, three pointers and
-free throws for 2020(top) vs. 2017-19(bottom). The histograms appear to be fairly
+free throws for 2020 (top) vs. 2017-19 (bottom). The histograms appear to be fairly
 similarly distributed between 2020 and 2017-19. Likewise, Figure \@ref(fig:Fig4),

### 2020-11-10 Manuscript/NBABubbleAnalysis.Rmd
Update NBABubbleAnalysis.Rmd

@@
 
-(ref:cap-Fig2) Shows the home(blue) and away (red) scoring distributions for 2020(bottom) and 2017-2019(top)"
+(ref:cap-Fig2) Shows the home(blue) and away (red) scoring distributions for 2020(bottom) and 2017-2019(top)
 ```{r Fig2, echo=FALSE, fig.cap="(ref:cap-Fig2)", fig.align='center'}

### 2020-11-10 Manuscript/NBABubbleAnalysis.Rmd
hyperlinks for citations with asa bibstyle

@@
    - \usepackage[margin=1in]{geometry}
+   - \usepackage {hyperref}
+   - \hypersetup {colorlinks = true, linkcolor = blue, citecolor=blue, urlcolor = red}
 toc: False

### 2020-11-10 Manuscript/NBABubbleAnalysis.Rmd
Update NBABubbleAnalysis.Rmd

@@
 order to compare these differences simple two-sample z-tests were conducted.
-Test for differences in means was conducted for comparing scoring totals, shown in 
-Equation \@ref(eq:zmean). Tests for differences in proportions were conducted for
-comparing winning percentage and the three different shooting percentages, shown
-in Equation \@ref(eq:zprop). All 9 tests are conducted under the assumption and null
-hypothesis that there should be no difference in statistics from 2020 compared
-to those from 2017-2019.
 
+A test for differences in means was conducted to compare scoring totals. Let
+$\bar{X}_1$ represent the average scoring total for 2020 and $\bar{X}_2$ represent
+average scoring total over 2017-19. Similarly, $s_1^2$ represent the sample variance
+of scoring for 2020 and $s_2^2$ represent sample variance of scoring over 2017-19.
+Finally, let $n_1$ represent the number of games in 2020 and $n_2$ represent the number
+of games from 2017-2019. Then the test statistic is constructed using Equation \@ref(eq:zmean).
 \begin{align}
@@
 \end{align}
+We will test the null hypothesis that $\mu_1$=$\mu_2$. The test will be conducted twice
+once comparing home scoring in 2020 to 2017-19 and once comparing away
+scoring in 2020 to 2017-2019.
 
+Tests for differences in proportions were conducted for comparing winning percentage
+and the three different shooting percentages in 2020 to those of 2017-2019. Let
+$\hat{p}_1$ represents sample proportion for 2020 and $\hat{p}_2$ represents sample
+proportion for 2017-2019. While, $\hat{p}$ is $\hat{p}_1$ and $\hat{p}_2$ pooled 
+together.Let $n_1$ represent number of observations for 2020 and $n_2$ represent 
+number of observations for 2017-2019. Then the test statistic is constructed using
+Equation \@ref(eq:zprop).
 \begin{align}
@@
 \end{align}
+We will test the null hypothesis that $p_1$=$p_2$. The test will be conducted seven times,
+once comparing home winning percentage for 2020 to that of 2017-2019 and once for each of
+two-point shooting, three-point shooting and free throws for both home and away teams
+in 2020 vs. 2017-2019.
 

### 2020-11-10 Manuscript/NBABubbleAnalysis.Rmd
Update NBABubbleAnalysis.Rmd

@@
 The 2020 NBA playoffs were played inside of a bubble in Disney World because of
-the COVID-19 pandemic. This meant that there would be no fans in attendance, 
+the COVID-19 pandemic. This meant that there were no fans in attendance, 
 games played on neutral courts and no traveling for teams, which in theory 
@@
 
-  Given the approximate normality of data, 9 different z-tests were performed in 
-  order to test the effects of the COVID bubble on the 2020 NBA playoffs. The 9 questions were:  
-* Is the home team winning percentage in 2020 different than that it was in 2017-2019?  
-* Is the average home team scoring different in 2020 than it was over 2017-2019?  
-* Is the average away team scoring different in 2020 than it was over 2017-2019?  
-* Are home teams making two pointers at the same rate in 2020 as 2017-2019?  
-* Are home teams making three pointers at the same rate in 2020 as 2017-2019?  
-* Are home teams making free throws at the same rate in 2020 as 2017-2019?  
-* Are away teams making two pointers at the same rate in 2020 as 2017-2019?  
-* Are away teams making three pointers at the same rate in 2020 as 2017-2019?  
-* Are away teams making free throws at the same rate in 2020 as 2017-2019?
-
 # Methods {#sec:methods}
@@
 order to compare these differences simple two-sample z-tests were conducted.
+
+Given the approximate normality of data, 9 different z-tests were performed in 
+  order to test the effects of the COVID bubble on the 2020 NBA playoffs. The 9 questions were:  
+* Is the home team winning percentage in 2020 different than that it was in 2017-2019?  
+* Is the average home team scoring different in 2020 than it was over 2017-2019?  
+* Is the average away team scoring different in 2020 than it was over 2017-2019?  
+* Are home teams making two pointers at the same rate in 2020 as 2017-2019?  
+* Are home teams making three pointers at the same rate in 2020 as 2017-2019?  
+* Are home teams making free throws at the same rate in 2020 as 2017-2019?  
+* Are away teams making two pointers at the same rate in 2020 as 2017-2019?  
+* Are away teams making three pointers at the same rate in 2020 as 2017-2019?  
+* Are away teams making free throws at the same rate in 2020 as 2017-2019?
 

### 2020-11-11 Manuscript/NBABubbleAnalysis.Rmd
edits and comments

@@
    - \usepackage {hyperref}
-   - \hypersetup {colorlinks = true, linkcolor = blue, citecolor=blue, urlcolor = red}
+   - \hypersetup {colorlinks = true, linkcolor = blue, citecolor=blue, urlcolor = blue}
 toc: False
@@
 games played on neutral courts and no traveling for teams, which in theory
-removes home-court advantage from the games. This setting has attracted a lot of
+removes home-court advantage from the games. This setting has attracted much
 discussion as analysts and fans debated the possible effects it may have on the
-outcome of games. Home-court advantage has historically played a huge role in 
-NBA playoff series outcomes. The 2020 playoff provided a unique
-opportunituy to study the effects of the bubble, and home-court
-advantage by comparing the 2020 season with the seasons in the past
+outcome of games. Home-court advantage has historically played an influential
+role in NBA playoff series outcomes. The 2020 playoff provided a unique
+opportunity to study the effects of the bubble and home-court
+advantage by comparing the 2020 season with the seasons in the past.
 While many factors contribute to the outcome of games, points scored
-is the deciding factor of who wins games so scoring is the primary focus 
+is the deciding factor of who wins games, so scoring is the primary focus
 of this study. The specific measures of interest are team scoring totals and
@@
 Comparing these measures for home teams and away teams in 2020 vs. 2017-2019
-shows that the 2020 playoffs favored away teams more than normal, particularly
+shows that the 2020 playoffs favored away teams more than usual, particularly
 with two point shooting and total scoring.
@@
 in a better chance of winning. Thus, the NBA playing the 2020 playoffs in a
-bubble brought a great deal of concern for fans, teams, journalists, and others. 
-Steve Aschburner discusses details about anticipated effects in his article 
-“How Orlando's Neutral Site Will Impact Teams, Referees and Games” @Aschburner. 
-Aschburner shares concerns from former players, coaches and other experts about
+bubble due to the COVID-19 pandemic
+brought a great deal of concern for fans, teams, journalists, and others.
+For example, @Aschburner discusses about anticipated effects, sharing
+concerns from former players, coaches and other experts about
 the potential effects of removing home-court advantage. Aschburner notes that the
@@
 the occasion and perform better not having to deal with the headache of going
-on the road? This paper seeks to answer that using scoring totals and shooting
+on the road? We seek to answer the questions using scoring totals and shooting
 percentages as indicators of team performance. This will deepen understanding
@@
 
-This study is quite different from earlier NBA home-court advantage studies. 
+Our study is quite different from earlier NBA home-court advantage studies.
 By using the neutral site games of 2020 we will get to compare home and away
 performance to a control. Typically, studies just compare home vs away
-performance. These studies don’t separate the effects of home-court advantage
+performance. These studies do not separate the effects of home-court advantage
 into the specific effect on the home team and the specific effect on the away
@@
 home-court advantage. Some of these studies are reviewed in greater detail in
-Section \@ref(sec:litrev) Literature Review.
+Section~\@ref(sec:litrev).
 
-This study will compare home team performance in 2020 at a neutral site with 
+We will compare home team performance in 2020 at a neutral site with
 no fans vs. 2017-19 playoffs with fans. Likewise, away team performance in 2020
@@
 teams in 2020 to home teams in 2017-19 and away teams in 2020 to away teams in
-2017-19, I can add a new perspective to the field of research. This will allow for
+2017-19, we add a new perspective to the field of research. This will allow for
 a more accurate understanding of the effects of home-court advantage on home and
-away teams in the NBA. We won’t just see that home-court advantage helps home 
-teams outperform away teams, but this study will actually clearly separate the 
+away teams in the NBA. We will not only see that home-court advantage helps home
+teams outperform away teams, but also separate the
 effects of home-court advantage on home teams and away teams performance individually.
 
-9 statistical tests are run to understand the differences in 2020 vs. earlier 
+Nine hypotheses were tested to understand the differences in 2020 vs. earlier
 years. First, whether or not the difference between home win percentage in 2020
-and 2017-19 is statistically significant. To confirm the difference we see in the
-winning percentage is truly significant. Then we can look for differences in home
+and 2017-19 is zero. This difference is found to be statistically
+significant from zero. Then we assess for differences in home
 scoring in 2020 vs 2017-2019. Similarly, we can do the same test, but for
 differences in away scoring in 2020 vs 2017-2019. Also, differences in team
-shooting(for two pointers, three pointers, and free throws) from 2020 vs 2017-2019
-for both home and away teams. The results from these simple tests will bring a new
-understanding of how home-court advantage is impacting games by altering the performance
+shooting (for two pointers, three pointers, and free throws) from 2020 vs 2017-2019
+for both home and away teams. The results from these tests bring a new perspective
+to understand of how home-court advantage impacts games by altering the performance
 of the home and away teams.
@@
 # Literature Review {#sec:litrev}
-There is tons of published literature on the effects of home-court advantage.
+There is a voluminous literature on the effects of home-court advantage.
 Many NBA home-court advantage studies analyze the effects by studying shooting
-percentages. @Kotecki examines the impact of home-court advantage on winning in the
-NBA. Kotecki proves the existence of home-court advantage using performance-based
+percentages. @Kotecki reported significant home-court advantage using performance-based
 statistics, specifically field goal percentage, free throw percentage, and
-points. All of which he proves significantly improved from home-court advantage, 
-indicating home-court helps teams play better. This conclusion comes from 
-comparing home performance vs. away performance in games. @Cao studied the effects
-of pressure on performance in the NBA. They use free throws as their measure of 
-interest, and research home and away factors in the study. They expected that home
-fans could distract and put pressure on opposing players to make free throws,
-which would negatively impact away team shooting. Likewise, the home team
-would see a positive effect on free throw shooting. However, they find insignificant
+points by comparing home performance vs. away performance in games.
+All of which he showed significantly indicates that
+home-court helps teams play better. @Cao studied the effects
+of pressure on performance in the NBA. Using throws as their measure of
+interest, they tested whether home
+fans could distract and put pressure on opposing players to make free
+throws. However, they found insignificant
 evidence that home status has a substantial impact on missing from the free throw
-line. @Harris use two point shots, three-point shots and free throws as measures
-of interest to study home-court advantage. They find that two point shots are
-the strongest predictor of home-court advantage. They suggest that evidence 
-supports that home teams should try to shoot more two point shots and force
+line. @Harris used two point shots, three-point shots and free throws as measures
+of interest to study home-court advantage. Two point shots were found to be
+the strongest predictor of home-court advantage. They suggested
+that home teams should try to shoot more two point shots and force
 their opponent to take more two point shot attempts. This strategy will maximize
@@
 Some studies focus less on shooting and more on scoring differences and other metrics.
-For example, @Greer focuses his study on the influence of spectator booing on 
+For example, @Greer focused on the influence of spectator booing on
 home-court advantage in basketball. The three methods of performance used
-in this study were scoring, violations and turnovers. This study was conducted 
+in this study were scoring, violations, and turnovers. This study was conducted
 using the men’s basketball programs at two large universities. The study finds that
@@
 referee calls is less clear. However, the data does seem to lean slightly in
-favor of affecting visiting team performance. One study conducted by @Harville studied the
+favor of affecting visiting team performance.  @Harville studied the
 effect of home-court advantage using the 1991-1992 college basketball season.
@@
 during the college basketball season. This allowed them to construct two samples,
-one of home teams and one of neutral teams. They formulate their study as a 
-regression equation predicting the expected difference in score for home teams. 
+one of home teams and one of neutral teams. They formulate their study in a
+regression predicting the expected difference in score for home teams.
 They set up their study to find if the home teams won games by more points when
 they had home-court advantage vs. when playing at neutral court. This study
-succeeds in proving that home-court advantage exists.
+concluded with evidence supporting home-court advantage.
 
-There are also studies focused on the factors contributing to home-court advantage. 
-@Carron1992 cite 4 main game location factors for home and away teams.
-First, the crowd factor, which is the impact of fans cheering. Second, learning
-factors, which they categorize as the advantage from home teams from playing at
-a familiar venue. Third, travel factors, the idea that away teams may face 
-fatigue and jet lag from traveling. Last, rule factors, which says that home teams
+There are also surveys on the factors contributing to home-court advantage.
+@Carron1992 gave four main game location factors for home and away
+teams, namely, the crowd factor, which is the impact of fans cheering; learning
+factors, which is an advantage from home teams from playing at
+a familiar venue; travel factors, the idea that away teams may face
+fatigue and jet lag from traveling; and, rule factors, which says that home teams
 may benefit from some advantages in rules and officiating. They acknowledge that
@@
 by @Carron2005. The 2005 review goes over the new findings from studies
-about the significance of these 4 game location factors. Since 1992 they found 
-that results on these 4 factors are mixed. However, there is some evidence 
+about the significance of these four game location factors. Since 1992 they found
+that results on these four factors are mixed. However, there is some evidence
 supporting crowd and travel factors impact games in the NBA. There is less
@@
 # Data {#sec:data}
-Data was collected from the official NBA website. The main variables of interest
+
+\emph{need to give some descriptive statistics on how many games in
+each sample}
+
+Data were collected from the official NBA website. The main variables of interest
 are whether or not the home team won, scoring totals for home and away teams, and
@@
 percentages since 2010. Notice in Figure \@ref(fig:Fig1), the 10 years before 2020, the home team
-winning percentage ranged from around 0.56 to 0.7 and never dipped below .5. The 
-2020 bubble broke this historic pattern dipping down below .5. Foreshadowing, 
+winning percentage ranged from around 0.56 to 0.7 and never dipped below 0.5. The
+2020 bubble broke this historic pattern dipping down below 0.5. Foreshadowing,
 the confirmation of the expectation that the effect of home-court advantage was removed
@@
       scale_x_continuous(name = "Year", breaks = years, labels = years)+
-      scale_y_continuous(limits = c(0.4, 0.8))+
+      scale_y_continuous(name =  "Winning Percentage", limits = c(0.4, 0.8))+
       geom_hline(aes(yintercept = 0.5), color = "green")+
@@
 
-Moving on to the main focus of the study, comparing 2020 to 2017-2019. First, 
-a brief look at the distribution of scoring depicted in Figure \@ref(fig:Fig2).
-The green histograms indicate home and red histograms indicate away scoring. The
-top row is scoring for 2020 and the bottom is scoring for 2017-2019. All histograms
-seem to be fairly normally distributed which is important for statistical analysis.
-It’s hard to see any clear differences between years, but it’s important to look at
-and understand the distribution of measures before moving onto the analysis.
+Moving on to the main focus of the study, comparing 2020 to 2017-2019.
+Figure \@ref(fig:Fig2) shows the histograms of the home (green) and
+away scoring (red) for 2020 vs. 2017-2019. All histograms are
+fairly bell shaped, which is important for statistical tests
+designed for normally distribute data. There appears to be little
+difference between the 2020 and 2017-2019 for home scoring, but for
+away scoring, a noticeable shift to the right in 2020 is observed
+compared with that in 2017-2019.
 
-(ref:cap-Fig2) Shows the home(blue) and away (red) scoring distributions for 2020(bottom) and 2017-2019(top)
+(ref:cap-Fig2)  Histograms of home (blue) and away (red) scoring
+for 2020 (bottom) and 2017-2019 (top)
 ```{r Fig2, echo=FALSE, fig.cap="(ref:cap-Fig2)", fig.align='center'}
@@
 
+Our second target of inference is shooting percentage for home and away teams.
 Figure \@ref(fig:Fig3) shows home shooting for two pointers, three pointers and
@@
 similarly distributed between 2020 and 2017-19. Likewise, Figure \@ref(fig:Fig4),
-shows the same data except for away teams. Just like Figure \@ref(fig:Fig3),
-Figure \@ref(fig:Fig4) histograms don’t show any obvious differences. The normality
-of the distributions in both Figure \@ref(fig:Fig3) and Figure \@ref(fig:Fig4) are
-important for meeting the required assumptions for statistical analysis.
+shows the same percentages except for away teams. It appears that the
+two-point shooting percentage for away teams has a small shift to the
+right in 2020 relative to that in 2017-2019.
 
-(ref:cap-Fig3) Home shooting percentages for two pointers, three pointers and free throws for 2020(top) vs. 2017-19 (bottom)
+(ref:cap-Fig3) Histograms of home shooting percentages for two
+pointers, three pointers and free throws for 2020 (top) vs. 2017-19 (bottom)
 ```{r Fig3, echo=FALSE, fig.cap="(ref:cap-Fig3)"}
@@
 
-(ref:cap-Fig4) Away shooting percentages for two pointers, three pointers and free throws for 2020(top) vs. 2017-19 (bottom)"
+(ref:cap-Fig4) Histograms of away shooting percentages for two
+pointers, three pointers and free throws for 2020 (top) vs. 2017-19 (bottom)"
 ```{r Fig4, echo=FALSE, fig.cap="(ref:cap-Fig4)", fig.align='center'}
@@
 back to other playoffs. For this study, the 2020 playoffs were compared against
-the three previous playoffs collectively. It is important to make sure that any
-observed differences are truly from a home-court advantage. So, in order to control
-for the changing play style of the NBA, we will limit the study to 2020 vs 2017-2019.
-The reason being the faster pace play and more common use of the three-point shot 
+the three previous playoffs collectively. To control
+for the changing play style of the NBA, we limit the study to 2020 vs 2017-2019
+for the faster pace play and more common use of the three-point shot
 in modern basketball. If we used data from say 10 years ago, or earlier, observed
@@
 context for how home-court advantage specifically affects performance.
-Shooting percentages aren’t the only possible metrics affected by home-court
+Shooting percentages are not the only possible metrics affected by home-court
 advantage, but they are the most obvious and certainly an important one. In
-order to compare these differences simple two-sample z-tests were conducted.
+order to compare these differences simple two-sample $z$-tests were conducted.
 
-Given the approximate normality of data, 9 different z-tests were performed in 
-  order to test the effects of the COVID bubble on the 2020 NBA playoffs. The 9 questions were:  
-* Is the home team winning percentage in 2020 different than that it was in 2017-2019?  
-* Is the average home team scoring different in 2020 than it was over 2017-2019?  
-* Is the average away team scoring different in 2020 than it was over 2017-2019?  
-* Are home teams making two pointers at the same rate in 2020 as 2017-2019?  
-* Are home teams making three pointers at the same rate in 2020 as 2017-2019?  
-* Are home teams making free throws at the same rate in 2020 as 2017-2019?  
-* Are away teams making two pointers at the same rate in 2020 as 2017-2019?  
-* Are away teams making three pointers at the same rate in 2020 as 2017-2019?  
-* Are away teams making free throws at the same rate in 2020 as 2017-2019?
+We formulate the following nine specific research questions to test
+the effects of the COVID bubble on the 2020 NBA playoffs:
 
-A test for differences in means was conducted to compare scoring totals. Let
+1. Is the home team winning percentage in 2020 different than that it was in 2017-2019?
+1. Is the average home team scoring different in 2020 than it was over 2017-2019?
+1. Is the average away team scoring different in 2020 than it was over 2017-2019?
+1. Are home teams making two pointers at the same rate in 2020 as 2017-2019?
+1. Are home teams making three pointers at the same rate in 2020 as 2017-2019?
+1. Are home teams making free throws at the same rate in 2020 as 2017-2019?
+1. Are away teams making two pointers at the same rate in 2020 as 2017-2019?
+1. Are away teams making three pointers at the same rate in 2020 as 2017-2019?
+1. Are away teams making free throws at the same rate in 2020 as 2017-2019?
+
+A test for differences in means was conducted to compare scoring
+totals. The null hypothesis is
+\[
+H_0: \mu_1 = \mu_2,
+\]
+where $\mu_1$ and $\mu_2$ are the mean scoring total in 2020 and
+2017-2019, respectively. Let
 $\bar{X}_1$ represent the average scoring total for 2020 and $\bar{X}_2$ represent
-average scoring total over 2017-19. Similarly, $s_1^2$ represent the sample variance
+average scoring total over 2017-19. Similarly, let $s_1^2$ represent the sample variance
 of scoring for 2020 and $s_2^2$ represent sample variance of scoring over 2017-19.
 Finally, let $n_1$ represent the number of games in 2020 and $n_2$ represent the number
-of games from 2017-2019. Then the test statistic is constructed using Equation \@ref(eq:zmean).
+of games from 2017-2019. Then the test statistic is
 \begin{align}
-    z = \frac{(\bar{X}_1 - \bar{X}_2)-0}{\sqrt{\frac{s_1^2}{n_1}+\frac{s_2^2}{n_2}}}.
+    z = \frac{\bar{X}_1 - \bar{X}_2}{\sqrt{\frac{s_1^2}{n_1}+\frac{s_2^2}{n_2}}}.
     (\#eq:zmean)
 \end{align}
-We will test the null hypothesis that $\mu_1$=$\mu_2$. The test will be conducted twice
+The null distribution of the test statistic is standard normal.
+The test will be conducted twice
 once comparing home scoring in 2020 to 2017-19 and once comparing away
@@
 Tests for differences in proportions were conducted for comparing winning percentage
-and the three different shooting percentages in 2020 to those of 2017-2019. Let
+and the three different shooting percentages in 2020 to those of
+2017-2019. The null hypothesis is 
+\[
+H_0: p_1 = p_2,
+\]
+where $p_1$ and $p_2$ are the shooting percentage in 2020 and
+2017-2019, respectively. Let
 $\hat{p}_1$ represents sample proportion for 2020 and $\hat{p}_2$ represents sample
-proportion for 2017-2019. While, $\hat{p}$ is $\hat{p}_1$ and $\hat{p}_2$ pooled 
-together.Let $n_1$ represent number of observations for 2020 and $n_2$ represent 
-number of observations for 2017-2019. Then the test statistic is constructed using
-Equation \@ref(eq:zprop).
+proportion for 2017-2019. The test statistic is
 \begin{align}
-    z = \frac{(\hat{p}_1 - \hat{p}_2)-0}{\sqrt{\hat{p}(1-\hat{p})(\frac{1}{n_1}+\frac{1}{n_2})}}.
+    z = \frac{\hat{p}_1 - \hat{p}_2}{\sqrt{\hat{p}(1-\hat{p})(\frac{1}{n_1}+\frac{1}{n_2})}},
     (\#eq:zprop)
 \end{align}
-We will test the null hypothesis that $p_1$=$p_2$. The test will be conducted seven times,
+where $\hat p$ is the pooled shooting percentage of 2017-2020.
+The null distribution of the test statistic is also standard normal
+The test will be conducted seven times,
 once comparing home winning percentage for 2020 to that of 2017-2019 and once for each of
@@
 
-# Analysis and Results {#sec:results}
+# Results {#sec:results}
 ```{r, echo=FALSE}
@@
 
-Starting from the top of the Table \@ref(tab:table), we see a statistically significant change
+Starting from the top of the 
+Table \@ref(tab:table) summarizes results of the nine hypotheses
+including the testing statistics and their p-values. Also reported are
+the point estimates of the two samples in each comparison and the
+confidence intervals of the true difference between the two.
+
+First, we see a statistically significant change
 in home win percentage in 2020 from 2017-19, with p-value of 0.0365. The 95%
@@
 advantage was lost in the 2020 NBA playoffs. However, after accounting for
-multiple tests using Bonferroni correction the p-value is no longer significant.
+multiple tests using the Bonferroni correction, the p-value is no longer significant.
 So, we may only cautiously say there is evidence that home-court advantage was
@@
 throw shooting all show no significant difference, on average, between 2020 vs.
-2017-19. Home teams do not appear to have any evidence suggesting they played at 
+2017-19. There appear to be no evidence suggesting home teams played at
 a lower level in 2020 than they did in previous years when they had home-court advantage.
 
-Away teams saw more of an impact than home teams. For starters, they saw a 
+\emph{focus on the two tests that are significant after Bonferroni
+correction.}
+
+Away teams saw more of an impact than home teams. For starters, there is a
 significant increase in mean points per game by about 5 points (95% CI: 2.083, 7.988)
@@
 than previous years on the road. Starting from the dip in home winning percentage
-to below .482 it’s clear something was different. Although the difference was not
-significant after a Bonferroni correction it’s still important to consider and
+to below 0.482 it is clear that something was different. Although the difference was not
+significant after a Bonferroni correction it is still informative to consider and
 understand that home teams seemed to struggle to win compared to normal conditions.
@@
 better record than away teams boasting about a 60.5% win percentage in his sample,
-the 48.2% home winning percentage of 2020 home teams is a shift from normal.
-It’s clear that home teams did not benefit from the usual advantages provided by
+the 48.2% home winning percentage of 2020 home teams is quite a shift.
+In this study, home teams did not benefit from the usual advantages provided by
 being the home team.
 
-While the home teams in 2020 did see a slight uptick in average scoring it was
-not large enough to be statistically significant. However, away team average 
+Away team average
 scoring did increase by a statistically significant amount. This goes hand in
@@
 more reason to believe the conclusion that there was a significant decrease in
-home winning percentage in 2020, despite failing to be significant after Bonferroni
+home winning percentage in 2020, despite failing to be significant
+after the Bonferroni
 correction. Only away team scoring being significantly impacted by playing on
@@
 An interesting finding is all shooting and scoring numbers for both home and
-away teams did make at least small increases. Although these increases weren’t
+away teams did make at least small increases. Although these increases
+were not
 all significant these increases are exactly what is reported in @Carron2005
@@
 increase in winning. The improvement of away teams confirms a proposition from
-@Greer who believed that the positive social impact of crowds benefiting 
+@Greer that the positive social impact of crowds benefiting
 home teams may be a result of inhibiting away teams.
@@
 the beginning of possibilities for studies using the 2020 NBA bubble as a case study
-for home-court advantage.
-
-Unfortunately, this study is limited by a one time sample. It seems unlikely that 
-these conditions will ever be repeated. That being said, it will not be possible
-to have a follow-up study using the same measures with a different sample. If
-possible, that type of study could help strengthen the conclusions
-made in this paper.
+for home-court advantage. Although the study is limited by a one time
+sample, it seems unlikely that these conditions will ever be
+repeated. It may not be possible to have a follow-up study using the
+same measures with a different sample. Otherwise, that type of study
+could help strengthen the conclusion in this paper.
 

### 2020-11-12 Manuscript/NBABubbleAnalysis.Rmd
Update NBABubbleAnalysis.Rmd

@@
 for both home and away teams. The results from these tests bring a new perspective
-to understand of how home-court advantage impacts games by altering the performance
+to understanding of how home-court advantage impacts games by altering the performance
 of the home and away teams.

### 2020-11-12 Manuscript/NBABubbleAnalysis.Rmd
Update NBABubbleAnalysis.Rmd

@@
 one <- prop.test(x=c(sum(game_log_2020_home$W.L), sum(game_log_2017_2019_home$W.L)),
-                 n=c(nrow(game_log_2020_home), nrow(game_log_2017_2019_home)), correct=FALSE)
-rowone <- c(as.numeric(one$estimate[1]), as.numeric(one$estimate[2]), one$conf.int,
-          sqrt(as.numeric(one$statistic)), one$p.value)
+                 n=c(nrow(game_log_2020_home), nrow(game_log_2017_2019_home)))
+onew <- wilcox.test(x = game_log_2020_home$W.L, y = game_log_2017_2019_home$W.L)
+rowone <- c(as.numeric(one$estimate[1]), as.numeric(one$estimate[2]), one$p.value, onew$p.value)
 
@@
                 sd(game_log_2020_home$PTS), sigma.y = sd(game_log_2017_2019_home$PTS))
+twow <- wilcox.test(x=game_log_2020_home$PTS, y=game_log_2017_2019_home$PTS)
 rowtwo <- c(as.numeric(two$estimate[1])/100, as.numeric(two$estimate[2])/100,
-            two$conf.int, two$statistic, two$p.value)
+            two$p.value, twow$p.value)
 
@@
                   sd(game_log_2020_away$PTS), sigma.y = sd(game_log_2017_2019_away$PTS))
+threew <- wilcox.test(x=game_log_2020_away$PTS, y=game_log_2017_2019_away$PTS)
 rowthree <- c(as.numeric(three$estimate[1])/100, as.numeric(three$estimate[2])/100,
-              three$conf.int, three$statistic, three$p.value)
+              three$p.value, threew$p.value)
 
@@
                   n=c(sum(game_log_2020_home$X2PA), sum(game_log_2017_2019_home$X2PA)))
+fourw <- wilcox.test(x = game_log_2020_home$X2P., y = game_log_2017_2019_home$X2P.)
 rowfour <- c(as.numeric(four$estimate[1]), as.numeric(four$estimate[2]),
-             four$conf.int, sqrt(as.numeric(four$statistic)), four$p.value)
+              four$p.value, fourw$p.value)
 
@@
                   n=c(sum(game_log_2020_home$X3PA), sum(game_log_2017_2019_home$X3PA)))
+fivew <- wilcox.test(x = game_log_2020_home$X3P., y = game_log_2017_2019_home$X3P.)
 rowfive <- c(as.numeric(five$estimate[1]), as.numeric(five$estimate[2]),
-             five$conf.int,  sqrt(as.numeric(five$statistic)), five$p.value)
+            five$p.value, fivew$p.value)
 
@@
                  n=c(sum(game_log_2020_home$FTA), sum(game_log_2017_2019_home$FTA)))
-rowsix <- c(as.numeric(six$estimate[1]), as.numeric(six$estimate[2]), six$conf.int,
-           sqrt(as.numeric(six$statistic)), six$p.value)
+sixw <- wilcox.test(x = game_log_2020_home$FT., y = game_log_2017_2019_home$FT.)
+rowsix <- c(as.numeric(six$estimate[1]), as.numeric(six$estimate[2]),
+            six$p.value, sixw$p.value)
 
@@
                    n=c(sum(game_log_2020_away$X2PA), sum(game_log_2017_2019_away$X2PA)))
+sevenw <- wilcox.test(x = game_log_2020_away$X2P., y = game_log_2017_2019_away$X2P.)
 rowseven <- c(as.numeric(seven$estimate[1]), as.numeric(seven$estimate[2]),
-              seven$conf.int,  sqrt(as.numeric(seven$statistic)), seven$p.value)
+             seven$p.value, sevenw$p.value)
 
@@
                    n=c(sum(game_log_2020_away$X3PA), sum(game_log_2017_2019_away$X3PA)))
+eightw <- wilcox.test(x = game_log_2020_away$X3P., y = game_log_2017_2019_away$X3P.)
 roweight <- c(as.numeric(eight$estimate[1]), as.numeric(eight$estimate[2]),
-              eight$conf.int,  sqrt(as.numeric(eight$statistic)), eight$p.value)
+              eight$p.value, eightw$p.value)
 
@@
                   n=c(sum(game_log_2020_away$FTA), sum(game_log_2017_2019_away$FTA)))
+ninew <- wilcox.test(x = game_log_2020_away$FT., y = game_log_2017_2019_away$FT.)
 rownine <- c(as.numeric(nine$estimate[1]), as.numeric(nine$estimate[2]),
-             nine$conf.int,  sqrt(as.numeric(nine$statistic)), nine$p.value)
+            nine$p.value, ninew$p.value)
 
-titles <- c(2020, "2017-19", "95% CI LL", "95% CI UL", "Test Stat.", "P-value")
+titles <- c(2020, "2017-19", "Z-test P-value", "Wilcoxon P-value")
 rows <- c("Home Win %", "Home Scoring", "Away Scoring", "Home 2P%", "Home 3P%",
@@
 resultsfinal <- data.frame(t(results))
-resultsfinal1 <- round(resultsfinal[1:5], 3)
-resultsfinal2 <- round(resultsfinal[6], 4)
+resultsfinal1 <- round(resultsfinal[1:2], 3)
+resultsfinal2 <- round(resultsfinal[3:4], 4)
 resultsfinal <-cbind(resultsfinal1, resultsfinal2)

### 2020-11-12 Manuscript/NBABubbleAnalysis.Rmd
Update NBABubbleAnalysis.Rmd

@@
 Finally, let $n_1$ represent the number of games in 2020 and $n_2$ represent the number
-of games from 2017-2019. Then the test statistic is
-\begin{align}
-    z = \frac{\bar{X}_1 - \bar{X}_2}{\sqrt{\frac{s_1^2}{n_1}+\frac{s_2^2}{n_2}}}.
-    (\#eq:zmean)
-\end{align}
-The null distribution of the test statistic is standard normal.
-The test will be conducted twice
-once comparing home scoring in 2020 to 2017-19 and once comparing away
-scoring in 2020 to 2017-2019.
+of games from 2017-2019. The null distribution of the test statistic is standard normal.
+The test will be conducted twice once comparing home scoring in 2020 to 2017-19 and once
+comparing away scoring in 2020 to 2017-2019. These tests will also be simaltaneously 
+conducted using the Wilcoxon Test. The advantage of running the Wilcoxon Test is as a
+nonparametric test it does not rely on any normality assumption
 
@@
 The null distribution of the test statistic is also standard normal
-The test will be conducted seven times,
-once comparing home winning percentage for 2020 to that of 2017-2019 and once for each of
-two-point shooting, three-point shooting and free throws for both home and away teams
-in 2020 vs. 2017-2019.
+The test will be conducted seven times,once comparing home winning percentage for 
+2020 to that of 2017-2019 and once for each oftwo-point shooting, three-point shooting and
+free throws for both home and away teams in 2020 vs. 2017-2019. Again these tests will
+run simaltaneuously as a z-test and a Wilcoxon test.
 

### 2020-11-12 Manuscript/NBABubbleAnalysis.Rmd
Update NBABubbleAnalysis.Rmd

@@
 free throws. These variables were very popular and frequently used in the related
-literature discussed earlier. While many other measures could be used for measuring
+literature discussed earlier. The data was collected on a game by game basis, giving us one observation
+for each variable per game played. There were 83 games played in the 2020 playoffs. Giving us 83 observations 
+for each variable in 2020 for both home and away teams. Likewise, there was 243 games played over 
+2017-2019. Giving us 243 observations of each variable for both home and away teams over 2017-2020. 
+While many other measures could be used for measuring
 the outcome of the game and team performance, scoring seemed to be the most
@@
 The test will be conducted twice once comparing home scoring in 2020 to 2017-19 and once
-comparing away scoring in 2020 to 2017-2019. These tests will also be simaltaneously 
+comparing away scoring in 2020 to 2017-2019. These tests will also be simultaneously
 conducted using the Wilcoxon Test. The advantage of running the Wilcoxon Test is as a
@@
 free throws for both home and away teams in 2020 vs. 2017-2019. Again these tests will
-run simaltaneuously as a z-test and a Wilcoxon test.
+run simultaneously as a z-test and a Wilcoxon test.
 

### 2020-11-12 Manuscript/NBABubbleAnalysis.Rmd
Update NBABubbleAnalysis.Rmd

@@
 # Data {#sec:data}
-
-\emph{need to give some descriptive statistics on how many games in
-each sample}
-
 Data were collected from the official NBA website. The main variables of interest
@@
 comparing away scoring in 2020 to 2017-2019. These tests will also be simultaneously
-conducted using the Wilcoxon Test. The advantage of running the Wilcoxon Test is as a
+conducted using the Wilcoxon Test. The advantage of running the Wilcoxon test is as a
 nonparametric test it does not rely on any normality assumption
@@
 Starting from the top of the 
-Table \@ref(tab:table) summarizes results of the nine hypotheses
-including the testing statistics and their p-values. Also reported are
-the point estimates of the two samples in each comparison and the
-confidence intervals of the true difference between the two.
+Table \@ref(tab:table) summarizes p-values of the nine hypotheses for both
+z-tests and Wilcoxon tests. The p-values are all fairly similar for both
+tests giving strong confidence in conclusions drawn. Also reported are 
+the point estimates of the two samples in each comparison.
 
 First, we see a statistically significant change
-in home win percentage in 2020 from 2017-19, with p-value of 0.0365. The 95%
-confidence interval of (-0.255, -0.008) confirms our belief that home-court
-advantage was lost in the 2020 NBA playoffs. However, after accounting for
-multiple tests using the Bonferroni correction, the p-value is no longer significant.
+in home win percentage in 2020 from 2017-19, with p-value of 0.0365 for the
+z-test and 0.0369 for the Wilcoxon test. The 95% confidence interval of (-0.255, -0.008) 
+confirms our belief that home-court advantage was lost in the 2020 NBA playoffs. However, after accounting for
+multiple tests using the Bonferroni correction, the p-values for both tests are no longer significant.
 So, we may only cautiously say there is evidence that home-court advantage was
@@
 Home team performance did not seem to be negatively impacted by losing home-court
-advantage like expected. Home scoring, two point, three point shooting and free
-throw shooting all show no significant difference, on average, between 2020 vs.
-2017-19. There appear to be no evidence suggesting home teams played at
+advantage like expected. Home scoring, two-point and three-point shooting
+shooting all show no significant difference, on average, between 2020 vs.
+2017-19 based on p-values from both tests. However, the Wilcoxon test and z-test have conflicting results
+for free throws. The z-test p-value of 0.0692 indictaes no significant difference, while
+the Wilcoxon test p-value of 0.0469 indicates a difference at the 5% signficance level. Since,
+the p-value of Wilcoxon test is so close to significance level and neither p-value is signifcant
+after a Bonferonni correction for multiple tests this difference is likely not very meaningful.
+There appears to be no strong evidence suggesting home teams played at
 a lower level in 2020 than they did in previous years when they had home-court advantage.
 
-\emph{focus on the two tests that are significant after Bonferroni
-correction.}
-
 Away teams saw more of an impact than home teams. For starters, there is a
-significant increase in mean points per game by about 5 points (95% CI: 2.083, 7.988)
-Likewise, the away team two point shooting efficiency increased by an average of
-about 0.03 (95% CI: 0.015, 0.050). However, unlike two point shooting, away
-teams did not see a statistically significant difference in three point and free
+significant increase in mean points per game, indicated by p-value of 0.0008 for z-test and
+0.0004 for Wilcoxon. It is important to note both p-values also remain significant after
+a Bonferonni correction giving strong indication of significance. The average difference in points is estimated to 
+be about 5 points (95% CI: 2.083, 7.988).Likewise, the away team two point
+shooting efficiency increased significantly based on p-value of 0.0003
+for both the z-test and Wilcoxon test. Again both p-values remain signficant after Bonferroni correction.
+The average difference can be estimated to be about 0.03 (95% CI: 0.015, 0.050). However,
+unlike two point shooting, away teams did not see a statistically significant difference in three point and free
 throw shooting. Overall, away teams have evidence of change in performance in

### 2020-11-12 Manuscript/NBABubbleAnalysis.Rmd
Update NBABubbleAnalysis.Rmd

@@
 proportion for 2017-2019. The test statistic is
-\begin{align}
-    z = \frac{\hat{p}_1 - \hat{p}_2}{\sqrt{\hat{p}(1-\hat{p})(\frac{1}{n_1}+\frac{1}{n_2})}},
-    (\#eq:zprop)
-\end{align}
 where $\hat p$ is the pooled shooting percentage of 2017-2020.
-The null distribution of the test statistic is also standard normal
-The test will be conducted seven times,once comparing home winning percentage for 
-2020 to that of 2017-2019 and once for each oftwo-point shooting, three-point shooting and
+The null distribution of the test statistic is also standard normal.
+The test will be conducted seven times, once comparing home winning percentage for 
+2020 to that of 2017-2019 and once for each of two-point shooting, three-point shooting and
 free throws for both home and away teams in 2020 vs. 2017-2019. Again these tests will

### 2020-11-12 Manuscript/NBABubbleAnalysis.Rmd
Update NBABubbleAnalysis.Rmd

@@
 a Bonferonni correction giving strong indication of significance. The average difference in points is estimated to 
-be about 5 points (95% CI: 2.083, 7.988).Likewise, the away team two point
+be about 5 points (95% CI: 2.083, 7.988). Likewise, the away team two point
 shooting efficiency increased significantly based on p-value of 0.0003
-for both the z-test and Wilcoxon test. Again both p-values remain signficant after Bonferroni correction.
+for both the z-test and Wilcoxon test. Again, both p-values remain signficant after Bonferroni correction.
 The average difference can be estimated to be about 0.03 (95% CI: 0.015, 0.050). However,
-unlike two point shooting, away teams did not see a statistically significant difference in three point and free
+unlike two-point shooting, away teams did not see a statistically significant difference in three-point and free
 throw shooting. Overall, away teams have evidence of change in performance in

### 2020-11-12 Manuscript/NBABubbleAnalysis.Rmd
Update NBABubbleAnalysis.Rmd

@@
 The test will be conducted twice once comparing home scoring in 2020 to 2017-19 and once
-comparing away scoring in 2020 to 2017-2019. These tests will also be simultaneously
-conducted using the Wilcoxon Test. The advantage of running the Wilcoxon test is as a
+comparing away scoring in 2020 to 2017-2019. These hypotheses will also be simultaneously
+tested using the Wilcoxon Test. The advantage of running the Wilcoxon test is as a
 nonparametric test it does not rely on any normality assumption
@@
 2020 to that of 2017-2019 and once for each of two-point shooting, three-point shooting and
-free throws for both home and away teams in 2020 vs. 2017-2019. Again these tests will
-run simultaneously as a z-test and a Wilcoxon test.
+free throws for both home and away teams in 2020 vs. 2017-2019. These hypotheses will also be
+tested using both a z-test and a Wilcoxon test.
 

### 2020-11-12 Manuscript/NBABubbleAnalysis.Rmd
Update NBABubbleAnalysis.Rmd

@@
 2017-19 based on p-values from both tests. However, the Wilcoxon test and z-test have conflicting results
-for free throws. The z-test p-value of 0.0692 indictaes no significant difference, while
-the Wilcoxon test p-value of 0.0469 indicates a difference at the 5% signficance level. Since,
-the p-value of Wilcoxon test is so close to significance level and neither p-value is signifcant
+for free throws. The z-test p-value of 0.0692 indicates no significant difference, while
+the Wilcoxon test p-value of 0.0469 indicates a difference at the 5% significance level. Since,
+the p-value of Wilcoxon test is so close to significance level and neither p-value is significant
 after a Bonferonni correction for multiple tests this difference is likely not very meaningful.

### 2020-11-12 Manuscript/NBABubbleAnalysis.Rmd
Update NBABubbleAnalysis.Rmd
comments on data, especially winning ratio

@@
 for each variable per game played. There were 83 games played in the 2020 playoffs. Giving us 83 observations 
-for each variable in 2020 for both home and away teams. Likewise, there was 243 games played over 
+for each variable in 2020 for both home and away teams. Likewise, there were 243 games played over 
 2017-2019. Giving us 243 observations of each variable for both home and away teams over 2017-2020. 
@@
 efficiency are important to understand.
+
+\emph{Does this mean that the data for winning percentage are binary
+(1/0)?}
 
@@
 
+\emph{Is shooting percentage the right terminology? It seems to mean
+something else (google true shooting percentage)}
+
 A test for differences in means was conducted to compare scoring
@@
 where $\mu_1$ and $\mu_2$ are the mean scoring total in 2020 and
-2017-2019, respectively. Let
-$\bar{X}_1$ represent the average scoring total for 2020 and $\bar{X}_2$ represent
-average scoring total over 2017-19. Similarly, let $s_1^2$ represent the sample variance
-of scoring for 2020 and $s_2^2$ represent sample variance of scoring over 2017-19.
-Finally, let $n_1$ represent the number of games in 2020 and $n_2$ represent the number
-of games from 2017-2019. The null distribution of the test statistic is standard normal.
+2017-2019, respectively. We considered both the $t$-test and the
+Wilcoxon rank-sum test as implemented in R [@R]. The Wilcoxon rank-sum
+test is preferred because, unlike the $t$-test, it is distribution-free.
 The test will be conducted twice once comparing home scoring in 2020 to 2017-19 and once
-comparing away scoring in 2020 to 2017-2019. These hypotheses will also be simultaneously
-tested using the Wilcoxon Test. The advantage of running the Wilcoxon test is as a
-nonparametric test it does not rely on any normality assumption
+comparing away scoring in 2020 to 2017-2019.
 
+\emph{What data were used for the test for the winning percentage?}
 Tests for differences in proportions were conducted for comparing winning percentage
@@
 where $p_1$ and $p_2$ are the shooting percentage in 2020 and
-2017-2019, respectively. Let
+2017-2019, respectively. 
 $\hat{p}_1$ represents sample proportion for 2020 and $\hat{p}_2$ represents sample
@@
 
+\emph{In the table, we have shooting rate and winning rate instead of
+percentages. So remove the percentage sign and make sure consistency throughout}
+
 Starting from the top of the 
@@
 shooting efficiency increased significantly based on p-value of 0.0003
-for both the z-test and Wilcoxon test. Again, both p-values remain signficant after Bonferroni correction.
+for both the z-test and Wilcoxon test. Again, both p-values remain significant after Bonferroni correction.
 The average difference can be estimated to be about 0.03 (95% CI: 0.015, 0.050). However,

### 2020-11-13 Manuscript/NBABubbleAnalysis.Rmd
Update NBABubbleAnalysis.Rmd

@@
 of this study. The specific measures of interest are team scoring totals and
-team shooting percentage on two pointers, three pointers, and free throws.
+team shooting percentage on two-pointers, three-pointers, and free throws.
 Comparing these measures for home teams and away teams in 2020 vs. 2017-2019
@@
 differences in away scoring in 2020 vs 2017-2019. Also, differences in team
-shooting (for two pointers, three pointers, and free throws) from 2020 vs 2017-2019
+shooting (for-two pointers, three-pointers, and free throws) from 2020 vs 2017-2019
 for both home and away teams. The results from these tests bring a new perspective
@@
 are whether or not the home team won, scoring totals for home and away teams, and
-shooting percentages for home and away teams on two pointers, three pointers, and
+shooting percentages for home and away teams on two-pointers, three-pointers, and
 free throws. These variables were very popular and frequently used in the related
-literature discussed earlier. The data was collected on a game by game basis, giving us one observation
-for each variable per game played. There were 83 games played in the 2020 playoffs. Giving us 83 observations 
-for each variable in 2020 for both home and away teams. Likewise, there were 243 games played over 
-2017-2019. Giving us 243 observations of each variable for both home and away teams over 2017-2020. 
+literature discussed earlier. The data was collected on a game by game basis, this gave us two observations
+for each variable per game played, one observation for each team(home and away).
+There were 83 games played in the 2020 playoffs. Giving us 83 observations 
+for each variable in 2020 for both the home and away teams(166 observations total). 
+Likewise, there were 243 games played over 2017-2019. Giving us 243 observations of each variable 
+for both home and away teams over 2017-2020(486 observations total).
 While many other measures could be used for measuring
 the outcome of the game and team performance, scoring seemed to be the most
-important. The winner of a game is determined by who scores more points. So,
-naturally when studying the outcome of winning, changes in scoring and scoring
-efficiency are important to understand.
-
-\emph{Does this mean that the data for winning percentage are binary
-(1/0)?}
+important. The winner of a game is determined by who scores more points. There can only be
+one winner and one loser making the outcome a binary variable. Each game has a winner 
+denoted by a 1 in the W.L column of our data and a loser denoted by 0. Every game must result in 
+exactly one team getting a 1 and exactly one team getting a 0. Winning percentage for home
+(or away) teams can be conducted by summing the 1's for those teams together then dividing 
+by the number of games played.
 
@@
 Our second target of inference is shooting percentage for home and away teams.
-Figure \@ref(fig:Fig3) shows home shooting for two pointers, three pointers and
+Figure \@ref(fig:Fig3) shows home shooting for two-pointers, three-pointers and
 free throws for 2020 (top) vs. 2017-19 (bottom). The histograms appear to be fairly
@@
 
-(ref:cap-Fig3) Histograms of home shooting percentages for two
-pointers, three pointers and free throws for 2020 (top) vs. 2017-19 (bottom)
+(ref:cap-Fig3) Histograms of home shooting percentages for two-
+pointers, three-pointers and free throws for 2020 (top) vs. 2017-19 (bottom)
 ```{r Fig3, echo=FALSE, fig.cap="(ref:cap-Fig3)"}
@@
 
-(ref:cap-Fig4) Histograms of away shooting percentages for two
-pointers, three pointers and free throws for 2020 (top) vs. 2017-19 (bottom)"
+(ref:cap-Fig4) Histograms of away shooting percentages for two-
+pointers, three-pointers and free throws for 2020 (top) vs. 2017-19 (bottom)"
 ```{r Fig4, echo=FALSE, fig.cap="(ref:cap-Fig4)", fig.align='center'}
@@
 1. Is the average away team scoring different in 2020 than it was over 2017-2019?
-1. Are home teams making two pointers at the same rate in 2020 as 2017-2019?
-1. Are home teams making three pointers at the same rate in 2020 as 2017-2019?
-1. Are home teams making free throws at the same rate in 2020 as 2017-2019?
-1. Are away teams making two pointers at the same rate in 2020 as 2017-2019?
-1. Are away teams making three pointers at the same rate in 2020 as 2017-2019?
+1. Are home teams making two-pointers at the same rate in 2020 as 2017-2019?
+1. Are home teams making three-pointers at the same rate in 2020 as 2017-2019?
+1. Are home teams making freethrows at the same rate in 2020 as 2017-2019?
+1. Are away teams making two-pointers at the same rate in 2020 as 2017-2019?
+1. Are away teams making three-pointers at the same rate in 2020 as 2017-2019?
 1. Are away teams making free throws at the same rate in 2020 as 2017-2019?
-
-\emph{Is shooting percentage the right terminology? It seems to mean
-something else (google true shooting percentage)}
 
@@
 
-\emph{What data were used for the test for the winning percentage?}
 Tests for differences in proportions were conducted for comparing winning percentage
@@
 where $p_1$ and $p_2$ are the shooting percentage in 2020 and
-2017-2019, respectively. 
+2017-2019, respectively. Or $p_1$ and $p_2$ are the home winning percentage for 2020 and
+2017-2019, respectively. This proportion comes from summing the number of home teams wins
+and dividing by the total number of games played.
 $\hat{p}_1$ represents sample proportion for 2020 and $\hat{p}_2$ represents sample
 proportion for 2017-2019. The test statistic is
-where $\hat p$ is the pooled shooting percentage of 2017-2020.
+where $\hat p$ is the pooled shooting percentage of 2017-2020. Or where $\hat p$ is the 
+pooled home team winning percentage of 2017-2020.
 The null distribution of the test statistic is also standard normal.
@@
 titles <- c(2020, "2017-19", "Z-test P-value", "Wilcoxon P-value")
-rows <- c("Home Win %", "Home Scoring", "Away Scoring", "Home 2P%", "Home 3P%",
-          "Home FT%", "Away 2P%", "Away 3P%", "Away FT%")
+rows <- c("Home Win ", "Home Scoring", "Away Scoring", "Home 2P", "Home 3P",
+          "Home FT", "Away 2P", "Away 3P", "Away FT")
 results <- data.frame(rowone, rowtwo, rowthree, rowfour, rowfive, rowsix,
@@
 ```
-
-\emph{In the table, we have shooting rate and winning rate instead of
-percentages. So remove the percentage sign and make sure consistency throughout}
 
@@
 a Bonferonni correction giving strong indication of significance. The average difference in points is estimated to 
-be about 5 points (95% CI: 2.083, 7.988). Likewise, the away team two point
+be about 5 points (95% CI: 2.083, 7.988). Likewise, the away team two-point
 shooting efficiency increased significantly based on p-value of 0.0003
@@
 At least some of that improvement from away teams came from significantly higher
-two point efficiency. This corresponds with the conclusion from @Harris,
-where they found home teams are best suited to capitalize on advantages from two point
-shots. Normally, by shooting more two pointers themselves and forcing away teams to
-shoot more two pointers the home team benefits from increasing effects of
-home-court advantage. However, with away teams significantly improving two point
+two-point efficiency. This corresponds with the conclusion from @Harris,
+where they found home teams are best suited to capitalize on advantages from two-point
+shots. Normally, by shooting more two-pointers themselves and forcing away teams to
+shoot more two-pointers the home team benefits from increasing effects of
+home-court advantage. However, with away teams significantly improving two-point
 shooting in the bubble this strategy was no longer viable and home-court

### 2020-11-13 Manuscript/NBABubbleAnalysis.Rmd
should use fisher.test for winning ratio

@@
 for each variable per game played, one observation for each team(home and away).
-There were 83 games played in the 2020 playoffs. Giving us 83 observations 
-for each variable in 2020 for both the home and away teams(166 observations total). 
-Likewise, there were 243 games played over 2017-2019. Giving us 243 observations of each variable 
-for both home and away teams over 2017-2020(486 observations total).
+There were 83 games played in the 2020 playoffs, giving 83 observations 
+for each variable in 2020 for both the home and away teams (166 observations total). 
+Likewise, there were 243 games played over 2017-2019,  giving 243 observations of each variable 
+for both home and away teams over 2017-2020 (486 observations total).
 While many other measures could be used for measuring
@@
 important. The winner of a game is determined by who scores more points. There can only be
-one winner and one loser making the outcome a binary variable. Each game has a winner 
-denoted by a 1 in the W.L column of our data and a loser denoted by 0. Every game must result in 
-exactly one team getting a 1 and exactly one team getting a 0. Winning percentage for home
-(or away) teams can be conducted by summing the 1's for those teams together then dividing 
-by the number of games played.
+one winner and one loser making the outcome a binary variable, with
+one indicating a win and zero indicating a loss.
 
@@
 comparing away scoring in 2020 to 2017-2019.
+
+\emph{separate winning percentage from the other comparisons. Use
+fisher.test for it}
 

### 2020-11-13 Manuscript/NBABubbleAnalysis.Rmd
Update NBABubbleAnalysis.Rmd

@@
 free throws for both home and away teams in 2020 vs. 2017-2019. These hypotheses will also be
-tested using both a z-test and a Wilcoxon test.
+tested using both a z-test and a nonparametric test. The Wilcoxon test will be used for shooting percentages and
+Fisher's test will be used to test the binary variable of home team winning.
 
@@
                  n=c(nrow(game_log_2020_home), nrow(game_log_2017_2019_home)))
+X2020 <- table(game_log_2020_home$W.L)
+X17_19 <- table(game_log_2017_2019_home$W.L)
+homewin<- rbind(X2020, X17_19)
 onew <- wilcox.test(x = game_log_2020_home$W.L, y = game_log_2017_2019_home$W.L)
@@
 
-titles <- c(2020, "2017-19", "Z-test P-value", "Wilcoxon P-value")
+titles <- c(2020, "2017-19", "Z-test P-value", "Nonparametric P-value")
 rows <- c("Home Win ", "Home Scoring", "Away Scoring", "Home 2P", "Home 3P",
@@
 in home win percentage in 2020 from 2017-19, with p-value of 0.0365 for the
-z-test and 0.0369 for the Wilcoxon test. The 95% confidence interval of (-0.255, -0.008) 
+z-test and 0.0400 for Fisher's test. The 95% confidence interval of (-0.255, -0.008) 
 confirms our belief that home-court advantage was lost in the 2020 NBA playoffs. However, after accounting for

### 2020-11-13 Manuscript/NBABubbleAnalysis.Rmd
Update NBABubbleAnalysis.Rmd

@@
 homewin<- rbind(X2020, X17_19)
-onew <- wilcox.test(x = game_log_2020_home$W.L, y = game_log_2017_2019_home$W.L)
+onew <- fisher.test(x = homewin)
 rowone <- c(as.numeric(one$estimate[1]), as.numeric(one$estimate[2]), one$p.value, onew$p.value)

### 2020-11-13 Manuscript/NBABubbleAnalysis.Rmd
Update NBABubbleAnalysis.Rmd

@@
 First, we see a statistically significant change
-in home win percentage in 2020 from 2017-19, with p-value of 0.0365 for the
+in home win percentage in 2020 from 2017-19, with p-value of 0.0497 for the
 z-test and 0.0400 for Fisher's test. The 95% confidence interval of (-0.255, -0.008) 

### 2020-11-14 Manuscript/NBABubbleAnalysis.Rmd
Update NBABubbleAnalysis.Rmd

@@
 
-\emph{separate winning percentage from the other comparisons. Use
-fisher.test for it}
+Tests for differences in proportions was conducted for comparing winning percentage of 
+home teams in 2020 to those of 2017-2019. The null hypothesis is 
+\[
+H_0: p_1 = p_2,
+\]
+where $p_1$ and $p_2$ are the home winning percentage for 2020 and
+2017-2019, respectively. $\hat{p}_1$ represents sample proportion for 2020 and $\hat{p}_2$ represents sample
+proportion for 2017-2019. This proportion comes from summing the number of home teams wins
+and dividing by the total number of games played. The test statistic is where $\hat p$ is the 
+pooled home team winning percentage of 2017-2020. The null distribution of the test statistic is also standard normal.
+This hypothesis will be tested using both a z-test and Fisher's test. Fisher's test is preferred, like Wilcoxon, 
+because it is distribution free.
 
-Tests for differences in proportions were conducted for comparing winning percentage
-and the three different shooting percentages in 2020 to those of
-2017-2019. The null hypothesis is 
+Tests for differences in proportions were also conducted for comparing shooting percentage on 
+two-pointers, three-pointers and free throws for home and away teams in 2020 to those of 2017-2019. 
+The null hypothesis is 
 \[
@@
 where $p_1$ and $p_2$ are the shooting percentage in 2020 and
-2017-2019, respectively. Or $p_1$ and $p_2$ are the home winning percentage for 2020 and
-2017-2019, respectively. This proportion comes from summing the number of home teams wins
-and dividing by the total number of games played.
-$\hat{p}_1$ represents sample proportion for 2020 and $\hat{p}_2$ represents sample
+2017-2019, respectively. $\hat{p}_1$ represents sample proportion for 2020 and $\hat{p}_2$ represents sample
 proportion for 2017-2019. The test statistic is
-where $\hat p$ is the pooled shooting percentage of 2017-2020. Or where $\hat p$ is the 
-pooled home team winning percentage of 2017-2020.
+where $\hat p$ is the pooled shooting percentage of 2017-2020.
 The null distribution of the test statistic is also standard normal.
-The test will be conducted seven times, once comparing home winning percentage for 
-2020 to that of 2017-2019 and once for each of two-point shooting, three-point shooting and
+The test will be conducted six times, once for each of two-point shooting, three-point shooting and
 free throws for both home and away teams in 2020 vs. 2017-2019. These hypotheses will also be
-tested using both a z-test and a nonparametric test. The Wilcoxon test will be used for shooting percentages and
-Fisher's test will be used to test the binary variable of home team winning.
+tested using both a z-test and the nonparametric Wilcoxon test.
 

### 2020-11-15 Manuscript/NBABubbleAnalysis.Rmd
smoothed text on testing

--- 2020-11-13 Manuscript/NBABubbleAnalysis.Rmd
@@
 Shooting percentages are not the only possible metrics affected by home-court
-advantage, but they are the most obvious and certainly an important one. In
-order to compare these differences simple two-sample $z$-tests were conducted.
+advantage, but they are the most obvious and certainly an important one.
 
@@
 
-A test for differences in means was conducted to compare scoring
-totals. The null hypothesis is
-\[
-H_0: \mu_1 = \mu_2,
-\]
-where $\mu_1$ and $\mu_2$ are the mean scoring total in 2020 and
-2017-2019, respectively. We considered both the $t$-test and the
-Wilcoxon rank-sum test as implemented in R [@R]. The Wilcoxon rank-sum
-test is preferred because, unlike the $t$-test, it is distribution-free.
-The test will be conducted twice once comparing home scoring in 2020 to 2017-19 and once
-comparing away scoring in 2020 to 2017-2019.
+All nine questions can be approached by a standard two-sample
+comparison with the $z$-test. The $z$-test statistic follows a
+standard normal distribution, which is a good approximation based on
+the central limit theorem given the sample size in this application.
 
-\emph{separate winning percentage from the other comparisons. Use
-fisher.test for it}
+We also conducted nonparametric tests that are distribution free to
+confirm the results from the $z$-test. For question~1, we used
+Fisher's exact test for a contigency table which summarizes the wins
+and losses of the home team in the 83 games in 2020 and the 243 games
+in 2017-2019. For all other eight questions, the data are the scores
+or shooting percentages from the 84 games in 2020 and the 243 games in
+2017-2019. We used Wilcoxan's rank-sum test.
 
-Tests for differences in proportions were conducted for comparing winning percentage
-and the three different shooting percentages in 2020 to those of
-2017-2019. The null hypothesis is 
-\[
-H_0: p_1 = p_2,
-\]
-where $p_1$ and $p_2$ are the shooting percentage in 2020 and
-2017-2019, respectively. Or $p_1$ and $p_2$ are the home winning percentage for 2020 and
-2017-2019, respectively. This proportion comes from summing the number of home teams wins
-and dividing by the total number of games played.
-$\hat{p}_1$ represents sample proportion for 2020 and $\hat{p}_2$ represents sample
-proportion for 2017-2019. The test statistic is
-where $\hat p$ is the pooled shooting percentage of 2017-2020. Or where $\hat p$ is the 
-pooled home team winning percentage of 2017-2020.
-The null distribution of the test statistic is also standard normal.
-The test will be conducted seven times, once comparing home winning percentage for 
-2020 to that of 2017-2019 and once for each of two-point shooting, three-point shooting and
-free throws for both home and away teams in 2020 vs. 2017-2019. These hypotheses will also be
-tested using both a z-test and a nonparametric test. The Wilcoxon test will be used for shooting percentages and
-Fisher's test will be used to test the binary variable of home team winning.
+All three tests, namely the $z$-test, Fisher's exact test, and
+Wilcoxan's rank-sum test, were performed using R [@R].
 
@@
 Table \@ref(tab:table) summarizes p-values of the nine hypotheses for both
-z-tests and Wilcoxon tests. The p-values are all fairly similar for both
+$z$-tests and Wilcoxon tests. The p-values are all fairly similar for both
 tests giving strong confidence in conclusions drawn. Also reported are 
@@
 in home win percentage in 2020 from 2017-19, with p-value of 0.0497 for the
-z-test and 0.0400 for Fisher's test. The 95% confidence interval of (-0.255, -0.008) 
+$z$-test and 0.0400 for Fisher's exact test. The 95% confidence
+interval (CI)  of $(-0.255, -0.008)$ 
 confirms our belief that home-court advantage was lost in the 2020 NBA playoffs. However, after accounting for
@@
 shooting all show no significant difference, on average, between 2020 vs.
-2017-19 based on p-values from both tests. However, the Wilcoxon test and z-test have conflicting results
-for free throws. The z-test p-value of 0.0692 indicates no significant difference, while
+2017-19 based on p-values from both tests. However, the Wilcoxon test and $z$-test have conflicting results
+for free throws. The $z$-test p-value of 0.0692 indicates no significant difference, while
 the Wilcoxon test p-value of 0.0469 indicates a difference at the 5% significance level. Since,
@@
 Away teams saw more of an impact than home teams. For starters, there is a
-significant increase in mean points per game, indicated by p-value of 0.0008 for z-test and
+significant increase in mean points per game, indicated by p-value of 0.0008 for $z$-test and
 0.0004 for Wilcoxon. It is important to note both p-values also remain significant after
-a Bonferonni correction giving strong indication of significance. The average difference in points is estimated to 
-be about 5 points (95% CI: 2.083, 7.988). Likewise, the away team two-point
+a Bonferonni correction giving strong indication of significance. The average difference in points was estimated to 
+be about 5 points, with 95% CI $(2.083, 7.988)$. Likewise, the away team two-point
 shooting efficiency increased significantly based on p-value of 0.0003
-for both the z-test and Wilcoxon test. Again, both p-values remain significant after Bonferroni correction.
-The average difference can be estimated to be about 0.03 (95% CI: 0.015, 0.050). However,
+for both the $z$-test and Wilcoxon test. Again, both p-values remain significant after Bonferroni correction.
+The average difference was estimated to be about 0.03, with 95% CI $(0.015, 0.050)$. However,
 unlike two-point shooting, away teams did not see a statistically significant difference in three-point and free

### 2020-12-01 Manuscript/NBABubbleAnalysis.Rmd
set up to separate the tex source

@@
 biblio-style: asa
-output: bookdown::pdf_document2
+output: 
+  bookdown::pdf_document2:
+    keep_tex: yes
 header-includes:

### 2020-12-01 Manuscript/manuscript.tex
added tex source for submission

--- 2020-11-15 Manuscript/NBABubbleAnalysis.Rmd
@@
----
-title: "The Effects of the NBA COVID Bubble on the NBA Playoffs"
-subtitle: "A Case Study for Home-Court Advantage"
-author: |
-  | Michael Price and Jun Yan
-  | michael.price@uconn.edu
-  | Department of Statistics, University of Connecticut
-date: "`r format(Sys.time(), '%d, %B, %Y')`"
-documentclass: article
-papersize: letter
-fontsize: 11pt
-bibliography: citations.bib
-biblio-style: asa
-output: bookdown::pdf_document2
-header-includes:
-   - \usepackage[margin=1in]{geometry}
-   - \usepackage {hyperref}
-   - \hypersetup {colorlinks = true, linkcolor = blue, citecolor=blue, urlcolor = blue}
-toc: False
----
-Abstract:
+\documentclass[12pt, letterpaper, titlepage]{article}
+
+\usepackage{amsmath}
+\usepackage{booktabs}
+\usepackage{amsthm}
+\usepackage{graphicx}
+\usepackage[margin=1in]{geometry}
+\usepackage {hyperref}
+\hypersetup {colorlinks = true, linkcolor = blue, citecolor=blue, urlcolor = blue}
+\usepackage[]{natbib}
+\usepackage{enumitem}
+\usepackage{setspace}
+
+\usepackage[pagewise]{lineno}
+\linenumbers*[1]
+% %% patches to make lineno work better with amsmath
+\newcommand*\patchAmsMathEnvironmentForLineno[1]{%
+ \expandafter\let\csname old#1\expandafter\endcsname\csname #1\endcsname
+ \expandafter\let\csname oldend#1\expandafter\endcsname\csname end#1\endcsname
+ \renewenvironment{#1}%
+ {\linenomath\csname old#1\endcsname}%
+ {\csname oldend#1\endcsname\endlinenomath}}%
+\newcommand*\patchBothAmsMathEnvironmentsForLineno[1]{%
+ \patchAmsMathEnvironmentForLineno{#1}%
+ \patchAmsMathEnvironmentForLineno{#1*}}%
+
+\AtBeginDocument{%
+ \patchBothAmsMathEnvironmentsForLineno{equation}%
+ \patchBothAmsMathEnvironmentsForLineno{align}%
+ \patchBothAmsMathEnvironmentsForLineno{flalign}%
+ \patchBothAmsMathEnvironmentsForLineno{alignat}%
+ \patchBothAmsMathEnvironmentsForLineno{gather}%
+ \patchBothAmsMathEnvironmentsForLineno{multline}%
+}
+
+\title{The Effects of the NBA COVID Bubble on the NBA Playoffs: A Case Study for Home-Court Advantage}
+
+% \author{Michael Price and Jun Yan\\
+% \href{mailto:michael.price@uconn.edu}{\nolinkurl{michael.price@uconn.edu}}\\
+% Department of Statistics, University of Connecticut}
+\author{annonymous}
+\date{December 1, 2020}
+
+\begin{document}
+\maketitle
+
+\doublespace
+
+\begin{abstract}
 The 2020 NBA playoffs were played inside of a bubble in Disney World because of
@@
 team shooting percentage on two-pointers, three-pointers, and free throws.
-Comparing these measures for home teams and away teams in 2020 vs. 2017-2019
+Comparing these measures for home teams and away teams in 2020 vs.~2017-2019
 shows that the 2020 playoffs favored away teams more than usual, particularly
 with two point shooting and total scoring.
+\end{abstract}
 
-```{r setup, echo = FALSE, message =  FALSE, warning = FALSE}
-library(BSDA)
-library(ggplot2)
-knitr::opts_chunk$set(fig.width = 6, fig.height = 4, dpi = 300,
-out.width = "90%", fig.align = "center", dev = "pdf")
-```
+\hypertarget{sec:intro}{%
+\section{Introduction}\label{sec:intro}}
 
-# Introduction {#sec:intro}
 Home-court advantage is often discussed in sports circles as a contributing
@@
 brought a great deal of concern for fans, teams, journalists, and others.
-For example, @Aschburner discusses about anticipated effects, sharing
+For example, \citet{Aschburner} discusses about anticipated effects, sharing
 concerns from former players, coaches and other experts about
 the potential effects of removing home-court advantage. Aschburner notes that the
-NBA did make attempts to recreate the effects by putting the "home” team logo on
-the court and allowing the "home" team to play crowd noise and music, but most
+NBA did make attempts to recreate the effects by putting the ``home'' team logo on
+the court and allowing the ``home'' team to play crowd noise and music, but most
 people doubted these small attempts would recreate a true playoff atmosphere.
-During the 2020 NBA playoffs, home teams only won about 48.2% of the games. This
-is lower than normal, which Aschburner claims usually floats around 60%. This
+During the 2020 NBA playoffs, home teams only won about 48.2\% of the games. This
+is lower than normal, which Aschburner claims usually floats around 60\%. This
 shift in the home team winning percentage surely indicates the opportunity for
@@
 home-court advantage. Some of these studies are reviewed in greater detail in
-Section~\@ref(sec:litrev).
+Section\textasciitilde\ref{sec:litrev}.
 
 We will compare home team performance in 2020 at a neutral site with
-no fans vs. 2017-19 playoffs with fans. Likewise, away team performance in 2020
-at a neutral site with no fans vs. 2017-19 playoffs with fans. By comparing home
+no fans vs.~2017-19 playoffs with fans. Likewise, away team performance in 2020
+at a neutral site with no fans vs.~2017-19 playoffs with fans. By comparing home
 teams in 2020 to home teams in 2017-19 and away teams in 2020 to away teams in
@@
 
-Nine hypotheses were tested to understand the differences in 2020 vs. earlier
+Nine hypotheses were tested to understand the differences in 2020 vs.~earlier
 years. First, whether or not the difference between home win percentage in 2020
@@
 
-# Literature Review {#sec:litrev}
+\hypertarget{sec:litrev}{%
+\section{Literature Review}\label{sec:litrev}}
+
 There is a voluminous literature on the effects of home-court advantage.
 Many NBA home-court advantage studies analyze the effects by studying shooting
-percentages. @Kotecki reported significant home-court advantage using performance-based
+percentages. \citet{Kotecki} reported significant home-court advantage using performance-based
 statistics, specifically field goal percentage, free throw percentage, and
-points by comparing home performance vs. away performance in games.
+points by comparing home performance vs.~away performance in games.
 All of which he showed significantly indicates that
-home-court helps teams play better. @Cao studied the effects
+home-court helps teams play better. \citet{Cao} studied the effects
 of pressure on performance in the NBA. Using throws as their measure of
@@
 evidence that home status has a substantial impact on missing from the free throw
-line. @Harris used two point shots, three-point shots and free throws as measures
+line. \citet{Harris} used two point shots, three-point shots and free throws as measures
 of interest to study home-court advantage. Two point shots were found to be
@@
 Some studies focus less on shooting and more on scoring differences and other metrics.
-For example, @Greer focused on the influence of spectator booing on
+For example, \citet{Greer} focused on the influence of spectator booing on
 home-court advantage in basketball. The three methods of performance used
 in this study were scoring, violations, and turnovers. This study was conducted
-using the men’s basketball programs at two large universities. The study finds that
+using the men's basketball programs at two large universities. The study finds that
 social support, like booing, is an important contributor to home-court advantage.
@@
 referee calls is less clear. However, the data does seem to lean slightly in
-favor of affecting visiting team performance.  @Harville studied the
+favor of affecting visiting team performance. \citet{Harville} studied the
 effect of home-court advantage using the 1991-1992 college basketball season.
@@
 They set up their study to find if the home teams won games by more points when
-they had home-court advantage vs. when playing at neutral court. This study
+they had home-court advantage vs.~when playing at neutral court. This study
 concluded with evidence supporting home-court advantage.
@@
 There are also surveys on the factors contributing to home-court advantage.
-@Carron1992 gave four main game location factors for home and away
+\citet{Carron1992} gave four main game location factors for home and away
 teams, namely, the crowd factor, which is the impact of fans cheering; learning
@@
 these factors would all be removed if games were played at a neutral site even
-if one team was designated as “home team”. This study was reviewed a decade later
-by @Carron2005. The 2005 review goes over the new findings from studies
+if one team was designated as ``home team''. This study was reviewed a decade later
+by \citet{Carron2005}. The 2005 review goes over the new findings from studies
 about the significance of these four game location factors. Since 1992 they found
@@
 evidence suggesting learning and rule factors impact the NBA. One interesting
-finding cited by @Carron2005 is that in the absence of crowds result
+finding cited by \citet{Carron2005} is that in the absence of crowds result
 in overall performance increases.
 
-```{r data, include=FALSE}
-#Data Collection
-game_log_2020 <- read.csv(file="../data/Game_Log_2020.csv", header=TRUE)
-game_log_2017_2019 <- read.csv(file="../data/2017-2019_game_data.csv", header=TRUE)
+\hypertarget{sec:data}{%
+\section{Data}\label{sec:data}}
 
-#Data Cleaning
-game_log_2020$W.L <- ifelse(game_log_2020$W.L=="W", 1, 0)
-game_log_2017_2019$W.L <- ifelse(game_log_2017_2019$W.L=="W", 1, 0)
-
-game_log_2020$Home.Away <- ifelse(game_log_2020$Home.Away == FALSE, "Home", "Away")
-game_log_2017_2019$Home.Away <- ifelse(game_log_2017_2019$Home.Away == FALSE, "Home", "Away")
-
-game_log_2020_home <- game_log_2020[game_log_2020$Home.Away=="Home",]
-game_log_2020_away <- game_log_2020[game_log_2020$Home.Away=="Away",]
-game_log_2017_2019_home <- game_log_2017_2019[game_log_2017_2019$Home.Away=="Home", ]
-game_log_2017_2019_away <- game_log_2017_2019[game_log_2017_2019$Home.Away=="Away", ]
-
-#combineddata
-combined <- rbind(game_log_2020, game_log_2017_2019)
-combined$Year <- as.factor(combined$Year)
-combined <-  na.omit(combined)
-
-years <- 2010:2020
-winpct <- c(.671, .667, .679, .635, .562, .593, .674, .570, .707, .561, .482)
-```
-
-# Data {#sec:data}
 Data were collected from the official NBA website. The main variables of interest
@@
 focusing on the 2017 to 2020 playoffs we can take a quick look at home team win
-percentages since 2010. Notice in Figure \@ref(fig:Fig1), the 10 years before 2020, the home team
+percentages since 2010. Notice in Figure \ref{fig:Fig1}, the 10 years before 2020, the home team
 winning percentage ranged from around 0.56 to 0.7 and never dipped below 0.5. The
@@
 
-(ref:cap-Fig1) Winning percentage of NBA home teams in the playoffs since 2010, the green line denotes .500
-```{r Fig1, echo=FALSE, fig.cap = " (ref:cap-Fig1)", fig.align='center'}
-winyear <- cbind(winpct, years)
-winyear <- as.data.frame(winyear)
-highlight <- data.frame(winpct = .482, years = 2020)
-ggplot(winyear, aes(x=years, y=winpct))+
-      geom_point()+
-      geom_point(data=highlight,
-             aes(x=years, y=winpct),
-             color='red')+
-      scale_x_continuous(name = "Year", breaks = years, labels = years)+
-      scale_y_continuous(name =  "Winning Percentage", limits = c(0.4, 0.8))+
-      geom_hline(aes(yintercept = 0.5), color = "green")+
-      geom_text(aes(label=winpct), hjust=0, vjust=-1, size=2)
-```
+\begin{figure}
+  \centering
+  \includegraphics[width=0.9\linewidth]{NBABubbleAnalysis_files/figure-latex/Fig1-1}
+  \caption{Winning percentage of NBA home teams in the playoffs since
+    2010, the green line denotes .500.}
+  \label{fig:Fig1}
+\end{figure}
 
 Moving on to the main focus of the study, comparing 2020 to 2017-2019.
-Figure \@ref(fig:Fig2) shows the histograms of the home (green) and
-away scoring (red) for 2020 vs. 2017-2019. All histograms are
+Figure \ref{fig:Fig2} shows the histograms of the home (green) and
+away scoring (red) for 2020 vs.~2017-2019. All histograms are
 fairly bell shaped, which is important for statistical tests
@@
 
-(ref:cap-Fig2)  Histograms of home (blue) and away (red) scoring
-for 2020 (bottom) and 2017-2019 (top)
-```{r Fig2, echo=FALSE, fig.cap="(ref:cap-Fig2)", fig.align='center'}
-ggplot(combined) +
-  geom_histogram(aes(x = PTS, y = ..density.., fill = Home.Away), bins = 10, binwidth = 5)+
-  facet_grid(Year ~ Home.Away)+
-  theme(legend.position = "none")
-```
+\begin{figure}
+  \centering
+  \includegraphics[width=0.9\linewidth]{NBABubbleAnalysis_files/figure-latex/Fig2-1}
+  \caption{Histograms of home (blue) and away (red) scoring
+    for 2020 (bottom) and 2017-2019 (top).}
+  \label{fig:Fig2}
+\end{figure}
 
 Our second target of inference is shooting percentage for home and away teams.
-Figure \@ref(fig:Fig3) shows home shooting for two-pointers, three-pointers and
-free throws for 2020 (top) vs. 2017-19 (bottom). The histograms appear to be fairly
-similarly distributed between 2020 and 2017-19. Likewise, Figure \@ref(fig:Fig4),
+Figure \ref{fig:Fig3} shows home shooting for two-pointers, three-pointers and
+free throws for 2020 (top) vs.~2017-19 (bottom). The histograms appear to be fairly
+similarly distributed between 2020 and 2017-19. Likewise, Figure \ref{fig:Fig4},
 shows the same percentages except for away teams. It appears that the
@@
 
-(ref:cap-Fig3) Histograms of home shooting percentages for two-
-pointers, three-pointers and free throws for 2020 (top) vs. 2017-19 (bottom)
-```{r Fig3, echo=FALSE, fig.cap="(ref:cap-Fig3)"}
-hometwo2020 <- data.frame(Year=game_log_2020_home$Year, Type=rep("Two Pointers", 83),
-                          Percentage=game_log_2020_home$X2P.)
-homethree2020 <- data.frame(Year=game_log_2020_home$Year, Type=rep("Three Pointers", 83),
-                            Percentage=game_log_2020_home$X3P.)
-homeFT2020 <- data.frame(Year=game_log_2020_home$Year, Type=rep("Free Throws", 83),
-                         Percentage=game_log_2020_home$FT.)
-hometwo2017_2019 <- data.frame(Year=game_log_2017_2019_home$Year, Type=rep("Two Pointers", 243),
-                               Percentage=game_log_2017_2019_home$X2P.)
-homethree2017_2019 <- data.frame(Year=game_log_2017_2019_home$Year, Type=rep("Three Pointers", 243),
-                                 Percentage=game_log_2017_2019_home$X3P.)
-homeFT2017_2019 <- data.frame(Year=game_log_2017_2019_home$Year, Type=rep("Free Throws", 243),
-                              Percentage=game_log_2017_2019_home$FT.)
-homegraph<-rbind(hometwo2020, homethree2020, homeFT2020, hometwo2017_2019, homethree2017_2019,
-                 homeFT2017_2019)
+\begin{figure}
+  \centering
+  \includegraphics[width=0.9\linewidth]{NBABubbleAnalysis_files/figure-latex/Fig3-1}
+  \caption{Histograms of home shooting percentages for two-
+    pointers, three-pointers and free throws for 2020 (top) vs.~2017-19
+    (bottom).}
+  \label{fig:Fig3}
+\end{figure}
 
-ggplot(homegraph) +
-  geom_histogram(aes(x = Percentage, y = ..density..), fill = "#00BFC4", bins = 10, binwidth = 5)+
-  facet_grid(Year ~ Type)
-```
+\begin{figure}
+  \centering
+  \includegraphics[width=0.9\linewidth]{NBABubbleAnalysis_files/figure-latex/Fig4-1}
+  \caption{Histograms of away shooting percentages for two-pointers,
+    three-pointers and free throws for 2020 (top) vs.~2017-19
+    (bottom).}
+  \label{fig:Fig4}  
+\end{figure}
 
-(ref:cap-Fig4) Histograms of away shooting percentages for two-
-pointers, three-pointers and free throws for 2020 (top) vs. 2017-19 (bottom)"
-```{r Fig4, echo=FALSE, fig.cap="(ref:cap-Fig4)", fig.align='center'}
-awaytwo2020 <- data.frame(Year=game_log_2020_away$Year, Type=rep("Two Pointers", 83),
-                          Percentage=game_log_2020_away$X2P.)
-awaythree2020 <- data.frame(Year=game_log_2020_away$Year, Type=rep("Three Pointers", 83),
-                            Percentage=game_log_2020_away$X3P.)
-awayFT2020 <- data.frame(Year=game_log_2020_away$Year, Type=rep("Free Throws", 83),
-                         Percentage=game_log_2020_away$FT.)
-awaytwo2017_2019 <- data.frame(Year=game_log_2017_2019_away$Year, Type=rep("Two Pointers", 243),
-                               Percentage=game_log_2017_2019_away$X2P.)
-awaythree2017_2019 <- data.frame(Year=game_log_2017_2019_away$Year, Type=rep("Three Pointers", 243),
-                                 Percentage=game_log_2017_2019_away$X3P.)
-awayFT2017_2019 <- data.frame(Year=game_log_2017_2019_away$Year, Type=rep("Free Throws", 243),
-                              Percentage=game_log_2017_2019_away$FT.)
-awaygraph<-rbind(awaytwo2020, awaythree2020, awayFT2020, awaytwo2017_2019, awaythree2017_2019,
-                 awayFT2017_2019)
+\hypertarget{sec:methods}{%
+\section{Methods}\label{sec:methods}}
 
-ggplot(awaygraph) +
-  geom_histogram(aes(x = Percentage, y = ..density..), fill = "#F8766D", bins = 10, binwidth = 5)+
-  facet_grid(Year ~ Type)
-```
-
-# Methods {#sec:methods}
 The 2020 bubble provides a new and exciting opportunity to study home-court
@@
 removed all 4 possible game location factors impacting home-court advantage
-hypothesized by @Carron1992. The NBA bubble featured 8 seeding games then a
+hypothesized by \citet{Carron1992}. The NBA bubble featured 8 seeding games then a
 standard playoff format. The focus of this study was the play during the playoff
@@
 
-1. Is the home team winning percentage in 2020 different than that it was in 2017-2019?
-1. Is the average home team scoring different in 2020 than it was over 2017-2019?
-1. Is the average away team scoring different in 2020 than it was over 2017-2019?
-1. Are home teams making two-pointers at the same rate in 2020 as 2017-2019?
-1. Are home teams making three-pointers at the same rate in 2020 as 2017-2019?
-1. Are home teams making freethrows at the same rate in 2020 as 2017-2019?
-1. Are away teams making two-pointers at the same rate in 2020 as 2017-2019?
-1. Are away teams making three-pointers at the same rate in 2020 as 2017-2019?
-1. Are away teams making free throws at the same rate in 2020 as 2017-2019?
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\item
+  Is the home team winning percentage in 2020 different than that it was in 2017-2019?
+\item
+  Is the average home team scoring different in 2020 than it was over 2017-2019?
+\item
+  Is the average away team scoring different in 2020 than it was over 2017-2019?
+\item
+  Are home teams making two-pointers at the same rate in 2020 as 2017-2019?
+\item
+  Are home teams making three-pointers at the same rate in 2020 as 2017-2019?
+\item
+  Are home teams making freethrows at the same rate in 2020 as 2017-2019?
+\item
+  Are away teams making two-pointers at the same rate in 2020 as 2017-2019?
+\item
+  Are away teams making three-pointers at the same rate in 2020 as 2017-2019?
+\item
+  Are away teams making free throws at the same rate in 2020 as 2017-2019?
+\end{enumerate}
 
 All nine questions can be approached by a standard two-sample
-comparison with the $z$-test. The $z$-test statistic follows a
+comparison with the \(z\)-test. The \(z\)-test statistic follows a
 standard normal distribution, which is a good approximation based on
@@
 We also conducted nonparametric tests that are distribution free to
-confirm the results from the $z$-test. For question~1, we used
+confirm the results from the \(z\)-test. For question\textasciitilde1, we used
 Fisher's exact test for a contigency table which summarizes the wins
@@
 
-All three tests, namely the $z$-test, Fisher's exact test, and
-Wilcoxan's rank-sum test, were performed using R [@R].
+All three tests, namely the \(z\)-test, Fisher's exact test, and
+Wilcoxan's rank-sum test, were performed using R \citep{R}.
 
-# Results {#sec:results}
-```{r, echo=FALSE}
-#Analysis
-one <- prop.test(x=c(sum(game_log_2020_home$W.L), sum(game_log_2017_2019_home$W.L)),
-                 n=c(nrow(game_log_2020_home), nrow(game_log_2017_2019_home)))
-X2020 <- table(game_log_2020_home$W.L)
-X17_19 <- table(game_log_2017_2019_home$W.L)
-homewin<- rbind(X2020, X17_19)
-onew <- fisher.test(x = homewin)
-rowone <- c(as.numeric(one$estimate[1]), as.numeric(one$estimate[2]), one$p.value, onew$p.value)
+\hypertarget{sec:results}{%
+\section{Results}\label{sec:results}}
 
-two <- z.test(x=game_log_2020_home$PTS, y=game_log_2017_2019_home$PTS, sigma.x =
-                sd(game_log_2020_home$PTS), sigma.y = sd(game_log_2017_2019_home$PTS))
-twow <- wilcox.test(x=game_log_2020_home$PTS, y=game_log_2017_2019_home$PTS)
-rowtwo <- c(as.numeric(two$estimate[1])/100, as.numeric(two$estimate[2])/100,
-            two$p.value, twow$p.value)
-
-three <- z.test(x=game_log_2020_away$PTS, y=game_log_2017_2019_away$PTS, sigma.x =
-                  sd(game_log_2020_away$PTS), sigma.y = sd(game_log_2017_2019_away$PTS))
-threew <- wilcox.test(x=game_log_2020_away$PTS, y=game_log_2017_2019_away$PTS)
-rowthree <- c(as.numeric(three$estimate[1])/100, as.numeric(three$estimate[2])/100,
-              three$p.value, threew$p.value)
-
-four <- prop.test(x=c(sum(game_log_2020_home$X2PM), sum(game_log_2017_2019_home$X2PM)),
-                  n=c(sum(game_log_2020_home$X2PA), sum(game_log_2017_2019_home$X2PA)))
-fourw <- wilcox.test(x = game_log_2020_home$X2P., y = game_log_2017_2019_home$X2P.)
-rowfour <- c(as.numeric(four$estimate[1]), as.numeric(four$estimate[2]),
-              four$p.value, fourw$p.value)
-
-five <- prop.test(x=c(sum(game_log_2020_home$X3PM), sum(game_log_2017_2019_home$X3PM)),
-                  n=c(sum(game_log_2020_home$X3PA), sum(game_log_2017_2019_home$X3PA)))
-fivew <- wilcox.test(x = game_log_2020_home$X3P., y = game_log_2017_2019_home$X3P.)
-rowfive <- c(as.numeric(five$estimate[1]), as.numeric(five$estimate[2]),
-            five$p.value, fivew$p.value)
-
-six <- prop.test(x=c(sum(game_log_2020_home$FTM), sum(game_log_2017_2019_home$FTM)),
-                 n=c(sum(game_log_2020_home$FTA), sum(game_log_2017_2019_home$FTA)))
-sixw <- wilcox.test(x = game_log_2020_home$FT., y = game_log_2017_2019_home$FT.)
-rowsix <- c(as.numeric(six$estimate[1]), as.numeric(six$estimate[2]),
-            six$p.value, sixw$p.value)
-
-seven <- prop.test(x=c(sum(game_log_2020_away$X2PM), sum(game_log_2017_2019_away$X2PM)),
-                   n=c(sum(game_log_2020_away$X2PA), sum(game_log_2017_2019_away$X2PA)))
-sevenw <- wilcox.test(x = game_log_2020_away$X2P., y = game_log_2017_2019_away$X2P.)
-rowseven <- c(as.numeric(seven$estimate[1]), as.numeric(seven$estimate[2]),
-             seven$p.value, sevenw$p.value)
-
-eight <- prop.test(x=c(sum(game_log_2020_away$X3PM), sum(game_log_2017_2019_away$X3PM)),
-                   n=c(sum(game_log_2020_away$X3PA), sum(game_log_2017_2019_away$X3PA)))
-eightw <- wilcox.test(x = game_log_2020_away$X3P., y = game_log_2017_2019_away$X3P.)
-roweight <- c(as.numeric(eight$estimate[1]), as.numeric(eight$estimate[2]),
-              eight$p.value, eightw$p.value)
-
-nine <- prop.test(x=c(sum(game_log_2020_away$FTM), sum(game_log_2017_2019_away$FTM)),
-                  n=c(sum(game_log_2020_away$FTA), sum(game_log_2017_2019_away$FTA)))
-ninew <- wilcox.test(x = game_log_2020_away$FT., y = game_log_2017_2019_away$FT.)
-rownine <- c(as.numeric(nine$estimate[1]), as.numeric(nine$estimate[2]),
-            nine$p.value, ninew$p.value)
-
-titles <- c(2020, "2017-19", "Z-test P-value", "Nonparametric P-value")
-rows <- c("Home Win ", "Home Scoring", "Away Scoring", "Home 2P", "Home 3P",
-          "Home FT", "Away 2P", "Away 3P", "Away FT")
-results <- data.frame(rowone, rowtwo, rowthree, rowfour, rowfive, rowsix,
-                    rowseven, roweight, rownine, stringsAsFactors = FALSE)
-resultsfinal <- data.frame(t(results))
-resultsfinal1 <- round(resultsfinal[1:2], 3)
-resultsfinal2 <- round(resultsfinal[3:4], 4)
-resultsfinal <-cbind(resultsfinal1, resultsfinal2)
-colnames(resultsfinal) <- titles
-row.names(resultsfinal) <- rows
-```
-
-```{r table, echo = FALSE}
-knitr::kable(resultsfinal, caption = 'The results from the 9 tests', align = "cccccc", booktabs = TRUE)
-```
+\begin{table}
+  \caption{The results from the 9 tests.}
+  \label{tab:table}
+\centering
+\begin{tabular}[t]{lcccc}
+  \toprule
+  & 2020 & 2017-19 & \multicolumn{2}{c}{P-value}\\
+  \cmidrule(lr){4-5}
+  &          &                & Z-test & Wilcoxan\\
+\midrule
+Home Win & 0.482 & 0.613 & 0.0497 & 0.0400\\
+Home Scoring & 1.101 & 1.081 & 0.2321 & 0.2985\\
+Away Scoring & 1.091 & 1.040 & 0.0008 & 0.0004\\
+Home 2P & 0.523 & 0.515 & 0.4335 & 0.5719\\
+Home 3P & 0.363 & 0.357 & 0.5733 & 0.8852\\
+Home FT & 0.793 & 0.774 & 0.0692 & 0.0496\\
+Away 2P & 0.536 & 0.504 & 0.0003 & 0.0003\\
+Away 3P & 0.357 & 0.346 & 0.3256 & 0.3081\\
+Away FT & 0.783 & 0.777 & 0.6601 & 0.8370\\
+\bottomrule
+\end{tabular}
+\end{table}
 
 Starting from the top of the
-Table \@ref(tab:table) summarizes p-values of the nine hypotheses for both
-$z$-tests and Wilcoxon tests. The p-values are all fairly similar for both
+Table \ref{tab:table} summarizes p-values of the nine hypotheses for both
+\(z\)-tests and Wilcoxon tests. The p-values are all fairly similar for both
 tests giving strong confidence in conclusions drawn. Also reported are
@@
 in home win percentage in 2020 from 2017-19, with p-value of 0.0497 for the
-$z$-test and 0.0400 for Fisher's exact test. The 95% confidence
-interval (CI)  of $(-0.255, -0.008)$ 
+\(z\)-test and 0.0400 for Fisher's exact test. The 95\% confidence
+interval (CI) of \((-0.255, -0.008)\)
 confirms our belief that home-court advantage was lost in the 2020 NBA playoffs. However, after accounting for
@@
 shooting all show no significant difference, on average, between 2020 vs.
-2017-19 based on p-values from both tests. However, the Wilcoxon test and $z$-test have conflicting results
-for free throws. The $z$-test p-value of 0.0692 indicates no significant difference, while
-the Wilcoxon test p-value of 0.0469 indicates a difference at the 5% significance level. Since,
+2017-19 based on p-values from both tests. However, the Wilcoxon test and \(z\)-test have conflicting results
+for free throws. The \(z\)-test p-value of 0.0692 indicates no significant difference, while
+the Wilcoxon test p-value of 0.0469 indicates a difference at the 5\% significance level. Since,
 the p-value of Wilcoxon test is so close to significance level and neither p-value is significant
@@
 Away teams saw more of an impact than home teams. For starters, there is a
-significant increase in mean points per game, indicated by p-value of 0.0008 for $z$-test and
+significant increase in mean points per game, indicated by p-value of 0.0008 for \(z\)-test and
 0.0004 for Wilcoxon. It is important to note both p-values also remain significant after
 a Bonferonni correction giving strong indication of significance. The average difference in points was estimated to
-be about 5 points, with 95% CI $(2.083, 7.988)$. Likewise, the away team two-point
+be about 5 points, with 95\% CI \((2.083, 7.988)\). Likewise, the away team two-point
 shooting efficiency increased significantly based on p-value of 0.0003
-for both the $z$-test and Wilcoxon test. Again, both p-values remain significant after Bonferroni correction.
-The average difference was estimated to be about 0.03, with 95% CI $(0.015, 0.050)$. However,
+for both the \(z\)-test and Wilcoxon test. Again, both p-values remain significant after Bonferroni correction.
+The average difference was estimated to be about 0.03, with 95\% CI \((0.015, 0.050)\). However,
 unlike two-point shooting, away teams did not see a statistically significant difference in three-point and free
@@
 
-# Discussion {#sec:disc}
+\hypertarget{sec:disc}{%
+\section{Discussion}\label{sec:disc}}
+
 Generally it seemed that away teams fared better in the 2020 NBA playoff bubble
@@
 understand that home teams seemed to struggle to win compared to normal conditions.
-Compared to @Kotecki who finds home teams consistently have a significantly
-better record than away teams boasting about a 60.5% win percentage in his sample,
-the 48.2% home winning percentage of 2020 home teams is quite a shift.
+Compared to \citet{Kotecki} who finds home teams consistently have a significantly
+better record than away teams boasting about a 60.5\% win percentage in his sample,
+the 48.2\% home winning percentage of 2020 home teams is quite a shift.
 In this study, home teams did not benefit from the usual advantages provided by
@@
 were not
-all significant these increases are exactly what is reported in @Carron2005
+all significant these increases are exactly what is reported in \citet{Carron2005}
 when they explain how evidence suggests that teams perform better with the absence
@@
 of home-court-advantage is the negative effects playing on the road away teams
-face. In 2020 there wasn’t any evidence of regression for home team performance,
+face. In 2020 there wasn't any evidence of regression for home team performance,
 based on the performance measures used, despite being stripped of home-court
-advantage. Yet, home teams lost about 12% more of games in the 2020 playoffs
+advantage. Yet, home teams lost about 12\% more of games in the 2020 playoffs
 than the typical average. This was because of the improvement of away teams.
@@
 increase in winning. The improvement of away teams confirms a proposition from
-@Greer that the positive social impact of crowds benefiting
+\citet{Greer} that the positive social impact of crowds benefiting
 home teams may be a result of inhibiting away teams.
@@
 At least some of that improvement from away teams came from significantly higher
-two-point efficiency. This corresponds with the conclusion from @Harris,
+two-point efficiency. This corresponds with the conclusion from \citet{Harris},
 where they found home teams are best suited to capitalize on advantages from two-point
@@
 
-# References {#sec:ref}
+\bibliographystyle{chicago}
+\bibliography{citations.bib}
+
+\end{document}

### 2020-12-02 Manuscript/NBABubbleAnalysis.Rmd
Update NBABubbleAnalysis.Rmd

@@
 highlight <- data.frame(winpct = .482, years = 2020)
+pdf("Fig1.pdf")
 ggplot(winyear, aes(x=years, y=winpct))+
@@
 ```{r Fig2, echo=FALSE, fig.cap="(ref:cap-Fig2)", fig.align='center'}
+pdf("Fig2.pdf")
 ggplot(combined) +
@@
                  homeFT2017_2019)
-
+pdf("Fig3.pdf")
 ggplot(homegraph) +
@@
                  awayFT2017_2019)
-
+pdf("Fig4.pdf")
 ggplot(awaygraph) +

### 2020-12-02 Manuscript/manuscript.tex
Update manuscript.tex

@@
   \centering
-  \includegraphics[width=0.9\linewidth]{NBABubbleAnalysis_files/figure-latex/Fig1-1}
+  \includegraphics[width=0.9\linewidth]{Fig1}
   \caption{Winning percentage of NBA home teams in the playoffs since
@@
   \centering
-  \includegraphics[width=0.9\linewidth]{NBABubbleAnalysis_files/figure-latex/Fig2-1}
+  \includegraphics[width=0.9\linewidth]{Fig2}
   \caption{Histograms of home (blue) and away (red) scoring
@@
   \centering
-  \includegraphics[width=0.9\linewidth]{NBABubbleAnalysis_files/figure-latex/Fig3-1}
+  \includegraphics[width=0.9\linewidth]{Fig3}
   \caption{Histograms of home shooting percentages for two-
@@
   \centering
-  \includegraphics[width=0.9\linewidth]{NBABubbleAnalysis_files/figure-latex/Fig4-1}
+  \includegraphics[width=0.9\linewidth]{Fig4}
   \caption{Histograms of away shooting percentages for two-pointers,

### 2021-03-23 Manuscript/NBABubbleAnalysis.Rmd
Update NBABubbleAnalysis.Rmd

@@
 
+# Appendix {#sec:Appendix}
+```{r, echo=FALSE}
+permutation_test <- function(data1, data2, test_column_num, n, R) {
+  #calculate test statsitic and store it
+  z <- z.test(x=data1[ ,test_column_num], y=data2[ ,test_column_num], sigma.x =
+                sd(data1[ ,test_column_num]), sigma.y = sd(data2[ ,test_column_num]))
+  z <- abs(z$statistic)
+
+  #calculate mean of combined samples
+  overallmean<-c(data1[ ,test_column_num], data2[ ,test_column_num])
+  overallmean<-mean(overallmean)
+
+  #transform datasets using xi*=xi-x̄-z̄
+  data1transformed <- data1[ ,test_column_num] - mean(data1[ ,test_column_num]) + overallmean
+  data1transformed <- data.frame(data1transformed)
+  data2transformed <- data2[ ,test_column_num] - mean(data2[ ,test_column_num]) + overallmean
+  data2transformed <- data.frame(data2transformed)
+
+  #collect R samples of n data points from each group and calculate test statistic between each pair of samples
+  bootstrap <- data.frame("Bootsrap T-stat"=numeric(), "Condition"=numeric())
+  set.seed(1234)
+  for (i in 1:R){
+    sample1 <- data1transformed[sample(1:nrow(data1transformed), nrow(data1transformed), replace=TRUE),]
+    sample2 <- data2transformed[sample(1:nrow(data2transformed), nrow(data2transformed), replace=TRUE),]
+
+    bt <- z.test(x=sample1, y=sample2, sigma.x =
+                  sd(sample1), sigma.y = sd(sample2))
+    bts <- abs(bt$statistic)
+
+    bootstrap[i, ] <- bts
+    #determine if bootstrap statistics is larger than observed and save result
+    if(bootstrap[i,1] >= z) {bootstrap[i,2]=1} else {bootstrap[i,2]=0}
+  }
+  #calculate proportion of bootsrap t-statistics wihch are larger than or equal to observed test statistic
+  p_value <- sum(bootstrap[, 2])/nrow(bootstrap)
+  return(p_value)
+}
+
+row_one <- permutation_test(game_log_2020_home, game_log_2017_2019_home, 6, 100, 10000)
+row_two <-permutation_test(game_log_2020_home, game_log_2017_2019_home, 8, 100, 10000)
+row_three <- permutation_test(game_log_2020_away, game_log_2017_2019_away, 8, 100, 10000)
+row_four <- permutation_test(game_log_2020_home, game_log_2017_2019_home, 14, 100, 10000)
+row_five <- permutation_test(game_log_2020_home, game_log_2017_2019_home, 17, 100, 10000)
+row_six <- permutation_test(game_log_2020_home, game_log_2017_2019_home, 20, 100, 10000)
+row_seven <-  permutation_test(game_log_2020_away, game_log_2017_2019_away, 14, 100, 10000)
+row_eight <- permutation_test(game_log_2020_away, game_log_2017_2019_away, 17, 100, 10000)
+row_nine <- permutation_test(game_log_2020_away, game_log_2017_2019_away, 20, 100, 10000)
+perm_test_res <- c(row_one, row_two, row_three, row_four, row_five, row_six, row_seven, row_eight, row_nine)
+perm_test_col <- matrix(data=perm_test_res, ncol = 1)
+app_resultsfinal <-cbind(resultsfinal, perm_test_col)
+colnames(app_resultsfinal)[5] <- "Bootstrap Z-test"
+```
+
+In an effort to advance and strengthen the paper we conducted bootstrap analysis for every
+hypothesis. A custom R function was developed and utilized to conduct this analysis, the code for 
+that function can be found below:
+
+```{r, echo = TRUE, eval=FALSE}
+permutation_test <- function(data1, data2, test_column_num, n, R) {
+  #calculate test statsitic and store it
+  z <- z.test(x=data1[ ,test_column_num], y=data2[ ,test_column_num], sigma.x =
+                sd(data1[ ,test_column_num]), sigma.y = sd(data2[ ,test_column_num]))
+  z <- abs(z$statistic)
+
+  #calculate mean of combined samples
+  overallmean<-c(data1[ ,test_column_num], data2[ ,test_column_num])
+  overallmean<-mean(overallmean)
+
+  #transform datasets using xi*=xi-x̄-z̄
+  data1transformed <- data1[ ,test_column_num] - mean(data1[ ,test_column_num]) + overallmean
+  data1transformed <- data.frame(data1transformed)
+  data2transformed <- data2[ ,test_column_num] - mean(data2[ ,test_column_num]) + overallmean
+  data2transformed <- data.frame(data2transformed)
+
+  #collect R samples of n data points from each group and calculate test statistic between each pair of samples
+  bootstrap <- data.frame("Bootsrap T-stat"=numeric(), "Condition"=numeric())
+  set.seed(1234)
+  for (i in 1:R){
+    sample1 <- data1transformed[sample(1:nrow(data1transformed), nrow(data1transformed), replace=TRUE),]
+    sample2 <- data2transformed[sample(1:nrow(data2transformed), nrow(data2transformed), replace=TRUE),]
+
+    bt <- z.test(x=sample1, y=sample2, sigma.x =
+                  sd(sample1), sigma.y = sd(sample2))
+    bts <- abs(bt$statistic)
+
+    bootstrap[i, ] <- bts
+    #determine if bootstrap statistics is larger than observed and save result
+    if(bootstrap[i,1] >= z) {bootstrap[i,2]=1} else {bootstrap[i,2]=0}
+  }
+  #calculate proportion of bootsrap t-statistics wihch are larger than or equal to observed test statistic
+  p_value <- sum(bootstrap[, 2])/nrow(bootstrap)
+  return(p_value)
+}
+```
+
+The results from the bootstrap tests can be compared to the previous tests results 
+in order to strengthen understanding. These comparisons can be made using the table, Table 
+\@ref(tab:table2), below.
+
+```{r table2, echo = FALSE}
+knitr::kable(app_resultsfinal, caption = 'Previous Results Side By Side With Bootstrap Results', align = "cccccc", booktabs = TRUE)
+```
+
+As you can see in the table, the p-values for the tests using bootstrapping
+align closely with p-values from the other tests. Starting with the home team
+win percentage, we see our p-value is 0.0405. This is important because 
+earlier in our paper we found that the p-values for z-test(.0497) and Fisher’s
+exact test(.0400) were initially significant, but not after a Bonferroni 
+correction. Thus, the bootstrapping result gives some extra level of confidence
+to the conclusion that the home team winning percentage was different from
+years past.
+
+For home teams, the bootstrapping tests seem to confirm the conclusion
+playing in the NBA bubble had very little impact on performance. The bootstrap
+tests for home scoring, two-point percentage and three-point percentage all
+have very high p-values and agree with previous tests that there is no evidence
+of a difference in these measures. One interesting finding is for the home team
+free throw percentage. Earlier, we found the z-test (p=0.0692) was insignificant,
+but the Wilcoxon test(p=0.0496) was significant. The bootstrapping test(p=0.551)
+agrees with the z-test that there is no significant change in home team free throw
+percentage. Since two of three tests run are insignificant and the third is very
+close, it is probably safest to conservatively say that there is not much evidence
+to support a claim that home team free throw percentage changed.
+
+For away teams, we saw the bootstrapping test confirm the results of the previous
+tests in every case. The bootstrapping methods corroborate the insignificant 
+results for away three-point (p=0.3494) and away free throw shooting (p=0.4743).
+Likewise, away team scoring(p=.0012) and two-point percentage(p=.0001) still had
+highly significant p-values for the bootstrap test, just like the other tests for
+these two measures. This gives us great confidence that away scoring and two-point
+percentage changed in the bubble. If you recall, the confidence intervals for 
+these measures support the claim that they increased. This is important because
+it suggests when home-court advantage is removed, away teams were able to score 
+more points by increased two-point efficiency, suggesting home-court advantage 
+stems from an adverse effect on away teams. Recall that @Harris found that 
+home-court advantage was maximized when home teams capitalized on their
+advantage shooting the two-pointer. When away teams no longer face more 
+challenging conditions for two-point shooting while going on the road, home
+teams lose their advantage.
+
+Overall, the addition of the bootstrapping methods to the study
+appear to strengthen the conclusions made in the paper.
+
 # References {#sec:ref}

### 2021-04-07 Manuscript/NBABubbleAnalysis.Rmd
Update NBABubbleAnalysis.Rmd

@@
 ```{r, echo = TRUE, eval=FALSE}
-permutation_test <- function(data1, data2, test_column_num, n, R) {
+permutation_test <- function(data1, data2, test_column_num, R) {
   #calculate test statsitic and store it

### 2021-04-16 Manuscript/NBABubbleAnalysis.Rmd
Update NBABubbleAnalysis.Rmd

@@
 
-```{r, echo = TRUE, eval=FALSE}
+```{r, echo = TRUE, eval=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=50)}
 permutation_test <- function(data1, data2, test_column_num, R) {

### 2022-01-26 Manuscript/NBABubbleAnalysis.Rmd
Update NBABubbleAnalysis.Rmd
Added reference lines for means on histograms

@@
 ```{r Fig2, echo=FALSE, fig.cap="(ref:cap-Fig2)", fig.align='center'}
+ref_line2 <- aggregate(data=combined, PTS ~ Year + Home.Away, FUN=mean)
+
 pdf("Fig2.pdf")
@@
   facet_grid(Year ~ Home.Away)+
-  theme(legend.position = "none")
+  theme(legend.position = "none")+
+  geom_vline(data=ref_line2, aes(xintercept=PTS))+
+  geom_text(data=ref_line2, aes(x=PTS+3.5, label="mean", y=0.045))
 ```
@@
                  homeFT2017_2019)
+ref_line3 <- aggregate(data=homegraph, Percentage ~ Year + Type, FUN=mean)
 pdf("Fig3.pdf")
@@
   geom_histogram(aes(x = Percentage, y = ..density..), fill = "#00BFC4", bins = 10, binwidth = 5)+
-  facet_grid(Year ~ Type)
+  facet_grid(Year ~ Type)+
+  geom_vline(data=ref_line3, aes(xintercept=Percentage))+
+  geom_text(data=ref_line3, aes(x=Percentage+6, label="mean", y=0.075))
 ```
@@
                  awayFT2017_2019)
+ref_line4 <- aggregate(data=awaygraph, Percentage ~ Year + Type, FUN=mean)
 pdf("Fig4.pdf")
@@
   geom_histogram(aes(x = Percentage, y = ..density..), fill = "#F8766D", bins = 10, binwidth = 5)+
-  facet_grid(Year ~ Type)
+  facet_grid(Year ~ Type)+
+  geom_vline(data=ref_line4, aes(xintercept=Percentage))+
+  geom_text(data=ref_line4, aes(x=Percentage+6, label="mean", y=0.075))
 ```

### 2022-01-26 Manuscript/NBABubbleAnalysis.Rmd
Update NBABubbleAnalysis.Rmd
Calculations for adjusted p-values

@@
 
-titles <- c(2020, "2017-19", "Z-test P-value", "Nonparametric P-value")
+titles <- c(2020, "2017-19", "Z-test P-value", "Nonparametric P-value", "Adj P-value Z", "Adj P-value W")
 rows <- c("Home Win ", "Home Scoring", "Away Scoring", "Home 2P", "Home 3P",
@@
 resultsfinal2 <- round(resultsfinal[3:4], 4)
-resultsfinal <-cbind(resultsfinal1, resultsfinal2)
+resultsfinal3 <- round(resultsfinal[3:4]*9, 4)
+resultsfinal <-cbind(resultsfinal1, resultsfinal2, resultsfinal3)
 colnames(resultsfinal) <- titles

### 2022-01-26 Manuscript/manuscript.tex
Update manuscript.tex
Updated table with adjusted p-values(still need to decide whether or not to remove 1 of parametric or non-parametric p-values like reviewer 1 suggested)

@@
 \centering
-\begin{tabular}[t]{lcccc}
-  \toprule
-  & 2020 & 2017-19 & \multicolumn{2}{c}{P-value}\\
-  \cmidrule(lr){4-5}
-  &          &                & Z-test & Wilcoxan\\
-\midrule
-Home Win & 0.482 & 0.613 & 0.0497 & 0.0400\\
-Home Scoring & 1.101 & 1.081 & 0.2321 & 0.2985\\
-Away Scoring & 1.091 & 1.040 & 0.0008 & 0.0004\\
-Home 2P & 0.523 & 0.515 & 0.4335 & 0.5719\\
-Home 3P & 0.363 & 0.357 & 0.5733 & 0.8852\\
-Home FT & 0.793 & 0.774 & 0.0692 & 0.0496\\
-Away 2P & 0.536 & 0.504 & 0.0003 & 0.0003\\
-Away 3P & 0.357 & 0.346 & 0.3256 & 0.3081\\
-Away FT & 0.783 & 0.777 & 0.6601 & 0.8370\\
-\bottomrule
+\begin{tabular}[t]{lcccccc}
+  & 2020 & 2017-19 & \multicolumn{2}{c}{P-value} & \multicolumn{2}{c}{Adjusted P-value}\\
+  &          &                & Z-test & Wilcoxan & Z-test & Wilcoxan\\
+Home Win & 0.482 & 0.613 & 0.0497* & 0.0400* & .4473 & 0.3598\\
+Home Scoring & 1.101 & 1.081 & 0.2321 & 0.2985 & 1 & 1\\
+Away Scoring & 1.091 & 1.040 & 0.0008* & 0.0004* & 0.0075* & 0.0035*\\
+Home 2P & 0.523 & 0.515 & 0.4335 & 0.5719 & 1 & 1\\
+Home 3P & 0.363 & 0.357 & 0.5733 & 0.8852 & 1 & 1\\
+Home FT & 0.793 & 0.774 & 0.0692 & 0.0496* & 0.6228 & 0.4464\\
+Away 2P & 0.536 & 0.504 & 0.0003* & 0.0003* & 0.0030 & 0.0023\\
+Away 3P & 0.357 & 0.346 & 0.3256 & 0.3081 & 1 & 1\\
+Away FT & 0.783 & 0.777 & 0.6601 & 0.8370 & 1 & 1\\
 \end{tabular}

### 2022-01-29 Manuscript/NBABubbleAnalysis.Rmd
Update NBABubbleAnalysis.Rmd
4 minor comments from reviewer 1 are addressed

@@
 home-court helps teams play better. @Cao studied the effects
-of pressure on performance in the NBA. Using throws as their measure of
+of pressure on performance in the NBA. Using free throws as their measure of
 interest, they tested whether home
 fans could distract and put pressure on opposing players to make free
-throws. However, they found insignificant
+throws. However, they did not find significant evidence
 evidence that home status has a substantial impact on missing from the free throw
@@
 that home teams should try to shoot more two point shots and force
-their opponent to take more two point shot attempts. This strategy will maximize
-the benefits of home-court advantage and give them the best chance to win.
+their opponent to take more two point shot attempts. This minimum variance strategy will 
+help home teams have a greater control of the game play and help maximize the benefits 
+home court advantange.
 
@@
 that results on these four factors are mixed. However, there is some evidence
-supporting crowd and travel factors impact games in the NBA. There is less
-evidence suggesting learning and rule factors impact the NBA. One interesting
-finding cited by @Carron2005 is that in the absence of crowds result
-in overall performance increases.
+supporting crowd and travel factors impact games across all major sports. There is less
+evidence suggesting learning and rule factors impact across the various collegiate 
+and professional sports, regardless, the NBA is not a league which has rules that
+may favor the home team like batting last in the MLB. One interesting finding cited by @Carron2005 is that in the absence of crowds result in overall performance increases.
 

### 2022-01-29 Manuscript/NBABubbleAnalysis.Rmd
Update NBABubbleAnalysis.Rmd
Addressing more concerns from reviewer 1

@@
 difference between the 2020 and 2017-2019 for home scoring, but for
-away scoring, a noticeable shift to the right in 2020 is observed
+away scoring, a more pronounced shift to the right in 2020 is observed
 compared with that in 2017-2019.
@@
 ```
+Note: Home and Away scoring totals are divided by 100
 

### 2022-01-29 Manuscript/NBABubbleAnalysis.Rmd
Update NBABubbleAnalysis.Rmd
Addressing reviewer 1 question on why 2 pointers were singled out in data by also elaborating on why all 3 shots are important to examine individually.

@@
 free throws. These variables were very popular and frequently used in the related
-literature discussed earlier. The data was collected on a game by game basis, this gave us two observations
+literature discussed earlier. While many other measures could be used for examining
+the outcome of the game and team performance, scoring seemed to be the most
+important. The winner of a game is determined by who scores more points. The three types of shots are anatural discussion point in basketball, as they're all
+important and directly impact scoring totals. Free throws can often determine the winner of close
+games, as losing teams are forced to foul the winning team to
+stop the clock and hope for some missed free throws. The three
+point shot has grown very prominent in basketball as teams have
+increased their volume of three point shots in response to the
+recent success of the Golden State Warriors and Stephen Curry. Due to this fact, is a growing consensus the three point shot is most important in basketball.
+However, two point shots are not to be overlooked and may
+actually be most important, especially for home teams, according
+to @Harris. In fact 2020 NBA champion Los Angeles Lakers actually
+led league in two point percentage throughout the playoffs and were bottom five in three point percentage
+
+The data was collected on a game by game basis, this gave us two observations
 for each variable per game played, one observation for each team(home and away).

### 2022-01-29 Manuscript/NBABubbleAnalysis.Rmd
Update NBABubbleAnalysis.Rmd
grammar fix

--- 2022-01-29 Manuscript/NBABubbleAnalysis.Rmd
@@
 free throws. These variables were very popular and frequently used in the related
-literature discussed earlier. The data was collected on a game by game basis, this gave us two observations
+literature discussed earlier. While many other measures could be used for examining
+the outcome of the game and team performance, scoring seemed to be the most
+important. The winner of a game is determined by who scores more points. The three types of shots are anatural discussion point in basketball, as they're all
+important and directly impact scoring totals. Free throws can often determine the winner of close
+games, as losing teams are forced to foul the winning team to
+stop the clock and hope for some missed free throws. The three
+point shot has grown very prominent in basketball as the entire NBA has increased their volume of three point shot attempts in response to the
+recent success of the Golden State Warriors and Stephen Curry. Due to this fact, there is a growing consensus the three point shot is most important in basketball. 
+However, two point shots are not to be overlooked and may
+actually be most important, especially for home teams, according
+to \citet{Harris}. In fact, 2020 NBA champion Los Angeles Lakers actually led league in two point percentage throughout the playoffs and were bottom five in three point percentage.
+
+The data was collected on a game by game basis, this gave us two observations
 for each variable per game played, one observation for each team(home and away).

### 2022-01-29 Manuscript/NBABubbleAnalysis.Rmd
Update NBABubbleAnalysis.Rmd
Reviewer 2 suggestion to discuss home court advantage factors in introduction

@@
 brought a great deal of concern for fans, teams, journalists, and others.
-For example, @Aschburner discusses about anticipated effects, sharing
+Among the biggest is the effect from the crowd, a home crowd has the ability
+to energize home teams, help control momentum in favor of the home team and 
+create a chaotic and stressful environment for away teams to play in. There 
+is no way replicate these effects in the bubble conditions. Also, there are 
+geographic factors that are lost like altitude and time-zone effects. There's 
+evidence teams across all sports in high altitude regions can rely on away
+teams coming in and struggling with lower oxygen levels at a higher altitude, 
+which the home teams are accustomed to playing with \citet{Lopez}. Examples,
+of high altitude teams in the NBA would be the Utah Jazz and Denver Nuggets, 
+both of whom participated in the 2020 NBA bubble. Also, referees are known to
+be subject to favoring home teams in large part due pressure from fans, which
+is no longer in play in the bubble. Refs may still have biases, but there's
+nothing to sway that bias towards the home team like normal. Any positive 
+effects of the home team playing in the arena they're most comfortable in 
+and living in the comfort of their own home are also lost. Travel is also 
+normally discussed as a factor hurting away teams, but for the context of
+the playoffs in the bubble losing this may not be significant. In the
+playoffs, many times (aside from the first game of the series when the 
+away team travels) both teams often travel at the same time since they're 
+both going back and forth between the two cities to complete the series.
+
+ @Aschburner discusses the anticipated effects, sharing
 concerns from former players, coaches and other experts about

### 2022-01-30 Manuscript/NBABubbleAnalysis.Rmd
Update NBABubbleAnalysis.Rmd
Discussion of significance of 3% change in made 2's

@@
 and professional sports, regardless, the NBA is not a league which has rules that
-may favor the home team like batting last in the MLB. One interesting finding cited by @Carron2005 is that in the absence of crowds result in overall performance increases.
+may favor the home team like batting last in the MLB. One interesting finding cited
+by @Carron2005 is that in the absence of crowds result in overall performance increases.
+Lastly, as discussed in the introduction @Lopez find evidence that geographic
+factors like altitude may influence and strengthen home court advantage for teams in
+high altitude regions.
 
@@
 the outcome of the game and team performance, scoring seemed to be the most
-important. The winner of a game is determined by who scores more points. The three types of shots are anatural discussion point in basketball, as they're all
+important. The winner of a game is determined by who scores more points. The three 
+types of shots are anatural discussion point in basketball, as they're all
 important and directly impact scoring totals. Free throws can often determine the winner of close
@@
 stop the clock and hope for some missed free throws. The three
-point shot has grown very prominent in basketball as the entire NBA has increased their volume of three point shot attempts in response to the
-recent success of the Golden State Warriors and Stephen Curry. Due to this fact, there is a growing consensus the three point shot is most important in basketball. 
+point shot has grown very prominent in basketball as the entire NBA has increased
+their volume of three point shot attempts in response to the
+recent success of the Golden State Warriors and Stephen Curry. Due to this fact, 
+there is a growing consensus the three point shot is most important in basketball. 
 However, two point shots are not to be overlooked and may
 actually be most important, especially for home teams, according
-to \citet{Harris}. In fact, 2020 NBA champion Los Angeles Lakers actually led league in two point percentage throughout the playoffs and were bottom five in three point percentage.
+to \citet{Harris}. In fact, 2020 NBA champion Los Angeles Lakers actually led league
+in two point percentage throughout the playoffs and were bottom five in three point percentage.
 
@@
 
+To get a bit more detailed, we saw an estimated 3\% increase in away two point
+shooting percentage. Likewise, NBA teams shot
+an average of about 53 two point shots per game, in our data, leading to a three 
+point increase in total scoring. Even this
+small swing in scoring for away teams can make a big difference in outcome of
+games. Around 1/3 of NBA games end by a decision of six points or less, those
+are the games where the three points matter most. Suppose there are 30 seconds
+left and the away team is losing. If the away team has the ball, only down three, 
+they can take their time finding a good shot which can
+still either be a two or three, given the score and remaining time. In the context 
+of the NBA bubble, this crucial possession also has the benefit of the away team
+getting to work in silence without the jeers from enemy fans. However, in the
+same situation while being down six the offense must rush to get a quick shot
+off, which probably has to be a three pointer. This is a much more desperate
+circumstance, less likely to have a positive outcome and even in the positive
+outcome case you're still going to be down three and likely have to foul the
+opposing team after you score. 
+Alternatively, if the away team is playing defense with 30 seconds left, only down 
+three, you can play regular defense without fouling, get a stop, then draw up a play 
+to score a three pointer with the remaining six seconds. However, in the same scenario 
+down six, you have no choice, but to foul. The free throws will likely put you down
+seven or eight points, but the alternative is let the opponent run the 24 second shot 
+clock down wasting time. Also, the away team may even be in a situation where they're
+leading by the three additional points, which for the same logic as explained above 
+puts them in far greater control of the game. Clearly, the three point swing can 
+drastically change probability of winning for a team down the stretch of a game. 
+Since, home teams picked up an additional two points per game on average vs. five 
+for away teams, the three additional points from two point shooting appear to be the 
+driving factor of away teams closing the gap with home teams.
+
 An interesting finding is all shooting and scoring numbers for both home and

### 2022-01-30 Manuscript/NBABubbleAnalysis.Rmd
Update NBABubbleAnalysis.Rmd
Discussion on why away teams more affected than home teams

@@
 
+At least some of that improvement from away teams came from significantly higher
+two-point efficiency. This corresponds with the conclusion from \citet{Harris},
+where they found home teams are best suited to capitalize on advantages from two-point
+shots. Normally, by shooting more two-pointers themselves and forcing away teams to
+shoot more two-pointers the home team benefits from increasing effects of
+home-court advantage. However, with away teams significantly improving two-point
+shooting in the bubble this strategy was no longer viable and home-court
+advantage disappeared.
+
 To get a bit more detailed, we saw an estimated 3\% increase in away two point
@@
 
-At least some of that improvement from away teams came from significantly higher
-two-point efficiency. This corresponds with the conclusion from @Harris,
-where they found home teams are best suited to capitalize on advantages from two-point
-shots. Normally, by shooting more two-pointers themselves and forcing away teams to
-shoot more two-pointers the home team benefits from increasing effects of
-home-court advantage. However, with away teams significantly improving two-point
-shooting in the bubble this strategy was no longer viable and home-court
-advantage disappeared.
+It's worth noting that much of this paper discussed the the effects from an
+offensive point of view, focusing on a teams ability to score. However, one 
+of the main factors that influence an offense's ability to score is the opposing
+defense. It's possible that home court advantage mainly functions as an extra defender
+for home teams, which is how it negatively impacts away scoring. There is not much
+home fans can do to help their team on offense, except being quiet so players can 
+easily communicate and focus. However, when on defense fans can support their team
+by cheering loudly and making it harder on the offense. This would explain how away 
+teams were able to improve offensively in the bubble. Home teams did not have the 
+regular help from the noise of fans that hinder the away team's offensive efforts. 
+This would also explain why home team offense didn't show significant change in the 
+bubble, the in-game offensive environment was largely unaltered for home teams
+since the bubble provided the same quiet environment they're used to at home while
+on offense. Meanwhile, removing the "sixth defender" from home team defensive efforts 
+allowed away teams to work more effectively on offense.
+
+An interesting finding is all shooting and scoring numbers for both home and
+away teams did make at least small increases. Although these increases
+were not all significant these increases are exactly what is reported in
+@Carron2005 when they explain how evidence suggests that teams perform
+better with the absence of fans. This is important because it coincides with our
+conclusion that home-court advantage mostly plays into games by negatively impacting away
+teams by acting as an extra defender. If fans cause overall performance to drop, then home 
+court advantage must come from a bigger drop in away performance than the drop in home
+performance. This is why away teams were able to close the gap with home teams with
+home-court advantage removed.
 

### 2022-01-30 Manuscript/NBABubbleAnalysis.Rmd
Update NBABubbleAnalysis.Rmd

@@
 of the main factors that influence an offense's ability to score is the opposing
-defense. It's possible that home court advantage mainly functions as an extra defender
-for home teams, which is how it negatively impacts away scoring. There is not much
-home fans can do to help their team on offense, except being quiet so players can 
-easily communicate and focus. However, when on defense fans can support their team
-by cheering loudly and making it harder on the offense. This would explain how away 
-teams were able to improve offensively in the bubble. Home teams did not have the 
-regular help from the noise of fans that hinder the away team's offensive efforts. 
-This would also explain why home team offense didn't show significant change in the 
-bubble, the in-game offensive environment was largely unaltered for home teams
-since the bubble provided the same quiet environment they're used to at home while
-on offense. Meanwhile, removing the "sixth defender" from home team defensive efforts 
-allowed away teams to work more effectively on offense.
+defense. It's possible that home court advantage mainly functions as an extra
+defender for home teams, which is how it negatively impacts away scoring. There
+is not much home fans can do to help their team on offense, except being quiet
+so players can easily communicate and focus. However, when on defense, fans can
+act as a "sixth defender" supporting their team by cheering loudly and making it
+harder on the offense. This would explain how away teams were able to improve
+offensively in the bubble. Home teams did not have the regular help from the
+noise of fans that hinder the away team's offensive efforts. This would also
+explain why home team offense didn't show significant change in the bubble, the
+in-game offensive environment was largely unaltered for home teams since the
+bubble provided the same quiet environment they're used to at home while on
+offense. I still believe home court advantage is better categorized as a
+negative effect on away team offense, since it's more likely
+noise from fans lowers level of play of away offense rather than raise the level
+of play of home defense. This will be further discussed below.
 

### 2022-02-01 Manuscript/NBABubbleAnalysis.Rmd
Update NBABubbleAnalysis.Rmd
Explaining why reviewer 1 suggestion of testing individual teams home court advantage was not possible given the available data size

@@
 same measures with a different sample. Otherwise, that type of study
-could help strengthen the conclusion in this paper.
+could help strengthen the conclusion in this paper. Also, you could
+answer more questions with a larger sample size. For example,  testing
+strength of home court advantage relative to specific teams, like the
+Denver Nuggets and Utah Jazz who we know from other literature may
+have a stronger home court advantage. The sample size of home games
+for the individual teams in the bubble was far too small to try and
+address this question, the Jazz and Nuggets only had 3 and 9 home
+games, respectively.
 

### 2022-02-01 Manuscript/NBABubbleAnalysis.Rmd
Update NBABubbleAnalysis.Rmd
Addressing reviewer 2 comment that we need to justify not using regression

@@
 
+Regression was also considered, but it did not seem to add any additional useful
+insight to answering the questions presented. Regression is a popular
+tool in home court advantage studies, but as mentioned previously, this
+study is very different from past studies. Rather than trying to prove the
+existence of home court advantage, like many previous studies, this paper
+was more interested in generating a clearer understanding of how home court
+advantage effects each team. The most effective way to do that is hypothesis
+testing to compare performance in the neutral bubble to previous years with
+normal game conditions for both the home and away team. Another potential
+benefit of regression would have been the ability to control for factors
+like strength of opponent. However, we felt that strength of team was
+already well enough controlled for by the fact that we focused on the
+playoffs. The playoffs only include the strongest 16 of 30 teams, then
+continue to remove the less competitive teams, so that the talent disparity
+between teams is much smaller than regular season play where controlling for
+opponent strength is likely more important.
+
 # Results {#sec:results}

### 2022-02-02 Manuscript/NBABubbleAnalysis.Rmd
Update NBABubbleAnalysis.Rmd
Generally proofreading, grammar and flow check

@@
 Abstract:
-The 2020 NBA playoffs were played inside of a bubble in Disney World because of
+The 2020 NBA playoffs were played inside of a bubble at Disney World because of
 the COVID-19 pandemic. This meant that there were no fans in attendance,
-games played on neutral courts and no traveling for teams, which in theory
-removes home-court advantage from the games. This setting has attracted much
-discussion as analysts and fans debated the possible effects it may have on the
+games were played on neutral courts and no traveling for teams. In theory,
+this should remove home-court advantage from the games. This setting genereated
+discussion and concerns, as analysts and fans debated the possible effects it may have on the
 outcome of games. Home-court advantage has historically played an influential
-role in NBA playoff series outcomes. The 2020 playoff provided a unique
+role in NBA playoff series outcomes. The 2020 playoffs provided a unique
 opportunity to study the effects of the bubble and home-court
@@
 While many factors contribute to the outcome of games, points scored
-is the deciding factor of who wins games, so scoring is the primary focus
+is the deciding factor of who wins. Thus, scoring is the primary focus
 of this study. The specific measures of interest are team scoring totals and
@@
 factor to the outcome of games. It is well-known that the home team typically
-benefits from some competitive edge from playing at their home-court, resulting
+benefits from some competitive edge while playing at their home-court, resulting
 in a better chance of winning. Thus, the NBA playing the 2020 playoffs in a
 bubble due to the COVID-19 pandemic
-brought a great deal of concern for fans, teams, journalists, and others.
-Among the biggest is the effect from the crowd, a home crowd has the ability
+brought a great deal of concern for fans, teams, journalists, and others. The bubble environment
+would not be able to replicate many of the factors that normal home-court advantage relies on.
+Among the biggest is the effect from the crowd. The home crowd has the ability
 to energize home teams, help control momentum in favor of the home team and 
@@
 is no way replicate these effects in the bubble conditions. Also, there are 
-geographic factors that are lost like altitude and time-zone effects. There's 
-evidence teams across all sports in high altitude regions can rely on away
+geographic factors that are lost, like altitude and time-zone effects. There's 
+evidence across all sports that teams in high altitude regions can rely on away
 teams coming in and struggling with lower oxygen levels at a higher altitude, 
@@
 both of whom participated in the 2020 NBA bubble. Also, referees are known to
-be subject to favoring home teams in large part due pressure from fans, which
+be subject to favoring home teams. This is due in large part to pressure from fans, which
 is no longer in play in the bubble. Refs may still have biases, but there's
@@
 normally discussed as a factor hurting away teams, but for the context of
-the playoffs in the bubble losing this may not be significant. In the
+the playoffs in the bubble, losing this may not be significant. In the
 playoffs, many times (aside from the first game of the series when the 
@@
 Our study is quite different from earlier NBA home-court advantage studies.
-By using the neutral site games of 2020 we will get to compare home and away
+By using the neutral site games of 2020, we will get to compare home and away
 performance to a control. Typically, studies just compare home vs away
@@
 # Literature Review {#sec:litrev}
-There is a voluminous literature on the effects of home-court advantage.
+There is voluminous literature on the effects of home-court advantage.
 Many NBA home-court advantage studies analyze the effects by studying shooting
-percentages. @Kotecki reported significant home-court advantage using performance-based
-statistics, specifically field goal percentage, free throw percentage, and
-points by comparing home performance vs. away performance in games.
-All of which he showed significantly indicates that
-home-court helps teams play better. @Cao studied the effects
+percentages. @Kotecki reported significant evidence of home-court advantage by 
+comparing field goal percentage, free throw percentage
+and points scored in home vs. away teams. He found all of these measures showed evidence that
+home-court advantage helps home teams play better. @Cao studied the effects
 of pressure on performance in the NBA. Using free throws as their measure of
@@
 evidence that home status has a substantial impact on missing from the free throw
-line. @Harris used two point shots, three-point shots and free throws as measures
+line. @Harris used two-point shots, three-point shots and free throws as measures
 of interest to study home-court advantage. Two point shots were found to be
@@
 that home teams should try to shoot more two point shots and force
-their opponent to take more two point shot attempts. This minimum variance strategy will 
+their opponent to take more two point shot attempts. This will 
 help home teams have a greater control of the game play and help maximize the benefits 
@@
 
-Some studies focus less on shooting and more on scoring differences and other metrics.
+Some studies focus less on shooting and more on other metrics.
 For example, @Greer focused on the influence of spectator booing on
@@
 evidence suggesting learning and rule factors impact across the various collegiate 
-and professional sports, regardless, the NBA is not a league which has rules that
-may favor the home team like batting last in the MLB. One interesting finding cited
+and professional sports. The NBA is not a league which has rules that
+may favor the home team, like batting last in the MLB, but these rule factors 
+also account for referee bias which may impact the NBA. One interesting finding cited
 by @Carron2005 is that in the absence of crowds result in overall performance increases.
@@
 the outcome of the game and team performance, scoring seemed to be the most
-important. The winner of a game is determined by who scores more points. The three 
-types of shots are anatural discussion point in basketball, as they're all
+important because the winner of a game is determined by who scores more points. Furthermore, the three 
+types of shots are a natural discussion point in basketball, as they're all
 important and directly impact scoring totals. Free throws can often determine the winner of close
@@
 stop the clock and hope for some missed free throws. The three
-point shot has grown very prominent in basketball as the entire NBA has increased
+point shot has grown very prominent in basketball. The entire NBA has increased
 their volume of three point shot attempts in response to the
-recent success of the Golden State Warriors and Stephen Curry. Due to this fact, 
-there is a growing consensus the three point shot is most important in basketball. 
+recent success of the Golden State Warriors and Stephen Curry.  Due to this
+fact, there is a growing consensus the three point shot is crucial because
+of the efficient scoring and floor spacing it provides offenses.  
 However, two point shots are not to be overlooked and may
@@
 to \citet{Harris}. In fact, 2020 NBA champion Los Angeles Lakers actually led league
-in two point percentage throughout the playoffs and were bottom five in three point percentage.
+in two-point shooting percentage throughout the playoffs and were bottom 
+five in three-point shooting percentage.
 
@@
 
-Home-court advantage is the basic idea that the home team is more likely to win.
+Home-court advantage is the basic idea that the home team is more likely to win given 
+they benefit from positive effects of a few factors we discussed earlier.
 So laying a foundation of typical home-court advantage is crucial. Before
@@
 fairly bell shaped, which is important for statistical tests
-designed for normally distribute data. There appears to be little
-difference between the 2020 and 2017-2019 for home scoring, but for
+designed for normally distributed data. There appears to be little
+difference between the 2020 and 2017-2019 for home scoring. For
 away scoring, a more pronounced shift to the right in 2020 is observed
@@
 two-point shooting percentage for away teams has a small shift to the
-right in 2020 relative to that in 2017-2019.
+right in 2020 relative 2017-2019.
 
@@
 exhibition/preseason games, the NBA always has a home and away team. So, for
-the first time in NBA history the bubble allows NBA home and away performance
-to be compared vs a control/neutral field. The NBA bubble, as a neutral court,
-removed all 4 possible game location factors impacting home-court advantage
-hypothesized by @Carron1992. The NBA bubble featured 8 seeding games then a
-standard playoff format. The focus of this study was the play during the playoff
-games since it followed the standard playoff format and can easily be compared
+the first time in NBA history, the bubble allows NBA home and away performance
+to be compared against a control/neutral field. The NBA bubble
+removed many, if not all, of the factors impacting home-court advantage. The NBA bubble featured 8 seeding games 
+for each team, then a standard playoff format. The focus of this study was on the play during the playoff
+games since it followed the standard playoff format and could be compared
 back to other playoffs. For this study, the 2020 playoffs were compared against
@@
 for the changing play style of the NBA, we limit the study to 2020 vs 2017-2019
-for the faster pace play and more common use of the three-point shot
+due to the faster pace play and more common use of the three-point shot
 in modern basketball. If we used data from say 10 years ago, or earlier, observed
@@
 away performance and how playing at home may positively impact home performance.
-If there are differences in scoring for home or away teams the differences can
+If there are differences in scoring for home or away teams, the differences can
 be used to show how home-court advantage affects overall performance of home
@@
 Shooting percentages are not the only possible metrics affected by home-court
-advantage, but they are the most obvious and certainly an important one.
+advantage, but they are the most obvious and likely most important one.
 
@@
 We also conducted nonparametric tests that are distribution free to
-confirm the results from the $z$-test. For question~1, we used
-Fisher's exact test for a contigency table which summarizes the wins
+confirm the results from the $z$-test. For question 1, we used
+Fisher's exact test for a contingency table which summarizes the wins
 and losses of the home team in the 83 games in 2020 and the 243 games
 in 2017-2019. For all other eight questions, the data are the scores
-or shooting percentages from the 84 games in 2020 and the 243 games in
+or shooting percentages from the 83 games in 2020 and the 243 games in
 2017-2019. We used Wilcoxan's rank-sum test.
@@
 $z$-tests and Wilcoxon tests. The p-values are all fairly similar for both
-tests giving strong confidence in conclusions drawn. Also reported are 
+tests giving strong confidence in conclusions drawn. Also, reported are 
 the point estimates of the two samples in each comparison.
@@
 # Discussion {#sec:disc}
-Generally it seemed that away teams fared better in the 2020 NBA playoff bubble
+Generally, it seemed that away teams fared better in the 2020 NBA playoff bubble
 than previous years on the road. Starting from the dip in home winning percentage
-to below 0.482 it is clear that something was different. Although the difference was not
+to below 0.482, it is clear that something was different. Although, the difference was not
 significant after a Bonferroni correction it is still informative to consider and
 understand that home teams seemed to struggle to win compared to normal conditions.
-Compared to @Kotecki who finds home teams consistently have a significantly
-better record than away teams boasting about a 60.5% win percentage in his sample,
+Compared to @Kotecki, who finds home teams consistently have a significantly
+better record than away teams boasting about a 60.5% win percentage, in his sample,
 the 48.2% home winning percentage of 2020 home teams is quite a shift.
-In this study, home teams did not benefit from the usual advantages provided by
+In this study, home teams did not appear to benefit from the usual advantages provided by
 being the home team.
@@
 shots. Normally, by shooting more two-pointers themselves and forcing away teams to
-shoot more two-pointers the home team benefits from increasing effects of
+shoot more two-pointers, the home team benefits from increasing effects of
 home-court advantage. However, with away teams significantly improving two-point
-shooting in the bubble this strategy was no longer viable and home-court
+shooting in the bubble, this strategy was no longer viable and home-court
 advantage disappeared.
 
-To get a bit more detailed, we saw an estimated 3\% increase in away two point
+To get a bit more detailed, we saw an estimated 3\% increase in away two-point
 shooting percentage. Likewise, NBA teams shot
@@
 they can take their time finding a good shot which can
-still either be a two or three, given the score and remaining time. In the context 
+still either be a two or three given the score and remaining time. In the context 
 of the NBA bubble, this crucial possession also has the benefit of the away team
@@
 same situation while being down six the offense must rush to get a quick shot
-off, which probably has to be a three pointer. This is a much more desperate
-circumstance, less likely to have a positive outcome and even in the positive
+off, which probably has to be a three pointer. This three point swing creates a much more desperate
+circumstance which is less likely to have a positive outcome. Even in the positive
 outcome case you're still going to be down three and likely have to foul the
@@
 seven or eight points, but the alternative is let the opponent run the 24 second shot 
-clock down wasting time. Also, the away team may even be in a situation where they're
+clock down wasting time. Similarly, the away team may even be in a situation where they're
 leading by the three additional points, which for the same logic as explained above 
@@
 driving factor of away teams closing the gap with home teams.
+
+There are a few possibilities that might create this increase in two-point shooting percentage. 
+One could be the possible removal of officiating bias in favor of the home team during the
+bubble. Away teams in the bubble may have benefited from more fouls called on drives to
+the basket. On drives to the basket there can often be a lot of contact with no clear
+foul, these calls are then up to the ref's discretion. Typically, refs may be more
+reluctant to blow the whistle against the home team, in front of their crowd, on this 
+type of play. The NBA only records an attempted shot when a foul is called on the shot
+if the shot is made. So, when away teams are fouled, but it is not called because it's 
+a often at the ref's discretion to make the call, they'll likely miss and be credited the 
+miss. However, in the bubble if fouls are called more fairly, the missed shots from uncalled
+fouls are removed and two-point shooting would increase. This wouldn't affect three-point 
+shooting because those fouls are more clear and less up to ref's discretion. Another possibility
+, without the crowd noise inhibiting their offense, away teams were able to more easily
+run their offensive sets that generate easy two-point looks at the rim. Also, away teams
+being generally more confident without opposing fans present, may have been more inclined
+to attack the basket and get an easy look close to the rim. It's hard to say for sure what 
+cause the increase in away two-point shooting, but these are all possibilities. 
+It may also be a combination of all of these.
 
@@
 the beginning of possibilities for studies using the 2020 NBA bubble as a case study
-for home-court advantage. Although the study is limited by a one time
+for home-court advantage. Although, the study is limited by a one time
 sample, it seems unlikely that these conditions will ever be

### 2022-02-02 Manuscript/NBABubbleAnalysis.Rmd
Update NBABubbleAnalysis.Rmd

@@
 shows that the 2020 playoffs favored away teams more than usual, particularly
-with two point shooting and total scoring.
+with two-point shooting and total scoring.
 
@@
 differences in away scoring in 2020 vs 2017-2019. Also, differences in team
-shooting (for-two pointers, three-pointers, and free throws) from 2020 vs 2017-2019
+shooting (for two-pointers, three-pointers, and free throws) from 2020 vs 2017-2019
 for both home and away teams. The results from these tests bring a new perspective
@@
 line. @Harris used two-point shots, three-point shots and free throws as measures
-of interest to study home-court advantage. Two point shots were found to be
+of interest to study home-court advantage. Two-point shots were found to be
 the strongest predictor of home-court advantage. They suggested
-that home teams should try to shoot more two point shots and force
-their opponent to take more two point shot attempts. This will 
+that home teams should try to shoot more two-point shots and force
+their opponent to take more two-point shot attempts. This will 
 help home teams have a greater control of the game play and help maximize the benefits 
-home court advantange.
+home-court advantange.
 
@@
 Lastly, as discussed in the introduction @Lopez find evidence that geographic
-factors like altitude may influence and strengthen home court advantage for teams in
+factors like altitude may influence and strengthen home-court advantage for teams in
 high altitude regions.
@@
 point shot has grown very prominent in basketball. The entire NBA has increased
-their volume of three point shot attempts in response to the
+their volume of three-point shot attempts in response to the
 recent success of the Golden State Warriors and Stephen Curry.  Due to this
-fact, there is a growing consensus the three point shot is crucial because
+fact, there is a growing consensus the three-point shot is crucial because
 of the efficient scoring and floor spacing it provides offenses.  
-However, two point shots are not to be overlooked and may
+However, two-point shots are not to be overlooked and may
 actually be most important, especially for home teams, according
@@
 insight to answering the questions presented. Regression is a popular
-tool in home court advantage studies, but as mentioned previously, this
+tool in home-court advantage studies, but as mentioned previously, this
 study is very different from past studies. Rather than trying to prove the
-existence of home court advantage, like many previous studies, this paper
-was more interested in generating a clearer understanding of how home court
+existence of home-court advantage, like many previous studies, this paper
+was more interested in generating a clearer understanding of how home-court
 advantage effects each team. The most effective way to do that is hypothesis
@@
 shooting percentage. Likewise, NBA teams shot
-an average of about 53 two point shots per game, in our data, leading to a three 
+an average of about 53 two-point shots per game, in our data, leading to a three 
 point increase in total scoring. Even this
@@
 same situation while being down six the offense must rush to get a quick shot
-off, which probably has to be a three pointer. This three point swing creates a much more desperate
+off, which probably has to be a three-pointer. This three point swing creates a much more desperate
 circumstance which is less likely to have a positive outcome. Even in the positive
@@
 three, you can play regular defense without fouling, get a stop, then draw up a play 
-to score a three pointer with the remaining six seconds. However, in the same scenario 
+to score a three-pointer with the remaining six seconds. However, in the same scenario 
 down six, you have no choice, but to foul. The free throws will likely put you down
@@
 Since, home teams picked up an additional two points per game on average vs. five 
-for away teams, the three additional points from two point shooting appear to be the 
+for away teams, the three additional points from two-point shooting appear to be the 
 driving factor of away teams closing the gap with home teams.
@@
 advantage mostly plays into games by negatively impacting away teams. If fans cause
-overall performance to drop, then home court advantage must come from a bigger drop
+overall performance to drop, then home-court advantage must come from a bigger drop
 in away performance than the drop in home performance. This is why away teams were
@@
 of the main factors that influence an offense's ability to score is the opposing
-defense. It's possible that home court advantage mainly functions as an extra
+defense. It's possible that home-court advantage mainly functions as an extra
 defender for home teams, which is how it negatively impacts away scoring. There
@@
 bubble provided the same quiet environment they're used to at home while on
-offense. I still believe home court advantage is better categorized as a
+offense. I still believe home-court advantage is better categorized as a
 negative effect on away team offense, since it's more likely
@@
 answer more questions with a larger sample size. For example,  testing
-strength of home court advantage relative to specific teams, like the
+strength of home-court advantage relative to specific teams, like the
 Denver Nuggets and Utah Jazz who we know from other literature may
-have a stronger home court advantage. The sample size of home games
+have a stronger home-court advantage. The sample size of home games
 for the individual teams in the bubble was far too small to try and

### 2022-02-04 Manuscript/NBABubbleAnalysis.Rmd
Update NBABubbleAnalysis.Rmd
General grammar check

@@
 factor to the outcome of games. It is well-known that the home team typically
-benefits from some competitive edge while playing at their home-court, resulting
+benefits from some competitive edge while playing at their homecourt, resulting
 in a better chance of winning. Thus, the NBA playing the 2020 playoffs in a
@@
 create a chaotic and stressful environment for away teams to play in. There 
-is no way replicate these effects in the bubble conditions. Also, there are 
+is no way to replicate these effects in the bubble conditions. Also, there are 
 geographic factors that are lost, like altitude and time-zone effects. There's 
@@
 teams coming in and struggling with lower oxygen levels at a higher altitude, 
-which the home teams are accustomed to playing with \citet{Lopez}. Examples,
+which the home teams are accustomed to playing with \citet{Lopez}. Examples
 of high altitude teams in the NBA would be the Utah Jazz and Denver Nuggets, 
@@
 and living in the comfort of their own home are also lost. Travel is also 
-normally discussed as a factor hurting away teams, but for the context of
+normally discussed as a factor hurting away teams, but in the context of
 the playoffs in the bubble, losing this may not be significant. In the
 playoffs, many times (aside from the first game of the series when the 
-away team travels) both teams often travel at the same time since they're 
+away team travels), both teams often travel at the same time, since they're 
 both going back and forth between the two cities to complete the series.
@@
 teams outperform away teams, but also separate the
-effects of home-court advantage on home teams and away teams performance individually.
+effects of home-court advantage on home teams' and away teams' performance individually.
 
@@
 for both home and away teams. The results from these tests bring a new perspective
-to understanding of how home-court advantage impacts games by altering the performance
+to the understanding of how home-court advantage impacts games by altering the performance
 of the home and away teams.
@@
 throws. However, they did not find significant evidence
-evidence that home status has a substantial impact on missing from the free throw
+that home status has a substantial impact on missing from the free throw
 line. @Harris used two-point shots, three-point shots and free throws as measures
@@
 help home teams have a greater control of the game play and help maximize the benefits 
-home-court advantange.
+home-court advantage.
 
@@
 during the college basketball season. This allowed them to construct two samples,
-one of home teams and one of neutral teams. They formulate their study in a
+one of home teams and one of neutral teams. They formulated their study in a
 regression predicting the expected difference in score for home teams.
@@
 may benefit from some advantages in rules and officiating. They acknowledge that
-these factors would all be removed if games were played at a neutral site even
+these factors would all be removed if games were played at a neutral site, even
 if one team was designated as “home team”. This study was reviewed a decade later
 by @Carron2005. The 2005 review goes over the new findings from studies
-about the significance of these four game location factors. Since 1992 they found
+about the significance of these four game location factors. Since 1992, they have found
 that results on these four factors are mixed. However, there is some evidence
@@
 also account for referee bias which may impact the NBA. One interesting finding cited
-by @Carron2005 is that in the absence of crowds result in overall performance increases.
-Lastly, as discussed in the introduction @Lopez find evidence that geographic
+by @Carron2005 is that the absence of crowds results in overall performance increases.
+Lastly, as discussed in the introduction, @Lopez find evidence that geographic
 factors like altitude may influence and strengthen home-court advantage for teams in
@@
 recent success of the Golden State Warriors and Stephen Curry.  Due to this
-fact, there is a growing consensus the three-point shot is crucial because
+fact, there is a growing consensus that the three-point shot is crucial because
 of the efficient scoring and floor spacing it provides offenses.  
@@
 actually be most important, especially for home teams, according
-to \citet{Harris}. In fact, 2020 NBA champion Los Angeles Lakers actually led league
+to \citet{Harris}. In fact, the 2020 NBA champion Los Angeles Lakers actually led the league
 in two-point shooting percentage throughout the playoffs and were bottom 
@@
 
-The data was collected on a game by game basis, this gave us two observations
+The data was collected on a game by game basis. This gave us two observations
 for each variable per game played, one observation for each team(home and away).
@@
 important. The winner of a game is determined by who scores more points. There can only be
-one winner and one loser making the outcome a binary variable, with
+one winner and one loser, making the outcome a binary variable, with
 one indicating a win and zero indicating a loss.
@@
 Home-court advantage is the basic idea that the home team is more likely to win given 
-they benefit from positive effects of a few factors we discussed earlier.
+they benefit from the positive effects of a few factors we discussed earlier.
 So laying a foundation of typical home-court advantage is crucial. Before
-focusing on the 2017 to 2020 playoffs we can take a quick look at home team win
+focusing on the 2017 to 2020 playoffs, we can take a quick look at home team win
 percentages since 2010. Notice in Figure \@ref(fig:Fig1), the 10 years before 2020, the home team
 winning percentage ranged from around 0.56 to 0.7 and never dipped below 0.5. The
-2020 bubble broke this historic pattern dipping down below 0.5. Foreshadowing,
+2020 bubble broke this historic pattern, dipping down below 0.5. Foreshadowing
 the confirmation of the expectation that the effect of home-court advantage was removed
@@
 away scoring, a more pronounced shift to the right in 2020 is observed
-compared with that in 2017-2019.
+compared to 2017-2019.
 
@@
 for each team, then a standard playoff format. The focus of this study was on the play during the playoff
-games since it followed the standard playoff format and could be compared
+games, since it followed the standard playoff format and could be compared
 back to other playoffs. For this study, the 2020 playoffs were compared against
@@
 for the changing play style of the NBA, we limit the study to 2020 vs 2017-2019
-due to the faster pace play and more common use of the three-point shot
+due to the faster pace of play and more common use of the three-point shot
 in modern basketball. If we used data from say 10 years ago, or earlier, observed
@@
 If there are differences in scoring for home or away teams, the differences can
-be used to show how home-court advantage affects overall performance of home
+be used to show how home-court advantage affects the overall performance of home
 and away teams. While testing for differences in shooting will provide added
@@
 was more interested in generating a clearer understanding of how home-court
-advantage effects each team. The most effective way to do that is hypothesis
+advantage affects each team. The most effective way to do that is hypothesis
 testing to compare performance in the neutral bubble to previous years with
-normal game conditions for both the home and away team. Another potential
+normal game conditions for both the home and away teams. Another potential
 benefit of regression would have been the ability to control for factors
-like strength of opponent. However, we felt that strength of team was
+like the strength of opponent. However, we felt that the strength of team was
 already well enough controlled for by the fact that we focused on the
@@
 continue to remove the less competitive teams, so that the talent disparity
-between teams is much smaller than regular season play where controlling for
-opponent strength is likely more important.
+between teams is much smaller than regular season play, where controlling for
+opponent strength is more likely to be important.
 
@@
 $z$-tests and Wilcoxon tests. The p-values are all fairly similar for both
-tests giving strong confidence in conclusions drawn. Also, reported are 
+tests, giving strong confidence in the conclusions drawn. Also, reported are 
 the point estimates of the two samples in each comparison.
@@
 First, we see a statistically significant change
-in home win percentage in 2020 from 2017-19, with p-value of 0.0497 for the
+in home win percentage in 2020 from 2017-19, with a p-value of 0.0497 for the
 $z$-test and 0.0400 for Fisher's exact test. The 95% confidence
@@
 for free throws. The $z$-test p-value of 0.0692 indicates no significant difference, while
-the Wilcoxon test p-value of 0.0469 indicates a difference at the 5% significance level. Since,
-the p-value of Wilcoxon test is so close to significance level and neither p-value is significant
-after a Bonferonni correction for multiple tests this difference is likely not very meaningful.
+the Wilcoxon test p-value of 0.0469 indicates a difference at the 5% significance level. Since
+the p-value of the Wilcoxon test is so close to significance level and neither p-value is significant
+after a Bonferonni correction for multiple tests, this difference is likely not very meaningful.
 There appears to be no strong evidence suggesting home teams played at
@@
 Away teams saw more of an impact than home teams. For starters, there is a
-significant increase in mean points per game, indicated by p-value of 0.0008 for $z$-test and
+significant increase in mean points per game, indicated by a p-value of 0.0008 for $z$-test and
 0.0004 for Wilcoxon. It is important to note both p-values also remain significant after
@@
 be about 5 points, with 95% CI $(2.083, 7.988)$. Likewise, the away team two-point
-shooting efficiency increased significantly based on p-value of 0.0003
-for both the $z$-test and Wilcoxon test. Again, both p-values remain significant after Bonferroni correction.
+shooting efficiency increased significantly based on a p-value of 0.0003
+for both the $z$-test and Wilcoxon test. Again, both p-values remain significant after the Bonferroni correction.
 The average difference was estimated to be about 0.03, with 95% CI $(0.015, 0.050)$. However,
@@
 than previous years on the road. Starting from the dip in home winning percentage
-to below 0.482, it is clear that something was different. Although, the difference was not
-significant after a Bonferroni correction it is still informative to consider and
+to below 0.482, it is clear that something was different. Although the difference was not
+significant after a Bonferroni correction, it is still informative to consider and
 understand that home teams seemed to struggle to win compared to normal conditions.
 Compared to @Kotecki, who finds home teams consistently have a significantly
-better record than away teams boasting about a 60.5% win percentage, in his sample,
+better record than away teams, boasting about a 60.5% win percentage in his sample,
 the 48.2% home winning percentage of 2020 home teams is quite a shift.
@@
 decreasing. If away teams are scoring significantly more and home teams are not,
-then we expect to see away teams winning a larger amount of games. This may give
+then we expect to see away teams winning a larger number of games. This may give
 more reason to believe the conclusion that there was a significant decrease in
@@
 shots. Normally, by shooting more two-pointers themselves and forcing away teams to
-shoot more two-pointers, the home team benefits from increasing effects of
+shoot more two-pointers, the home team benefits most from increasing effects of
 home-court advantage. However, with away teams significantly improving two-point
@@
 small swing in scoring for away teams can make a big difference in outcome of
-games. Around 1/3 of NBA games end by a decision of six points or less, those
+games. Around 1/3 of NBA games end by a decision of six points or less, these
 are the games where the three points matter most. Suppose there are 30 seconds
@@
 getting to work in silence without the jeers from enemy fans. However, in the
-same situation while being down six the offense must rush to get a quick shot
+same situation, while being down six the offense must rush to get a quick shot
 off, which probably has to be a three-pointer. This three point swing creates a much more desperate
@@
 three, you can play regular defense without fouling, get a stop, then draw up a play 
-to score a three-pointer with the remaining six seconds. However, in the same scenario 
+to score a three-pointer with the remaining six seconds. In the same scenario, 
 down six, you have no choice, but to foul. The free throws will likely put you down
-seven or eight points, but the alternative is let the opponent run the 24 second shot 
+seven or eight points, but the alternative is letting the opponent run the 24 second shot 
 clock down wasting time. Similarly, the away team may even be in a situation where they're
@@
 if the shot is made. So, when away teams are fouled, but it is not called because it's 
-a often at the ref's discretion to make the call, they'll likely miss and be credited the 
-miss. However, in the bubble if fouls are called more fairly, the missed shots from uncalled
+often at the ref's discretion to make the call, they'll likely miss and be credited the 
+miss. However, in the bubble, if fouls are called more fairly, the missed shots from uncalled
 fouls are removed and two-point shooting would increase. This wouldn't affect three-point 
-shooting because those fouls are more clear and less up to ref's discretion. Another possibility
+shooting because those fouls are clearer and less up to the ref's discretion. Another possibility
 , without the crowd noise inhibiting their offense, away teams were able to more easily
@@
 to attack the basket and get an easy look close to the rim. It's hard to say for sure what 
-cause the increase in away two-point shooting, but these are all possibilities. 
+causes the increase in away two-point shooting, but these are all possibilities. 
 It may also be a combination of all of these.
@@
 
-It's worth noting that much of this paper discussed the the effects from an
-offensive point of view, focusing on a teams ability to score. However, one
-of the main factors that influence an offense's ability to score is the opposing
+It's worth noting that much of this paper discusses the effects from an
+offensive point of view, focusing on a team's ability to score. However, one
+of the main factors that affects an offense's ability to score is the opposing
 defense. It's possible that home-court advantage mainly functions as an extra
@@
 offensively in the bubble. Home teams did not have the regular help from the
-noise of fans that hinder the away team's offensive efforts. This would also
+noise of fans that hinders the away team's offensive efforts. This would also
 explain why home team offense didn't show significant change in the bubble, the
@@
 negative effect on away team offense, since it's more likely
-noise from fans lowers level of play of away offense rather than raise the level
+noise from fans lowers level of play of away offense rather than raises the level
 of play of home defense. This will be further discussed below.
@@
 could help strengthen the conclusion in this paper. Also, you could
-answer more questions with a larger sample size. For example,  testing
+answer more questions with a larger sample size. For example, testing the
 strength of home-court advantage relative to specific teams, like the

### 2022-02-04 Manuscript/manuscript.tex
Update manuscript.tex
Updated Tex for submission file

@@
-\documentclass[12pt, letterpaper, titlepage]{article}
+% Created by Bonita Graham
+% Last update: December 2019 By Kestutis Bendinskas
 
-\usepackage{amsmath}
-\usepackage{booktabs}
-\usepackage{amsthm}
+% Authors: 
+% Please do not make changes to the preamble until after the solid line of %s.
+
+\documentclass[10pt]{article}
+\usepackage[explicit]{titlesec}
+\setlength{\parindent}{0pt}
+\setlength{\parskip}{1em}
+\usepackage{hyphenat}
+\usepackage{ragged2e}
+\RaggedRight
+
+% These commands change the font. If you do not have Garamond on your computer, you will need to install it.
+\usepackage{garamondx}
+\usepackage[T1]{fontenc}
+\usepackage{amsmath, amsthm}
 \usepackage{graphicx}
-\usepackage[margin=1in]{geometry}
-\usepackage {hyperref}
-\hypersetup {colorlinks = true, linkcolor = blue, citecolor=blue, urlcolor = blue}
-\usepackage[]{natbib}
-\usepackage{enumitem}
-\usepackage{setspace}
 
-\usepackage[pagewise]{lineno}
-\linenumbers*[1]
-% %% patches to make lineno work better with amsmath
-\newcommand*\patchAmsMathEnvironmentForLineno[1]{%
- \expandafter\let\csname old#1\expandafter\endcsname\csname #1\endcsname
- \expandafter\let\csname oldend#1\expandafter\endcsname\csname end#1\endcsname
- \renewenvironment{#1}%
- {\linenomath\csname old#1\endcsname}%
- {\csname oldend#1\endcsname\endlinenomath}}%
-\newcommand*\patchBothAmsMathEnvironmentsForLineno[1]{%
- \patchAmsMathEnvironmentForLineno{#1}%
- \patchAmsMathEnvironmentForLineno{#1*}}%
+% This adjusts the underline to be in keeping with word processors.
+\usepackage{soul}
+\setul{.6pt}{.4pt}
 
-\AtBeginDocument{%
- \patchBothAmsMathEnvironmentsForLineno{equation}%
- \patchBothAmsMathEnvironmentsForLineno{align}%
- \patchBothAmsMathEnvironmentsForLineno{flalign}%
- \patchBothAmsMathEnvironmentsForLineno{alignat}%
- \patchBothAmsMathEnvironmentsForLineno{gather}%
- \patchBothAmsMathEnvironmentsForLineno{multline}%
+% The following sets margins to 1 in. on top and bottom and .75 in on left and right, and remove page numbers.
+\usepackage{geometry}
+\geometry{vmargin={1in,1in}, hmargin={.75in, .75in}}
+\usepackage{fancyhdr}
+\pagestyle{fancy}
+\pagenumbering{gobble}
+\renewcommand{\headrulewidth}{0.0pt}
+\renewcommand{\footrulewidth}{0.0pt}
+
+% These Commands create the label style for tables, figures and equations.
+\usepackage[labelfont={footnotesize,bf} , textfont=footnotesize]{caption}
+\captionsetup{labelformat=simple, labelsep=period}
+\newcommand\num{\addtocounter{equation}{1}\tag{\theequation}}
+\renewcommand{\theequation}{\arabic{equation}}
+\makeatletter
+\renewcommand\tagform@[1]{\maketag@@@ {\ignorespaces {\footnotesize{\textbf{Equation}}} #1.\unskip \@@italiccorr }}
+\makeatother
+\setlength{\intextsep}{10pt}
+\setlength{\abovecaptionskip}{2pt}
+\setlength{\belowcaptionskip}{-10pt}
+
+\renewcommand{\textfraction}{0.10}
+\renewcommand{\topfraction}{0.85}
+\renewcommand{\bottomfraction}{0.85}
+\renewcommand{\floatpagefraction}{0.90}
+
+% These commands set the paragraph and line spacing
+\titleformat{\section}
+  {\normalfont}{\thesection}{1em}{\MakeUppercase{\textbf{#1}}}
+\titlespacing\section{0pt}{0pt}{-10pt}
+\titleformat{\subsection}
+  {\normalfont}{\thesubsection}{1em}{\textit{#1}}
+\titlespacing\subsection{0pt}{0pt}{-8pt}
+\renewcommand{\baselinestretch}{1.15}
+
+% This designs the title display style for the maketitle command
+\makeatletter
+\newcommand\sixteen{\@setfontsize\sixteen{17pt}{6}}
+\renewcommand{\maketitle}{\bgroup\setlength{\parindent}{0pt}
+\begin{flushleft}
+\sixteen\bfseries \@title
+\medskip
+\end{flushleft}
+\textit{\@author}
+\egroup}
+\makeatother
+
+% This styles the bibliography and citations.
+%\usepackage[biblabel]{cite}
+\usepackage[sort&compress]{natbib}
+\setlength\bibindent{2em}
+\makeatletter
+\renewcommand\@biblabel[1]{\textbf{#1.}\hfill}
+\makeatother
+\renewcommand{\citenumfont}[1]{\textbf{#1}}
+\bibpunct{}{}{,~}{s}{,}{,}
+\setlength{\bibsep}{0pt plus 0.3ex}
+
+\usepackage{float}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+% Authors: Add additional packages and new commands here.  
+% Limit your use of new commands and special formatting.
+
+% Place your title below. Use Title Capitalization.
+\title{The Effects of the NBA COVID Bubble on the NBA Playoffs: A Case Study for Home-Court Advantage}
+
+% Add author information below. Communicating author is indicated by an asterisk, the affiliation is shown by superscripted lower case letter if several affiliations need to be noted.
+\author{
+Michael Price*$^{a}$ \\ \medskip 
+$^{a}$Department of Statistics, University of Connecticut, Storrs, CT  \\  \medskip 
+\\ \medskip 
+Students: michael.price@uconn.edu* \\
+Mentor: jun.yan@uconn.edu 
 }
 
-\title{The Effects of the NBA COVID Bubble on the NBA Playoffs: A Case Study for Home-Court Advantage}
+\pagestyle{empty}
+\begin{document}
 
-% \author{Michael Price and Jun Yan\\
-% \href{mailto:michael.price@uconn.edu}{\nolinkurl{michael.price@uconn.edu}}\\
-% Department of Statistics, University of Connecticut}
-\author{annonymous}
-\date{December 1, 2020}
+% Makes the title and author information appear.
+\vspace*{.01 in}
+\maketitle
+\vspace{.12 in}
 
-\begin{document}
-\maketitle
-
-\doublespace
-
-\begin{abstract}
-The 2020 NBA playoffs were played inside of a bubble in Disney World because of
+% Abstracts are required.
+\section*{abstract}
+The 2020 NBA playoffs were played inside of a bubble at Disney World because of
 the COVID-19 pandemic. This meant that there were no fans in attendance,
-games played on neutral courts and no traveling for teams, which in theory
-removes home-court advantage from the games. This setting has attracted much
-discussion as analysts and fans debated the possible effects it may have on the
+games were played on neutral courts and no traveling for teams. In theory, these conditions should
+remove home-court advantage from the games. This setting generated
+discussion and concern, as analysts and fans debated the possible effects it may have on the
 outcome of games. Home-court advantage has historically played an influential
-role in NBA playoff series outcomes. The 2020 playoff provided a unique
+role in NBA playoff series outcomes. The 2020 playoffs provided a unique
 opportunity to study the effects of the bubble and home-court
@@
 shows that the 2020 playoffs favored away teams more than usual, particularly
-with two point shooting and total scoring.
-\end{abstract}
+with two-point shooting and total scoring.
 
-\hypertarget{sec:intro}{%
-\section{Introduction}\label{sec:intro}}
+% Keywords are required.
+\section*{keywords} 
+NBA; NBA Covid; NBA Bubble; HomeCourt Advantage;
+
+\vspace{.12 in}
+
+% Start the main part of the manuscript here.
+% Comment out section headings if inappropriate to your discipline.
+% If you add additional section or subsection headings, use an asterisk * to avoid numbering.
+
+\section{Introduction}
 
@@
 factor to the outcome of games. It is well-known that the home team typically
-benefits from some competitive edge from playing at their home-court, resulting
+benefits from some competitive edge while playing at their homecourt, resulting
 in a better chance of winning. Thus, the NBA playing the 2020 playoffs in a
 bubble due to the COVID-19 pandemic
-brought a great deal of concern for fans, teams, journalists, and others.
-For example, \citet{Aschburner} discusses about anticipated effects, sharing
+brought a great deal of concern for fans, teams, journalists, and others. The bubble
+environment would not be able to replicate many of the factors that normal home-court
+advantage relies on. Among the biggest is the effect from the crowd. The home crowd has
+the ability to energize home teams, help control momentum in favor of the home team and 
+create a chaotic and stressful environment for away teams to play in. There is no way
+to replicate these effects in the bubble conditions. Also, there are geographic factors
+that are lost, like altitude and time-zone effects. There's evidence across all sports
+that teams in high altitude regions can rely on away teams coming in and struggling with
+the lower oxygen levels at a higher altitude, which the home teams are accustomed to playing
+with \citet{Lopez}. Examples of high altitude teams in the NBA would be the Utah Jazz
+and Denver Nuggets, both of whom participated in the 2020 NBA bubble. Also, referees 
+are known to be subject to favoring home teams. This is due in large part to pressure
+from fans, which is no longer in play in the bubble. Refs may still have biases, but 
+there's nothing to sway that bias towards the home team like normal. Any positive
+effects of the home team playing in the arena they're most comfortable in and
+living in the comfort of their own home are also lost. Travel is also normally 
+discussed as a factor hurting away teams, but in the context of the playoffs
+being played in the bubble, losing this may not be significant. In the playoffs,
+many times (aside from the first game of the series when the away team travels), 
+both teams often travel at the same time, since they're both going back and forth
+between the two cities to complete the series.
+
+ \citet{Aschburner} discusses the anticipated effects, sharing
 concerns from former players, coaches and other experts about
@@
 result of home teams overperforming or away teams underperforming because of
-home-court advantage. Some of these studies are reviewed in greater detail in
-Section\textasciitilde\ref{sec:litrev}.
+home-court advantage. Some of these studies are reviewed in greater detail in Section 2.
 
@@
 differences in away scoring in 2020 vs 2017-2019. Also, differences in team
-shooting (for-two pointers, three-pointers, and free throws) from 2020 vs 2017-2019
+shooting (for two-pointers, three-pointers, and free throws) from 2020 vs 2017-2019
 for both home and away teams. The results from these tests bring a new perspective
-to understanding of how home-court advantage impacts games by altering the performance
+to the understanding of how home-court advantage impacts games by altering the performance
 of the home and away teams.
 
-\hypertarget{sec:litrev}{%
-\section{Literature Review}\label{sec:litrev}}
+\section{Literature Review}
+There is voluminous literature on the effects of home-court advantage.
+Many NBA home-court advantage studies analyze the effects by studying shooting
+percentages. \citet{Kotecki} reported significant evidence of home-court
+advantage by comparing field goal
+percentage, free throw percentage and points scored in home vs.~away teams.
+He found all of these measures showed evidence that
+home-court advantage helps home teams play better. \citet{Cao} studied the effects
+of pressure on performance in the NBA. Using free throws as their measure of interest, 
+they tested whether home
+fans could distract and put pressure on opposing players to make free
+throws. However, they did not find significant evidence
+that home status has a substantial impact on missing from the free throw line. 
+\citet{Harris} used two-point shots, three-point shots and free
+throws as measures of interest to study home-court advantage. Two-point
+shots were found to be the strongest predictor of home-court advantage. They
+suggested that home teams should try to shoot more two-point shots and force
+their opponent to take more two-point shot attempts. This help home teams have a
+greater control of the game play and help maximize the benefits 
+home-court advantage.
 
-There is a voluminous literature on the effects of home-court advantage.
-Many NBA home-court advantage studies analyze the effects by studying shooting
-percentages. \citet{Kotecki} reported significant home-court advantage using performance-based
-statistics, specifically field goal percentage, free throw percentage, and
-points by comparing home performance vs.~away performance in games.
-All of which he showed significantly indicates that
-home-court helps teams play better. \citet{Cao} studied the effects
-of pressure on performance in the NBA. Using throws as their measure of
-interest, they tested whether home
-fans could distract and put pressure on opposing players to make free
-throws. However, they found insignificant
-evidence that home status has a substantial impact on missing from the free throw
-line. \citet{Harris} used two point shots, three-point shots and free throws as measures
-of interest to study home-court advantage. Two point shots were found to be
-the strongest predictor of home-court advantage. They suggested
-that home teams should try to shoot more two point shots and force
-their opponent to take more two point shot attempts. This strategy will maximize
-the benefits of home-court advantage and give them the best chance to win.
-
-Some studies focus less on shooting and more on scoring differences and other metrics.
+Some studies focus less on shooting and more on other metrics.
 For example, \citet{Greer} focused on the influence of spectator booing on
@@
 may benefit from some advantages in rules and officiating. They acknowledge that
-these factors would all be removed if games were played at a neutral site even
+these factors would all be removed if games were played at a neutral site, even
 if one team was designated as ``home team''. This study was reviewed a decade later
 by \citet{Carron2005}. The 2005 review goes over the new findings from studies
-about the significance of these four game location factors. Since 1992 they found
+about the significance of these four game location factors. Since 1992, they have found
 that results on these four factors are mixed. However, there is some evidence
-supporting crowd and travel factors impact games in the NBA. There is less
-evidence suggesting learning and rule factors impact the NBA. One interesting
-finding cited by \citet{Carron2005} is that in the absence of crowds result
-in overall performance increases.
+supporting crowd and travel factors impact games across all major sports. There is less
+evidence suggesting learning and rule factors impact across the various collegiate 
+and professional sports. The NBA is not a league which has rules that
+may favor the home team, like batting last in the MLB, but these rule
+factors also account for referee bias which may impact the NBA.  One
+interesting
+finding cited by \citet{Carron2005} is that the absence of crowds results
+in overall performance increases. Lastly, as discussed in the introduction, \citet{Lopez}
+find evidence that geographic factors like altitude may influence and strengthen home-court
+advantage for teams in high altitude regions.
 
-\hypertarget{sec:data}{%
-\section{Data}\label{sec:data}}
+\section{Data}
 
@@
 free throws. These variables were very popular and frequently used in the related
-literature discussed earlier. The data was collected on a game by game basis, this gave us two observations
+literature discussed earlier. While many other measures could be used for examining
+the outcome of the game and team performance, scoring seemed to be the most
+important because the winner of a game is determined by who scores more points. Furthermore, 
+the three types of shots are a
+natural discussion point in basketball, as they're all
+important and directly impact scoring totals. Free throws can often determine the winner of close
+games, especially when losing teams are forced to foul the winning team to
+stop the clock and hope for some missed free throws. The three
+point shot has grown very prominent in basketball. The entire NBA has increased their
+volume of three-point shot attempts in response to the
+recent success of the Golden State Warriors and Stephen Curry. Due to this
+fact, there is a growing consensus that the three-point shot is crucial because of 
+the efficient scoring and floor spacing it provides offenses. 
+However, two-point shots are not to be overlooked and may
+actually be most important, especially for home teams, according
+to \citet{Harris}. In fact, the 2020 NBA champion Los Angeles Lakers actually led the 
+league in two-point shooting percentage throughout the playoffs and were bottom five
+in three-point shooting percentage.
+
+The data was collected on a game by game basis. This gave us two observations
 for each variable per game played, one observation for each team(home and away).
@@
 for both home and away teams over 2017-2020 (486 observations total).
-While many other measures could be used for measuring
-the outcome of the game and team performance, scoring seemed to be the most
-important. The winner of a game is determined by who scores more points. There can only be
-one winner and one loser making the outcome a binary variable, with
+ There can only be
+one winner and one loser, making the outcome a binary variable, with
 one indicating a win and zero indicating a loss.
 
-Home-court advantage is the basic idea that the home team is more likely to win.
+Home-court advantage is the basic idea that the home team is more likely to win given 
+they benefit from positive effects of a few factors we discussed earlier.
 So laying a foundation of typical home-court advantage is crucial. Before
-focusing on the 2017 to 2020 playoffs we can take a quick look at home team win
+focusing on the 2017 to 2020 playoffs, we can take a quick look at home team win
 percentages since 2010. Notice in Figure \ref{fig:Fig1}, the 10 years before 2020, the home team
 winning percentage ranged from around 0.56 to 0.7 and never dipped below 0.5. The
-2020 bubble broke this historic pattern dipping down below 0.5. Foreshadowing,
+2020 bubble broke this historic pattern, dipping down below 0.5. Foreshadowing
 the confirmation of the expectation that the effect of home-court advantage was removed
@@
 
-\begin{figure}
+\begin{figure}[H]
   \centering
-  \includegraphics[width=0.9\linewidth]{Fig1}
+  \includegraphics[width=0.4\linewidth]{Fig1}
   \caption{Winning percentage of NBA home teams in the playoffs since
@@
 fairly bell shaped, which is important for statistical tests
-designed for normally distribute data. There appears to be little
-difference between the 2020 and 2017-2019 for home scoring, but for
-away scoring, a noticeable shift to the right in 2020 is observed
-compared with that in 2017-2019.
+designed for normally distributed data. There appears to be little
+difference between the 2020 and 2017-2019 for home scoring. For
+away scoring, a more pronounced shift to the right in 2020 is observed
+compared to 2017-2019.
 
-\begin{figure}
+\begin{figure}[H]
   \centering
-  \includegraphics[width=0.9\linewidth]{Fig2}
+  \includegraphics[width=0.4\linewidth]{Fig2}
   \caption{Histograms of home (blue) and away (red) scoring
@@
 two-point shooting percentage for away teams has a small shift to the
-right in 2020 relative to that in 2017-2019.
+right in 2020 relative to 2017-2019.
 
-\begin{figure}
+\begin{figure}[H]
   \centering
-  \includegraphics[width=0.9\linewidth]{Fig3}
+  \includegraphics[width=0.4\linewidth]{Fig3}
   \caption{Histograms of home shooting percentages for two-
@@
 
-\begin{figure}
+\begin{figure}[H]
   \centering
-  \includegraphics[width=0.9\linewidth]{Fig4}
+  \includegraphics[width=0.4\linewidth]{Fig4}
   \caption{Histograms of away shooting percentages for two-pointers,
@@
 
-\hypertarget{sec:methods}{%
-\section{Methods}\label{sec:methods}}
+\section{Methods}
 
@@
 exhibition/preseason games, the NBA always has a home and away team. So, for
-the first time in NBA history the bubble allows NBA home and away performance
-to be compared vs a control/neutral field. The NBA bubble, as a neutral court,
-removed all 4 possible game location factors impacting home-court advantage
-hypothesized by \citet{Carron1992}. The NBA bubble featured 8 seeding games then a
-standard playoff format. The focus of this study was the play during the playoff
-games since it followed the standard playoff format and can easily be compared
+the first time in NBA history, the bubble allows NBA home and away performance
+to be compared against a control/neutral field. The NBA bubble
+removed many, if not all, factors impacting home-court advantage. The NBA 
+bubble featured 8 seeding games for each team, then a
+standard playoff format. The focus of this study was on the play during the playoff
+games, since it followed the standard playoff format and could be compared
 back to other playoffs. For this study, the 2020 playoffs were compared against
@@
 for the changing play style of the NBA, we limit the study to 2020 vs 2017-2019
-for the faster pace play and more common use of the three-point shot
+due to the faster pace of play and more common use of the three-point shot
 in modern basketball. If we used data from say 10 years ago, or earlier, observed
@@
 away performance and how playing at home may positively impact home performance.
-If there are differences in scoring for home or away teams the differences can
-be used to show how home-court advantage affects overall performance of home
+If there are differences in scoring for home or away teams, the differences can
+be used to show how home-court advantage affects the overall performance of home
 and away teams. While testing for differences in shooting will provide added
@@
 We also conducted nonparametric tests that are distribution free to
-confirm the results from the \(z\)-test. For question\textasciitilde1, we used
-Fisher's exact test for a contigency table which summarizes the wins
+confirm the results from the \(z\)-test. For question 1, we used
+Fisher's exact test for a contingency table which summarizes the wins
 and losses of the home team in the 83 games in 2020 and the 243 games
 in 2017-2019. For all other eight questions, the data are the scores
-or shooting percentages from the 84 games in 2020 and the 243 games in
+or shooting percentages from the 83 games in 2020 and the 243 games in
 2017-2019. We used Wilcoxan's rank-sum test.
@@
 
-\hypertarget{sec:results}{%
-\section{Results}\label{sec:results}}
+Regression was also considered, but it did not seem to add any additional 
+useful insight to answering the questions presented. Regression is a popular
+tool in home-court advantage studies, but as mentioned previously, this
+study is very different from past studies. Rather than trying to prove the
+existence of home-court advantage, like many previous studies, this paper
+was more interested in generating a clearer understanding of how home-court
+advantage affects each team. The most effective way to do that is hypothesis
+testing to compare performance in the neutral bubble to previous years with
+normal game conditions for both the home and away teams. Another potential
+benefit of regression would have been the ability to control for factors
+like the strength of opponent. However, we felt that the strength of team was
+already well enough controlled for by the fact that we focused on the
+playoffs. The playoffs only include the strongest 16 of 30 teams, then
+continue to remove the less competitive teams, so that the talent disparity
+between teams is much smaller than regular season play, where controlling for
+opponent strength is more likely to be important.
 
-\begin{table}
-  \caption{The results from the 9 tests.}
+\section{Results}
+\bigskip
+\begin{table}[H]
+  \caption{The results from the 9 tests(* denotes significant p-value at alpha=0.05)
+  \\*Note: Home and Away scoring totals are divided by 100}
   \label{tab:table}
@@
 \(z\)-tests and Wilcoxon tests. The p-values are all fairly similar for both
-tests giving strong confidence in conclusions drawn. Also reported are
+tests giving strong confidence in the conclusions drawn. Also, reported are
 the point estimates of the two samples in each comparison.
@@
 First, we see a statistically significant change
-in home win percentage in 2020 from 2017-19, with p-value of 0.0497 for the
+in home win percentage in 2020 from 2017-19, with a p-value of 0.0497 for the
 \(z\)-test and 0.0400 for Fisher's exact test. The 95\% confidence
 interval (CI) of \((-0.255, -0.008)\)
-confirms our belief that home-court advantage was lost in the 2020 NBA playoffs. However, after accounting for
-multiple tests using the Bonferroni correction, the p-values for both tests are no longer significant.
-So, we may only cautiously say there is evidence that home-court advantage was
+confirms our belief that home-court advantage was lost in the 2020 NBA 
+playoffs. However, after accounting for
+multiple tests using the Bonferroni correction, the p-values for both
+tests are no longer significant. So, we may only cautiously say there is 
+evidence that home-court advantage was
 not a factor in 2020.
@@
 for free throws. The \(z\)-test p-value of 0.0692 indicates no significant difference, while
-the Wilcoxon test p-value of 0.0469 indicates a difference at the 5\% significance level. Since,
-the p-value of Wilcoxon test is so close to significance level and neither p-value is significant
-after a Bonferonni correction for multiple tests this difference is likely not very meaningful.
+the Wilcoxon test p-value of 0.0469 indicates a difference at the 5\% significance level. Since
+the p-value of the Wilcoxon test is so close to significance level and neither p-value is significant
+after a Bonferonni correction for multiple tests, this difference is likely not very meaningful.
 There appears to be no strong evidence suggesting home teams played at
@@
 
-\hypertarget{sec:disc}{%
-\section{Discussion}\label{sec:disc}}
+\section{Discussion}
 
-Generally it seemed that away teams fared better in the 2020 NBA playoff bubble
-than previous years on the road. Starting from the dip in home winning percentage
-to below 0.482 it is clear that something was different. Although the difference was not
-significant after a Bonferroni correction it is still informative to consider and
-understand that home teams seemed to struggle to win compared to normal conditions.
-Compared to \citet{Kotecki} who finds home teams consistently have a significantly
-better record than away teams boasting about a 60.5\% win percentage in his sample,
-the 48.2\% home winning percentage of 2020 home teams is quite a shift.
-In this study, home teams did not benefit from the usual advantages provided by
-being the home team.
+Generally, it seemed that away teams fared better in the 2020 NBA playoff bubble
+than previous years on the road. Starting from the dip in home winning percentage to
+below 0.482, it is clear that something was different. Although the difference
+was not significant after a Bonferroni correction, it is still informative to 
+consider and understand that home teams seemed to struggle to win compared to
+normal conditions. Compared to \citet{Kotecki}, who finds home teams consistently 
+have a significantly better record than away teams, boasting about a 60.5\% win 
+percentage in his sample, the 48.2\% home winning percentage of 2020 home teams 
+is quite a shift. In this study, home teams did not appear to benefit from the 
+usual advantages provided by being the home team.
 
-Away team average
-scoring did increase by a statistically significant amount. This goes hand in
-hand with our intuition and conclusion about the home winning percentage
-decreasing. If away teams are scoring significantly more and home teams are not,
-then we expect to see away teams winning a larger amount of games. This may give
-more reason to believe the conclusion that there was a significant decrease in
-home winning percentage in 2020, despite failing to be significant
-after the Bonferroni
-correction. Only away team scoring being significantly impacted by playing on
-a neutral court indicates that home-court advantage stems mainly from adverse
-effects on the visiting team.
+Away team average scoring did increase by a statistically significant amount.
+This goes hand in hand with our intuition and conclusion about the home winning 
+percentage decreasing. If away teams are scoring significantly more and home 
+teams are not, then we expect to see away teams winning a larger number of games. 
+This may give more reason to believe the conclusion that there was a significant 
+decrease in home winning percentage in 2020, despite failing to be significant after 
+the Bonferroni correction. Only away team scoring being significantly impacted by 
+playing on a neutral court and not the home scoring indicates that home-court 
+advantage stems mainly from adverse effects on the visiting team.
 
-An interesting finding is all shooting and scoring numbers for both home and
-away teams did make at least small increases. Although these increases
-were not
-all significant these increases are exactly what is reported in \citet{Carron2005}
-when they explain how evidence suggests that teams perform better with the absence
-of fans. This is important because it coincides with our conclusion that home-court
-advantage mostly plays into games by negatively impacting away teams. If fans cause
-overall performance to drop, then home court advantage must come from a bigger drop
-in away performance than the drop in home performance. This is why away teams were
-able to close the gap with home teams with home-court advantage removed.
+At least some of that improvement from away teams came from significantly higher
+two-point efficiency. This corresponds with the conclusion from \citet{Harris},
+where they found home teams are best suited to capitalize on advantages from two-point
+shots. Normally, by shooting more two-pointers themselves and forcing away teams to
+shoot more two-pointers, the home team benefits most from effects of
+home-court advantage. However, with away teams significantly improving two-point
+shooting in the bubble, this strategy was no longer viable and home-court
+advantage disappeared.
+
+To get a bit more detailed, we saw an estimated 3\% increase in away two-point
+shooting percentage. Likewise, NBA teams shot
+an average of about 53 two-point shots per game, in our data, leading to a three-point increase in total scoring. Even this
+small swing in scoring for away teams can make a big difference in outcome of
+games. Around 1/3 of NBA games end by a decision of six points or less, these
+are the games where the three points matter most. Suppose there are 30 seconds
+left and the away team is losing. If the away team has the ball, only down three, they can take their time finding a good shot which can
+still either be a two or three given the score and remaining time. In the context 
+of the NBA bubble, this crucial possession also has the benefit of the away team
+getting to work in silence without the jeers from enemy fans. However, in the
+same situation, while being down six, the offense must rush to get a quick shot
+off which probably has to be a three-pointer. The three point swing creates a much more desperate
+circumstance that is less likely to have a positive outcome. Even in the positive
+outcome case you're still going to be down three and likely have to foul the
+opposing team after you score. 
+Alternatively, if the away team is playing defense with 30 seconds left, only down 
+three, you can play regular defense without fouling, get a stop, then draw up a play
+to score a three-pointer with the remaining six seconds. In the same scenario, down 
+six, you have no choice, but to foul. The free throws will likely put you down seven
+or eight points, but the alternative is letting the opponent run the 24 second shot 
+clock down wasting time. Similarly, the away team may even be in a situation where 
+they're leading by the three additional points, which for the same logic as explained
+above puts them in far greater control of the game. Clearly, the three point swing 
+can drastically change probability of winning for a team down the stretch of a game. 
+Since, home teams picked up an additional two-points per game on average vs. five for 
+away teams, the three additional points from two-point shooting appear to be the driving
+factor of away teams closing the gap with home teams.
+
+There are a few possibilities that might create this increase in two-point shooting
+percentage. One could be the possible removal of officiating bias in favor of the 
+home team during the bubble. Away teams in the bubble may have benefited from more 
+fouls called on drives to the basket. On drives to the basket there can often be a 
+lot of contact with no clear foul, these calls are then up to the ref's discretion.
+Typically, refs may be more reluctant to blow the whistle against the home team, in
+front of their crowd, on this type of play. The NBA only records an attempted shot 
+when a foul is called on the shot if the shot is made. So, when away teams are 
+fouled, but it is not called because it's often at the ref's discretion to make 
+the call, they'll likely miss and be credited the miss. However, in the bubble, if
+fouls are called more fairly, the missed shots from uncalled fouls are removed and 
+two-point shooting would increase. This wouldn't affect three-point shooting because
+those fouls are clearer and less up to the ref's discretion. Another possibility,
+without the crowd noise inhibiting their offense, away teams were able to more 
+easily run their offensive sets that generate easy two-point looks at the rim. 
+Also, away teams being generally more confident without opposing fans present, 
+may have been more inclined to attack the basket and get an easy look close to 
+the rim. It's hard to say for sure what causes the increase in away two-point shooting,
+but these are all possibilities. It may also be a combination of all of these.
 
@@
 
-At least some of that improvement from away teams came from significantly higher
-two-point efficiency. This corresponds with the conclusion from \citet{Harris},
-where they found home teams are best suited to capitalize on advantages from two-point
-shots. Normally, by shooting more two-pointers themselves and forcing away teams to
-shoot more two-pointers the home team benefits from increasing effects of
-home-court advantage. However, with away teams significantly improving two-point
-shooting in the bubble this strategy was no longer viable and home-court
-advantage disappeared.
+It's worth noting that much of this paper discusses the effects from an
+offensive point of view, focusing on a team's ability to score. However, one of 
+the main factors that affects an offense's ability to score is the opposing
+defense. It's possible that home-court advantage mainly functions as an extra
+defender for home teams, which is how it negatively impacts away scoring. There
+is not much home fans can do to help their team on offense, except being quiet
+so players can easily communicate and focus. However, when on defense, fans can
+act as a "sixth defender" supporting their team by cheering loudly and making it
+harder on the offense. This would explain how away teams were able to improve
+offensively in the bubble. Home teams did not have the regular help from the
+noise of fans that hinders the away team's offensive efforts. This would also
+explain why home team offense didn't show significant change in the bubble, the
+in-game offensive environment was largely unaltered for home teams since the
+bubble provided the same quiet environment they're used to at home while on
+offense. I still believe home-court advantage is better categorized as a
+negative effect on away team offense, since it's more likely
+noise from fans lowers level of play of away offense rather than raises the level
+of play of home defense. This will be further discussed below.
+
+An interesting finding is all shooting and scoring numbers for both home and
+away teams did make at least small increases. Although, these increases
+were not all significant these increases are exactly what is reported in
+\citet{Carron2005} when they explain how evidence suggests that teams perform
+better with the absence of fans. This is important because it coincides with our
+conclusion that home-court advantage mostly plays into games by negatively impacting 
+away teams by acting as an extra defender. If fans cause overall performance to drop,
+then home-court advantage must come from a bigger drop in away performance than the 
+drop in home performance. This is why away teams were able to close the gap with 
+home teams with home-court advantage removed.
 
@@
 same measures with a different sample. Otherwise, that type of study
-could help strengthen the conclusion in this paper.
+could help strengthen the conclusion in this paper. Also, you could
+answer more questions with a larger sample size. For example, testing
+the strength of home-court advantage relative to specific teams, like the
+Denver Nuggets and Utah Jazz who we know from other literature may
+have a stronger home-court advantage. The sample size of home games
+for the individual teams in the bubble was far too small to try and
+address this question, the Jazz and Nuggets only had 3 and 9 home
+games, respectively.
 
@@
 
+\section*{acknowledgements}
+The authors thank the University of Connecticut and the UConn Department of Statistics.
+%Note correct LaTeX quotations above. Do not use the " symbol, but rather double ` followed by double '
+
+% The About the Student Author section is NOT optional.  Write a paragraph about the student; see previous journal editions for examples.
+% If there is more than one student author, you must move the comment below.
+\section*{about the student author}
+%\section*{about the student authors}
+Michael Price graduated in May 2021 from the University of Connecticut where he double-majored in Statistics and Economics and minored in Data Analytics.
+He began his graduate studies at the University of Delaware studying Applied Statistics in August 2021.
+
+% The Press Summary section is NOT optional.  Write a paragraph describing the paper in a manner suitable for the press; see previous journal editions for examples.
+\section*{press summary}
+The purpose of this study was to examine the 2020 NBA playoffs, which were played inside of a bubble in Disney World because of the COVID-19 pandemic.
+The hope was to generate new insights about the effects of home-court advantage because the bubble created an unprecedented neutral playing field for the NBA.
+For the first time the effects of home-court advantage on home and away teams could be easily separated and studied individually, typically studies can only compare them relative to each other.
+This study is focused on team scoring totals and team shooting percentage on two-pointers, three-pointers, and free throws.
+Comparing these measures for home teams and away teams in 2020 vs.
+2017-2019 shows that the 2020 playoffs favored away teams more than usual, particularly with two-point shooting and total scoring.
+The implication of these findings is home-court advantage seems to be the result of negative effects on away team, not positive effects on home-team.
+
 \end{document}

### 2022-02-05 Response_To_Comments.tex
Create Response_To_Comments.tex

\documentclass[12pt]{article}
\usepackage[margin=1in]{geometry}

\usepackage{xr-hyper}
\RequirePackage[colorlinks,citecolor=blue,urlcolor=blue]{hyperref}
\usepackage{listings}
\usepackage{rotating, graphicx}
\usepackage{booktabs, natbib}

\usepackage{amsmath, enumerate, quoting, amssymb, subfigure, url}

\usepackage{color}
\newcommand{\red}[1]{\textcolor{red}{#1}}
\newcommand{\blue}[1]{\textcolor{blue}{#1}}
\newcommand{\gv}[1]{\textcolor{blue}{(GV: #1)}}
\DeclareMathOperator*{\argmin}{arg\,min}
\DeclareMathOperator*{\cov}{Cov}
\DeclareMathOperator*{\corr}{Corr}

\usepackage{xr}
\externaldocument{isgee}
\externaldocument{supp}

% remove left indentation in itemize
\usepackage{enumitem}
\setlist[itemize]{leftmargin=*}

\usepackage{xcolor}
\newcommand{\jy}[1]{\textcolor{red}{(JY: #1)}}

% graph path
\graphicspath{{./figs/}}

\quotingsetup{leftmargin = 0pt}

\newenvironment{comment}%
{\begin{quoting}\noindent\small\it\ignorespaces%
  }{\end{quoting}}

\begin{document}

\begin{center}
  {\Large\bf Response to the Comments on Manuscript}
\end{center}

\section*{Summary}

We thank the two referees for their insightful, constructive
comments. In particular, we are grateful to referee~1 for providing many helpful, direct and concise comments using annotations. Also, to referee~2, who took the time to provide detailed explanations for each suggestion making it clear how and why the changes were important.

The manuscript has been revised to incorporate the comments from the
referees with the following most notable changes:

\begin{enumerate}
\item
  Additional commentary and thoughts on what caused away teams to be more affected by neutral courts more than home teams.
  \item
  A deeper discussion on the significance and impact of the shift in two-point shooting.
\item
  Alterations to graphs and charts to better and more clearly describe the data and findings.
\end{enumerate}

Point-by-point responses are as follows.

\section*{To Referee~1}

\begin{comment}
  I’d have the paper examined for grammar.  I think there are commas missing or misplaced periodically. Couldn’t add that to the pdf effectively.
\end{comment}

 The paper was carefully reread and edited, line by line, in order to correct grammar errors and improve the overall flow of the paper.

\begin{comment}
Comments in the pdf. 
\end{comment}

There were a series of brief comments provided on an annotated version of our manuscript submission. This included many small comments for missing words or asking for a brief clarification about a point made. These comments were addressed accordingly and were resolved. The more extensive comments/concerns are addressed individually below.

\begin{comment}
  Reference lines (mean/median) would be nice in the graphs to more clearly see the small shifts.
\end{comment}

 Code was added to the ggplot code which added a reference line at the mean for all applicable histograms.

\begin{comment}
Use of adjusted p-values was appreciated, but those should be listed in the table.  I would also recommend using either the nonparametric (Wilcoxon) or parametric (t-test) exclusively—no need for both.  I think the assumptions are satisfied for the parametric test.  Fisher’s exact test is appropriate
\end{comment}

 Columns were added to the results table to accommodate including the adjusted p-values. As for including both the non-parametric and parametric tests, we think including both is justified. We felt it strengthens the confidence in the results when the conclusions from each the non-parametric and parametric tests agree with one another. While the assumptions appear to be met for parametric test, it does not hurt to show tests not dependent on assumptions also show significant results.

\begin{comment}
 Why are two pointers singled out in the data? Any insight on why 2P\% is affected more? I would provide some insight. My hypothesis is officiating bias as shown in many papers (less discretion on 3P fouls). 
\end{comment}

 We thank the reviewer for bringing this concern to our attention. We realize we missed discussing the significance of studying each type of shot individually, not just 2-pointers. Discussion on the importance of examining each of the three types of shot was added, including the 2-point shot.

\begin{comment}
Comment a bit on these effect sizes. Does a 3\% decrease in 2P\% matter in a game? That may only be 1 shot. While it's statistically significant, I wonder if it's practically significant.
\end{comment}

 A paragraph was added to the discussion section detailing the practical significance of the 3\% change in 2-point shooting. In short, many NBA games are decided by 6 points or less and in those games late-game strategy is completely dependent exactly how many points one team is behind. Even a small shift in score can drastically shift probability of outcomes and strategy for teams. The paper now goes into detail on how a 3-point swing in these close games can completely change the game.

\begin{comment}
Would be interesting to examine if these home court advantages are more pronounced for specific teams.
\end{comment}

 We are appreciative of this suggestion for another possible topic of analysis. Cross referencing this comment with the source provided by referee~2, we definitely have reason to believe the shift in home-court advantage would be more pronounced for specific teams. However, we didn't have the sample size to test for strength home-court advantage in individual teams. The teams we'd expect to have the most pronounced affect, the Denver Nuggets and Utah Jazz, only played 9 and 3 home games, respectively, in the bubble. We did add some comments about this in the conclusion explaining how this analysis would be possible if generating a larger sample were possible and what we'd expect to see.

\section*{Referee 2}

\subsection*{Major comments}

\begin{comment}
Good motivation of the problem and why we have additional useful information to answer questions about
home advantage.
It would be good to mention here, or elsewhere, different (potential) sources of home advantage, and discuss
which may/may not still exist in the NBA bubble. For example,
\item
• Travel miles/time
\item
• Time zone changes
\item
• Days rest (back-to-back)
\item
• Altitude
\item
• Attendance/crowd
\item
• Referee bias
\item
• Being at home, familiar surroundings, sleeping in own bed
\item
This is partly in the literature review where you discuss Courneya and Carron’s article, but I think you could
touch on that there are different sources of home advantage somewhere in the introduction. There are, for
example, no travel miles differences between the two teams and (I think) no back-to-back games, so home
advantage from those probably doesn’t exist in the bubble. There may, for example, still be referee bias in
the bubble, though. Referee bias may only exist when there’s a home crowd, this case the video monitors
and the handful of attendees don’t affect referee. Or it might exist because referees know who the home
team is.
\end{comment}

 We thank the reviewer for being appreciative and understanding of the good motivations driving this paper. We agree that a deeper discussion of why home-court advantage would not exist in the bubble would be beneficial to setting up the paper. We added in a discussion of why all the different factors play into creating home-court advantage and how each one may or may not still exist in the bubble. For the most part, these factors disappeared in the context of the bubble, except travel/time and miles for reasons we discuss in the paper.

\begin{comment}
Also, this paper https://arxiv.org/abs/1701.05976 finds that altitude matters (Colorado and Utah teams
have the highest home advantage in multiple sports), so that could be added/discussed. So changes in home
advantage in the bubble could impact teams differently, and NBA teams Nuggets and Jazz could be hurt
more than other teams since they no longer benefit from an altitude advantage.
\end{comment}

 We are very grateful referee~2 brought this additional source to our attention. It was added to Literature Review and discussed in the Introduction when setting up and explaining the factors that would be missing from normal home-court advantage when playing in the NBA bubble. The most noteworthy point added from this paper, as pointed out by referee~2, was the evidence supporting geographic factors impact home-court advantage. Particularly, teams like the Utah Jazz and Denver Nuggets are the NBA teams who would benefit from the high-altitude effects. This is discussed in more detail in the manuscript.

\begin{comment}
In your plots with histograms, it would be useful to have a vertical line showing the mean to visually compare
the two distributions.

It seems like you made this using facet\_grid(shot.type ~ season) or something similar. I think it would
be easy to add vertical lines fairly easily by using group\_by and summarise to make a data frame df.means
like this

season shot.type mean.perc

2020 2p xx.x

2020 3p xx.x

2020 ft xx.x

2017-2019 2p xx.x

2017-2019 3p xx.x

2017-2019 ft xx.x

and then adding
geom\_vline(data=df.means, aes(x=mean.perc)) +
somewhere in your ggplot code.

If the columns season and shot type are the same as in your other data
set, and the same names used in facet\_grid, a different vertical line should show up for each season and shot type.
\end{comment}

 While this was not the exact coding solution used to add reference lines to the histograms, it was greatly appreciated and was the perfect stepping stone to finding the code solution. As mentioned earlier, the code was added to include a reference line at the mean on each histogram we provided.

\begin{comment}
If you use a regression, you could take into account strength of team, strength of opponent, and home
advantage simultaneously. This could be done with any of the metrics you considered. For example,
Score ~ 1 + HomeTeamOnOffenseIndicator + OffensiveTeam + DefensiveTeam
Did you consider a regression? Why would it, or would it not, be useful? If strength of opponent (in the
example above, the DefensiveTeam) doesn’t vary much from team to team (in other words, if there is a
balanced schedule), then it might not be that helpful to account for strength of opponent. If it does vary a
lot, then it might help.
In a typical regular season, the NBA schedule is pretty balanced. Is the schedule a balance schedule in the
bubble?
Would accounting for team strength change the home advantage estimates? If so, in what situations would
it increase home advantage estimates? Decrease?
You don’t necessarily have to do all of this analysis, but you should at least discuss somewhere why you
didn’t use a regression, and justify your decision. I’m not saying that a regression is necessarily better,
and the answers to the above questions might suggest it wouldn’t help. But a lot of people would think of
estimating home advantage using a regression by including a home indicator like the example above, and
then using the hypothesis test where the null hypothesis is that the regression coefficient for home indicator
is zero. So it is important to, at minimum, discuss your choice to not use a regression, and justify that
choice.
\end{comment}

 This is certainly a interesting suggestion and something we explored. However, we elected to not add the regression to the paper. We did take the suggestion to add a paragraph to the paper explaining why the hypothesis testing alone sufficed to answering our research questions without need for the regression. We did not consider accounting for strength of opponent necessary. Given the study was conducted using the playoff teams, match-ups were largely between strong teams with small gaps in ability. Had the study been done using regular season play, controlling for strength of opponent would have been beneficial since bottom-tier teams would be included. More details on why the regression was not required are included in the manuscript.

\begin{comment}
I’d be interested in hearing more thoughts about why away teams were helped, but home teams weren’t
hurt, in the bubble. You did mention “No longer having to face the struggle of traveling, pressure from
opposing fans, or playing on an unfamiliar court”. A little more detail would be nice. Of the factors that
impact home advantage, which ones were unchanged for the home team, but changed for the away team,
and in what specific situation?
For example, when a team has two home games in a row, they didn’t travel pre-bubble in between those
games, and still didn’t travel in the bubble. If their opponent, the away team, had 2 away games in a row,
they did travel pre-bubble, but didn’t have to travel in the bubble. So in that case there is a difference in
home advantage.
But if the home team instead had an away game before their current home game, then they had to travel
before their home game, and a lack of travel advantage for the home team doesn’t apply to that game. I
would imagine this happens frequently, and that 30-40\% of home games are preceded by an away game.
Also, if away teams were impacted more, wouldn’t they also be better on defense, which would cause the
home team’s score to go down, for example? Do the home advantage only impact offensive performance?
\end{comment}

 Thank you for bringing this up, this comment allowed us to revisit the discussion and derive a more clear well-thought out conclusion from the results. A few paragraphs were added to the discussion section of the paper detailing why it is likely only away teams performance were significantly changed. The discussion strengthened our determination that home-court advantage is best categorized as negative effects on away team.

\begin{comment}
About study author-remove “will”
\end{comment}

The misplaced word "will" was removed from the about the author section.

\bibliographystyle{chicago}
\bibliography{fpref}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:

### 2022-02-05 Manuscript/manuscript.tex
fixed author and compiling errors/warnings

@@
 \usepackage{graphicx}
+\usepackage{hyperref}
 
@@
 \author{
-Michael Price*$^{a}$ \\ \medskip 
+Michael Price*$^{a}$, Jun Yan \\ \medskip 
 $^{a}$Department of Statistics, University of Connecticut, Storrs, CT  \\  \medskip 
-\\ \medskip 
+% \\ \medskip 
 Students: michael.price@uconn.edu* \\
@@
 \pagestyle{empty}
+
 \begin{document}

### 2022-02-05 Manuscript/Response_To_Comments.tex
comments on revised manuscript and response

--- 2022-02-05 Response_To_Comments.tex
@@
 \section*{Summary}
+\jy{Practice the line width style in tex source}
 
 We thank the two referees for their insightful, constructive
-comments.
-In particular, we are grateful to referee~1 for providing many helpful, direct and concise comments using annotations.
-Also, to referee~2, who took the time to provide detailed explanations for each suggestion making it clear how and why the changes were important.
+comments. In particular, we are grateful to referee~1 for providing
+many helpful, direct and concise comments using annotations. Also, to
+referee~2, who took the time to provide detailed explanations for each
+suggestion making it clear how and why the changes were important.
 
 The manuscript has been revised to incorporate the comments from the
-referees with the following most notable changes:
+referees with the following most notable additions:
 
@@
 
-Point-by-point responses are as follows.
+Point-by-point responses are as follows with the original comments in \emph{italic}.
+
+\jy{In each response, be more specific on the location of the changes made (give
+  paragraph numbers in which section)}
 
@@
 
-\begin{comment}
-Comments in the pdf. 
-\end{comment}
+% \begin{comment}
+% Comments in the pdf. 
+% \end{comment}
 
@@
 
- Code was added to the ggplot code which added a reference line at the mean for all applicable histograms.
+The figures were remade with added reference line at the mean for all applicable histograms.
 
@@
 
- We are very grateful referee~2 brought this additional source to our attention.
+ We are very grateful that referee~2 brought \citet{Lopez} to our attention.
 It was added to Literature Review and discussed in the Introduction when setting up and explaining the factors that would be missing from normal home-court advantage when playing in the NBA bubble.
@@
 This is discussed in more detail in the manuscript.
+
+ \jy{Check the references of Lopez's paper to see if we can cite
+   anything there; this is also a place to mention the additional
+   references you found. }
 
@@
 
-The misplaced word "will" was removed from the about the author section.
+Done.
 
 \bibliographystyle{chicago}
-\bibliography{fpref}
+\bibliography{citations}
 

### 2022-02-05 Manuscript/manuscript.tex
comments on revised manuscript and response

@@
 \usepackage{hyperref}
+\usepackage{xcolor}
+\newcommand{\jy}[1]{\textcolor{orange}{JY: (#1)}}
+\usepackage{booktabs}
 
@@
 \author{
-Michael Price*$^{a}$, Jun Yan \\ \medskip 
-$^{a}$Department of Statistics, University of Connecticut, Storrs, CT  \\  \medskip 
-% \\ \medskip 
+Michael Price*, Jun Yan \\ \medskip 
+Department of Statistics, University of Connecticut, Storrs, CT  \\  \medskip 
 Students: michael.price@uconn.edu* \\
@@
 factor to the outcome of games. It is well-known that the home team typically
-benefits from some competitive edge while playing at their homecourt, resulting
+benefits from some competitive edge while playing at their home court, resulting
 in a better chance of winning. Thus, the NBA playing the 2020 playoffs in a
@@
 We will compare home team performance in 2020 at a neutral site with
-no fans vs.~2017-19 playoffs with fans. Likewise, away team performance in 2020
-at a neutral site with no fans vs.~2017-19 playoffs with fans. By comparing home
-teams in 2020 to home teams in 2017-19 and away teams in 2020 to away teams in
-2017-19, we add a new perspective to the field of research. This will allow for
+no fans vs.~2017--19 playoffs with fans. Likewise, away team performance in 2020
+at a neutral site with no fans vs.~2017--19 playoffs with fans. By comparing home
+teams in 2020 to home teams in 2017--19 and away teams in 2020 to away teams in
+2017--19, we add a new perspective to the field of research. This will allow for
 a more accurate understanding of the effects of home-court advantage on home and
@@
 years. First, whether or not the difference between home win percentage in 2020
-and 2017-19 is zero. This difference is found to be statistically
+and 2017--19 is zero. This difference is found to be statistically
 significant from zero. Then we assess for differences in home
@@
 \section{Literature Review}
+
+\jy{add some references from google scholar search ``home team
+  advantage nba''; we have only 10 references now.}
+
 There is voluminous literature on the effects of home-court advantage.
@@
 focusing on the 2017 to 2020 playoffs, we can take a quick look at home team win
-percentages since 2010. Notice in Figure \ref{fig:Fig1}, the 10 years before 2020, the home team
+percentages since 2010. Notice in Figure~\ref{fig:Fig1}, the 10 years before 2020, the home team
 winning percentage ranged from around 0.56 to 0.7 and never dipped below 0.5. The
@@
 
-\begin{figure}[H]
+\jy{use `tbp' for figure/table locations instead of `H', unless it is
+  required by the journal}
+
+\jy{remake the figure~\ref{fig:Fig1}: make aspect 3:4 (height:width;
+  increase the font size of the annonations; control the height/weight
+  from within R now that we know how the figures will appear in the
+  published version (so that you don't need to control width in latex.}
+
+\jy{explain what the vertical lines are in all the figure captions}
+
+\begin{figure}[tbp]
   \centering
@@
 Moving on to the main focus of the study, comparing 2020 to 2017-2019.
-Figure \ref{fig:Fig2} shows the histograms of the home (green) and
+Figure~\ref{fig:Fig2} shows the histograms of the home (green) and
 away scoring (red) for 2020 vs.~2017-2019. All histograms are
@@
 compared to 2017-2019.
+
+\jy{remake the figure~\ref{fig:Fig2}: make each panel aspect ratio 3:4; same for the
+  next figures.}
 
@@
 Our second target of inference is shooting percentage for home and away teams.
-Figure \ref{fig:Fig3} shows home shooting for two-pointers, three-pointers and
-free throws for 2020 (top) vs.~2017-19 (bottom). The histograms appear to be fairly
-similarly distributed between 2020 and 2017-19. Likewise, Figure \ref{fig:Fig4},
+Figure~\ref{fig:Fig3} shows home shooting for two-pointers, three-pointers and
+free throws for 2020 (top) vs.~2017--19 (bottom). The histograms appear to be fairly
+similarly distributed between 2020 and 2017--19. Likewise, Figure~\ref{fig:Fig4},
 shows the same percentages except for away teams. It appears that the
@@
   \caption{Histograms of home shooting percentages for two-
-    pointers, three-pointers and free throws for 2020 (top) vs.~2017-19
+    pointers, three-pointers and free throws for 2020 (top) vs.~2017--19
     (bottom).}
@@
   \caption{Histograms of away shooting percentages for two-pointers,
-    three-pointers and free throws for 2020 (top) vs.~2017-19
+    three-pointers and free throws for 2020 (top) vs.~2017--19
     (bottom).}
@@
 
-Comparisons between 2020 and 2017-19 home and away teams were made on home team
+Comparisons between 2020 and 2017--19 home and away teams were made on home team
 winning percentage, total team scoring and two-point, three-point and free throw
@@
 \section{Results}
-\bigskip
-\begin{table}[H]
+
+\jy{use booktabs package to add horizontal rules; replace 1 with
+  1.0000; make sure ``Z-test'' is in the same font as in the text.}
+
+\begin{table}[tbp]
   \caption{The results from the 9 tests(* denotes significant p-value at alpha=0.05)
@@
 \centering
-\begin{tabular}[t]{lcccccc}
-  & 2020 & 2017-19 & \multicolumn{2}{c}{P-value} & \multicolumn{2}{c}{Adjusted P-value}\\
-  &          &                & Z-test & Wilcoxan & Z-test & Wilcoxan\\
-Home Win & 0.482 & 0.613 & 0.0497* & 0.0400* & .4473 & 0.3598\\
+\begin{tabular}[t]{lccllll}
+  \toprule
+  & 2020 & 2017--19 & \multicolumn{2}{c}{P-value} & \multicolumn{2}{c}{Adjusted P-value}\\
+\cmidrule(lr){4-5}\cmidrule(lr){6-7}
+  &          &                & Z-test & Wilcoxon & Z-test & Wilcoxon\\
+  \midrule
+  Home Win & 0.482 & 0.613 & 0.0497* & 0.0400* & 0.4473 & 0.3598\\
 Home Scoring & 1.101 & 1.081 & 0.2321 & 0.2985 & 1 & 1\\
@@
 Away FT & 0.783 & 0.777 & 0.6601 & 0.8370 & 1 & 1\\
+  \bottomrule
 \end{tabular}
@@
 
-Starting from the top of the
-Table \ref{tab:table} summarizes p-values of the nine hypotheses for both
+Starting from the top,
+Table~\ref{tab:table} summarizes p-values of the nine hypotheses for both
 \(z\)-tests and Wilcoxon tests. The p-values are all fairly similar for both
@@
 
+\jy{add a reference to Bonferroni correction}
+
 First, we see a statistically significant change
-in home win percentage in 2020 from 2017-19, with a p-value of 0.0497 for the
+in home win percentage in 2020 from 2017--19, with a p-value of 0.0497 for the
 \(z\)-test and 0.0400 for Fisher's exact test. The 95\% confidence
@@
 Home team performance did not seem to be negatively impacted by losing home-court
-advantage like expected. Home scoring, two-point and three-point shooting
+advantage like expected. Home scoring, two-point and three-point
 shooting all show no significant difference, on average, between 2020 vs.
-2017-19 based on p-values from both tests. However, the Wilcoxon test and \(z\)-test have conflicting results
+2017--19 based on p-values from both tests. However, the Wilcoxon test and \(z\)-test have conflicting results
 for free throws. The \(z\)-test p-value of 0.0692 indicates no significant difference, while
@@
 0.0004 for Wilcoxon. It is important to note both p-values also remain significant after
-a Bonferonni correction giving strong indication of significance. The average difference in points was estimated to
+a Bonferonni correction giving strong indication of significance. The
+average difference in points was estimated to
 be about 5 points, with 95\% CI \((2.083, 7.988)\). Likewise, the away team two-point
@@
 
-\section*{acknowledgements}
-The authors thank the University of Connecticut and the UConn Department of Statistics.
+% \section*{acknowledgements}
+% The authors thank the University of Connecticut and the UConn Department of Statistics.
 %Note correct LaTeX quotations above. Do not use the " symbol, but rather double ` followed by double '
@@
 %\section*{about the student authors}
-Michael Price graduated in May 2021 from the University of Connecticut where he double-majored in Statistics and Economics and minored in Data Analytics.
-He began his graduate studies at the University of Delaware studying Applied Statistics in August 2021.
+Michael Price graduated in May 2021 from the University of Connecticut
+where he double-majored in Statistics and Economics and minored in
+Data Analytics. He began his graduate studies at the University of
+Delaware studying Applied Statistics in August 2021.
+
+\jy{are you interested in phd in statistics?}
 
@@
 \section*{press summary}
-The purpose of this study was to examine the 2020 NBA playoffs, which were played inside of a bubble in Disney World because of the COVID-19 pandemic.
-The hope was to generate new insights about the effects of home-court advantage because the bubble created an unprecedented neutral playing field for the NBA.
-For the first time the effects of home-court advantage on home and away teams could be easily separated and studied individually, typically studies can only compare them relative to each other.
-This study is focused on team scoring totals and team shooting percentage on two-pointers, three-pointers, and free throws.
-Comparing these measures for home teams and away teams in 2020 vs.
-2017-2019 shows that the 2020 playoffs favored away teams more than usual, particularly with two-point shooting and total scoring.
-The implication of these findings is home-court advantage seems to be the result of negative effects on away team, not positive effects on home-team.
+The purpose of this study was to examine the 2020 NBA playoffs, which
+were played inside of a bubble in Disney World because of the COVID-19
+pandemic. The hope was to generate new insights about the effects of
+home-court advantage because the bubble created an unprecedented
+neutral playing field for the NBA. For the first time the effects of
+home-court advantage on home and away teams could be easily separated
+and studied individually, typically studies can only compare them
+relative to each other. This study is focused on team scoring totals
+and team shooting percentage on two-pointers, three-pointers, and free
+throws. Comparing these measures for home teams and away teams in 2020
+vs. 2017--2019 shows that the 2020 playoffs favored away teams more
+than usual, particularly with two-point shooting and total
+scoring. The implication of these findings is home-court advantage
+seems to be the result of negative effects on away team, not positive
+effects on home-team.
 

### 2022-02-07 Manuscript/Response_To_Comments.tex
Update Response_To_Comments.tex
Correcting line width

@@
 \item
-  Additional commentary and thoughts on what caused away teams to be more affected by neutral courts more than home teams.
+  Additional commentary and thoughts on what caused away teams to
+  be more affected by neutral courts more than home teams.
   \item
-  A deeper discussion on the significance and impact of the shift in two-point shooting.
+  A deeper discussion on the significance and impact of the shift 
+  in two-point shooting.
 \item
-  Alterations to graphs and charts to better and more clearly describe the data and findings.
+  Alterations to graphs and charts to better and more clearly describe
+  the data and findings.
 \end{enumerate}
@@
 \begin{comment}
-  I’d have the paper examined for grammar.  I think there are commas missing or misplaced periodically. Couldn’t add that to the pdf effectively.
+  I’d have the paper examined for grammar.  I think there are commas missing or 
+  misplaced periodically. Couldn’t add that to the pdf effectively.
 \end{comment}
 
- The paper was carefully reread and edited, line by line, in order to correct grammar errors and improve the overall flow of the paper.
+ The paper was carefully reread and edited, line by line, in order to
+ correct grammar errors and improve the overall flow of the paper.
 
@@
 
-There were a series of brief comments provided on an annotated version of our manuscript submission.
-This included many small comments for missing words or asking for a brief clarification about a point made.
-These comments were addressed accordingly and were resolved.
-The more extensive comments/concerns are addressed individually below.
+There were a series of brief comments provided on an annotated version of our
+manuscript submission. This included many small comments for missing words or 
+asking for a brief clarification about a point made. These comments were addressed
+accordingly and were resolved. The more extensive comments/concerns are addressed 
+individually below.
 
 \begin{comment}
-  Reference lines (mean/median) would be nice in the graphs to more clearly see the small shifts.
+  Reference lines (mean/median) would be nice in the graphs to more clearly
+  see the small shifts.
 \end{comment}
 
-The figures were remade with added reference line at the mean for all applicable histograms.
+The figures were remade with added reference line at the mean for all 
+applicable histograms.
 
 \begin{comment}
-Use of adjusted p-values was appreciated, but those should be listed in the table.
-I would also recommend using either the nonparametric (Wilcoxon) or parametric (t-test) exclusively—no need for both.
-I think the assumptions are satisfied for the parametric test.
-Fisher’s exact test is appropriate
+Use of adjusted p-values was appreciated, but those should be listed in 
+the table.  I would also recommend using either the nonparametric (Wilcoxon)
+or parametric (t-test) exclusively—no need for both.  I think the 
+assumptions are satisfied for the parametric test.  Fisher’s exact test 
+is appropriate
 \end{comment}
 
- Columns were added to the results table to accommodate including the adjusted p-values.
-As for including both the non-parametric and parametric tests, we think including both is justified.
-We felt it strengthens the confidence in the results when the conclusions from each the non-parametric and parametric tests agree with one another.
-While the assumptions appear to be met for parametric test, it does not hurt to show tests not dependent on assumptions also show significant results.
+ Columns were added to the results table to accommodate including the 
+ adjusted p-values. As for including both the non-parametric and parametric 
+ tests, we think including both is justified. We felt it strengthens the confidence
+ in the results when the conclusions from each the non-parametric and parametric 
+ tests agree with one another. While the assumptions appear to be met for parametric
+ test, it does not hurt to show tests not dependent on assumptions also show
+ significant results.
 
 \begin{comment}
- Why are two pointers singled out in the data?
-Any insight on why 2P\% is affected more?
-I would provide some insight.
-My hypothesis is officiating bias as shown in many papers (less discretion on 3P fouls).
+ Why are two pointers singled out in the data? Any insight on why 2P\% is 
+ affected more? I would provide some insight. My hypothesis is officiating 
+ bias as shown in many papers (less discretion on 3P fouls). 
 \end{comment}
 
- We thank the reviewer for bringing this concern to our attention.
-We realize we missed discussing the significance of studying each type of shot individually, not just 2-pointers.
-Discussion on the importance of examining each of the three types of shot was added, including the 2-point shot.
+ We thank the reviewer for bringing this concern to our attention. We realize
+ we missed discussing the significance of studying each type of shot individually,
+ not just 2-pointers. Discussion on the importance of examining each of the three
+ types of shot was added, including the 2-point shot.
 
 \begin{comment}
-Comment a bit on these effect sizes. Does a 3\% decrease in 2P\% matter in a game? That may only be 1 shot. While it's statistically significant, I wonder if it's practically significant.
+Comment a bit on these effect sizes. Does a 3\% decrease in 2P\% matter in a game?
+That may only be 1 shot. While it's statistically significant, I wonder 
+if it's practically significant.
 \end{comment}
 
- A paragraph was added to the discussion section detailing the practical significance of the 3\% change in 2-point shooting.
-In short, many NBA games are decided by 6 points or less and in those games late-game strategy is completely dependent exactly how many points one team is behind.
-Even a small shift in score can drastically shift probability of outcomes and strategy for teams.
-The paper now goes into detail on how a 3-point swing in these close games can completely change the game.
+ A paragraph was added to the discussion section detailing the practical 
+ significance of the 3\% change in 2-point shooting. In short, many NBA games 
+ are decided by 6 points or less and in those games late-game strategy is 
+ completely dependent exactly how many points one team is behind. Even a small
+ shift in score can drastically shift probability of outcomes and strategy 
+ for teams. The paper now goes into detail on how a 3-point swing in these 
+ close games can completely change the game.
 
 \begin{comment}
-Would be interesting to examine if these home court advantages are more pronounced for specific teams.
+Would be interesting to examine if these home court advantages are 
+more pronounced for specific teams.
 \end{comment}
@@
  We are appreciative of this suggestion for another possible topic of analysis. 
-Cross referencing this comment with the source provided by referee~2, we definitely have reason to believe the shift in home-court advantage would be more pronounced for specific teams.
-However, we didn't have the sample size to test for strength home-court advantage in individual teams.
-The teams we'd expect to have the most pronounced affect, the Denver Nuggets and Utah Jazz, only played 9 and 3 home games, respectively, in the bubble.
-We did add some comments about this in the conclusion explaining how this analysis would be possible if generating a larger sample were possible and what we'd expect to see.
+ Cross referencing this comment with the source provided by referee~2, we definitely 
+ have reason to believe the shift in home-court advantage would be more pronounced 
+ for specific teams. However, we didn't have the sample size to test for strength 
+ home-court advantage in individual teams. The teams we'd expect to have the 
+ most pronounced affect, the Denver Nuggets and Utah Jazz, only played 9 and 3
+ home games, respectively, in the bubble. We did add some comments about this 
+ in the conclusion explaining how this analysis would be possible if generating
+ a larger sample were possible and what we'd expect to see.
 
@@
 \begin{comment}
-Good motivation of the problem and why we have additional useful information to answer questions about
+Good motivation of the problem and why we have additional useful information
+to answer questions about
 home advantage.
-It would be good to mention here, or elsewhere, different (potential) sources of home advantage, and discuss
+It would be good to mention here, or elsewhere, different (potential) 
+sources of home advantage, and discuss
 which may/may not still exist in the NBA bubble. For example,
@@
 \item
-This is partly in the literature review where you discuss Courneya and Carron’s article, but I think you could
-touch on that there are different sources of home advantage somewhere in the introduction. There are, for
-example, no travel miles differences between the two teams and (I think) no back-to-back games, so home
-advantage from those probably doesn’t exist in the bubble. There may, for example, still be referee bias in
-the bubble, though. Referee bias may only exist when there’s a home crowd, this case the video monitors
-and the handful of attendees don’t affect referee. Or it might exist because referees know who the home
-team is.
+This is partly in the literature review where you discuss Courneya and Carron’s article,
+but I think you could touch on that there are different sources of home advantage 
+somewhere in the introduction. There are, for example, no travel miles differences between the 
+two teams and (I think) no back-to-back games, so homeadvantage from those probably doesn’t 
+exist in the bubble. There may, for example, still be referee bias in
+the bubble, though. Referee bias may only exist when there’s a home crowd, this
+case the video monitors and the handful of attendees don’t affect referee. 
+Or it might exist because referees know who the home team is.
 \end{comment}
 
- We thank the reviewer for being appreciative and understanding of the good motivations driving this paper.
-We agree that a deeper discussion of why home-court advantage would not exist in the bubble would be beneficial to setting up the paper.
-We added in a discussion of why all the different factors play into creating home-court advantage and how each one may or may not still exist in the bubble.
-For the most part, these factors disappeared in the context of the bubble, except travel/time and miles for reasons we discuss in the paper.
+ We thank the reviewer for being appreciative and understanding of the 
+ good motivations driving this paper. We agree that a deeper discussion 
+ of why home-court advantage would not exist in the bubble would be beneficial 
+ to setting up the paper. We added in a discussion of why all the different 
+ factors play into creating home-court advantage and how each one may or may 
+ not still exist in the bubble. For the most part, these factors disappeared 
+ in the context of the bubble, except travel/time and miles for reasons we
+ discuss in the paper.
 
 \begin{comment}
-Also, this paper https://arxiv.org/abs/1701.05976 finds that altitude matters (Colorado and Utah teams
-have the highest home advantage in multiple sports), so that could be added/discussed. So changes in home
-advantage in the bubble could impact teams differently, and NBA teams Nuggets and Jazz could be hurt
+Also, this paper https://arxiv.org/abs/1701.05976 finds that altitude matters
+(Colorado and Utah teams have the highest home advantage in multiple sports),
+so that could be added/discussed. So changes in home advantage in the bubble
+could impact teams differently, and NBA teams Nuggets and Jazz could be hurt
 more than other teams since they no longer benefit from an altitude advantage.
@@
  We are very grateful that referee~2 brought \citet{Lopez} to our attention. 
-It was added to Literature Review and discussed in the Introduction when setting up and explaining the factors that would be missing from normal home-court advantage when playing in the NBA bubble.
-The most noteworthy point added from this paper, as pointed out by referee~2, was the evidence supporting geographic factors impact home-court advantage.
-Particularly, teams like the Utah Jazz and Denver Nuggets are the NBA teams who would benefit from the high-altitude effects.
-This is discussed in more detail in the manuscript.
+ It was added to Literature Review and discussed in the Introduction when setting 
+ up and explaining the factors that would be missing from normal home-court 
+ advantage when playing in the NBA bubble. The most noteworthy point added 
+ from this paper, as pointed out by referee~2, was the evidence supporting 
+ geographic factors impact home-court advantage. Particularly, teams like 
+ the Utah Jazz and Denver Nuggets are the NBA teams who would benefit from 
+ the high-altitude effects. This is discussed in more detail in the manuscript.
 
@@
 \begin{comment}
-In your plots with histograms, it would be useful to have a vertical line showing the mean to visually compare
+In your plots with histograms, it would be useful to have a vertical 
+line showing the mean to visually compare
 the two distributions.
 
-It seems like you made this using facet\_grid(shot.type ~ season) or something similar. I think it would
-be easy to add vertical lines fairly easily by using group\_by and summarise to make a data frame df.means
+It seems like you made this using facet\_grid(shot.type ~ season) or
+something similar. I think it would be easy to add vertical lines fairly 
+easily by using group\_by and summarise to make a data frame df.means
 like this
@@
 If the columns season and shot type are the same as in your other data
-set, and the same names used in facet\_grid, a different vertical line should show up for each season and shot type.
+set, and the same names used in facet\_grid, a different vertical line
+should show up for each season and shot type.
 \end{comment}
 
- While this was not the exact coding solution used to add reference lines to the histograms, it was greatly appreciated and was the perfect stepping stone to finding the code solution.
-As mentioned earlier, the code was added to include a reference line at the mean on each histogram we provided.
+ While this was not the exact coding solution used to add reference lines
+ to the histograms, it was greatly appreciated and was the perfect stepping 
+ stone to finding the code solution. As mentioned earlier, the code was
+ added to include a reference line at the mean on each histogram we provided.
 
 \begin{comment}
-If you use a regression, you could take into account strength of team, strength of opponent, and home
-advantage simultaneously. This could be done with any of the metrics you considered. For example,
+If you use a regression, you could take into account strength of team,
+strength of opponent, and home advantage simultaneously. This could 
+be done with any of the metrics you considered. For example,
 Score ~ 1 + HomeTeamOnOffenseIndicator + OffensiveTeam + DefensiveTeam
-Did you consider a regression? Why would it, or would it not, be useful? If strength of opponent (in the
-example above, the DefensiveTeam) doesn’t vary much from team to team (in other words, if there is a
-balanced schedule), then it might not be that helpful to account for strength of opponent. If it does vary a
-lot, then it might help.
-In a typical regular season, the NBA schedule is pretty balanced. Is the schedule a balance schedule in the
-bubble?
-Would accounting for team strength change the home advantage estimates? If so, in what situations would
-it increase home advantage estimates? Decrease?
-You don’t necessarily have to do all of this analysis, but you should at least discuss somewhere why you
-didn’t use a regression, and justify your decision. I’m not saying that a regression is necessarily better,
-and the answers to the above questions might suggest it wouldn’t help. But a lot of people would think of
-estimating home advantage using a regression by including a home indicator like the example above, and
-then using the hypothesis test where the null hypothesis is that the regression coefficient for home indicator
-is zero. So it is important to, at minimum, discuss your choice to not use a regression, and justify that
+Did you consider a regression? Why would it, or would it not, be useful? 
+If strength of opponent (in the example above, the DefensiveTeam) doesn’t
+vary much from team to team (in other words, if there is a
+balanced schedule), then it might not be that helpful to account for 
+strength of opponent. If it does vary a lot, then it might help.
+In a typical regular season, the NBA schedule is pretty balanced. 
+Is the schedule a balance schedule in the bubble?
+Would accounting for team strength change the home advantage estimates? If so, 
+in what situations would it increase home advantage estimates? Decrease?
+You don’t necessarily have to do all of this analysis, but you should at least 
+discuss somewhere why you didn’t use a regression, and justify your decision.
+I’m not saying that a regression is necessarily better, and the answers to 
+the above questions might suggest it wouldn’t help. But a lot of people
+would think of estimating home advantage using a regression by including a home 
+indicator like the example above, and then using the hypothesis test where 
+the null hypothesis is that the regression coefficient for home indicator
+is zero. So it is important to, at minimum, discuss your choice to not use
+a regression, and justify that
 choice.
@@
 
- This is certainly a interesting suggestion and something we explored.
-However, we elected to not add the regression to the paper.
-We did take the suggestion to add a paragraph to the paper explaining why the hypothesis testing alone sufficed to answering our research questions without need for the regression.
-We did not consider accounting for strength of opponent necessary.
-Given the study was conducted using the playoff teams, match-ups were largely between strong teams with small gaps in ability.
-Had the study been done using regular season play, controlling for strength of opponent would have been beneficial since bottom-tier teams would be included.
-More details on why the regression was not required are included in the manuscript.
+ This is certainly a interesting suggestion and something we explored. However, we 
+ elected to not add the regression to the paper. We did take the suggestion to add
+ a paragraph to the paper explaining why the hypothesis testing alone sufficed to
+ answering our research questions without need for the regression. We did not
+ consider accounting for strength of opponent necessary. Given the study was 
+ conducted using the playoff teams, match-ups were largely between strong teams 
+ with small gaps in ability. Had the study been done using regular season play, 
+ controlling for strength of opponent would have been beneficial since bottom-tier 
+ teams would be included. More details on why the regression was not required are 
+ included in the manuscript.
 
 \begin{comment}
-I’d be interested in hearing more thoughts about why away teams were helped, but home teams weren’t
-hurt, in the bubble. You did mention “No longer having to face the struggle of traveling, pressure from
-opposing fans, or playing on an unfamiliar court”. A little more detail would be nice. Of the factors that
-impact home advantage, which ones were unchanged for the home team, but changed for the away team,
-and in what specific situation?
-For example, when a team has two home games in a row, they didn’t travel pre-bubble in between those
-games, and still didn’t travel in the bubble. If their opponent, the away team, had 2 away games in a row,
-they did travel pre-bubble, but didn’t have to travel in the bubble. So in that case there is a difference in
-home advantage.
-But if the home team instead had an away game before their current home game, then they had to travel
-before their home game, and a lack of travel advantage for the home team doesn’t apply to that game. I
-would imagine this happens frequently, and that 30-40\% of home games are preceded by an away game.
-Also, if away teams were impacted more, wouldn’t they also be better on defense, which would cause the
-home team’s score to go down, for example? Do the home advantage only impact offensive performance?
+I’d be interested in hearing more thoughts about why away teams were helped,
+but home teams weren’t hurt, in the bubble. You did mention “No longer having
+to face the struggle of traveling, pressure from opposing fans, or playing on
+an unfamiliar court”. A little more detail would be nice. Of the factors that
+impact home advantage, which ones were unchanged for the home team, but
+changed for the away team, and in what specific situation?
+For example, when a team has two home games in a row, they didn’t travel 
+pre-bubble in between those games, and still didn’t travel in the bubble. If 
+their opponent, the away team, had 2 away games in a row, they did travel pre-bubble,
+but didn’t have to travel in the bubble. So in that case there is a difference in home advantage.
+But if the home team instead had an away game before their current home game,
+then they had to travel before their home game, and a lack of travel advantage
+for the home team doesn’t apply to that game. I would imagine this happens 
+frequently, and that 30-40\% of home games are preceded by an away game.
+Also, if away teams were impacted more, wouldn’t they also be better on 
+defense, which would cause the home team’s score to go down, for example?
+Do the home advantage only impact offensive performance?
 \end{comment}
 
- Thank you for bringing this up, this comment allowed us to revisit the discussion and derive a more clear well-thought out conclusion from the results.
-A few paragraphs were added to the discussion section of the paper detailing why it is likely only away teams performance were significantly changed.
-The discussion strengthened our determination that home-court advantage is best categorized as negative effects on away team.
+ Thank you for bringing this up, this comment allowed us to revisit the discussion 
+ and derive a more clear well-thought out conclusion from the results. A few 
+ paragraphs were added to the discussion section of the paper detailing why
+ it is likely only away teams performance were significantly changed. The 
+ discussion strengthened our determination that home-court advantage is best 
+ categorized as negative effects on away team.
 

### 2022-02-08 Manuscript/Response_To_Comments.tex
Update Response_To_Comments.tex
Line width problems were addressed
Update Response_To_Comments.tex
Adding specific location of changes

@@
 \section*{Summary}
-\jy{Practice the line width style in tex source}
 
@@
 Point-by-point responses are as follows with the original comments in \emph{italic}.
-
-\jy{In each response, be more specific on the location of the changes made (give
-  paragraph numbers in which section)}
 
@@
 The figures were remade with added reference line at the mean for all 
-applicable histograms.
+applicable histograms, in the Data section.
 
@@
 
- Columns were added to the results table to accommodate including the 
- adjusted p-values. As for including both the non-parametric and parametric 
- tests, we think including both is justified. We felt it strengthens the confidence
- in the results when the conclusions from each the non-parametric and parametric 
+ Columns were added to the results table, located at the beginning of the Results section,
+ to accommodate including the adjusted p-values. As for including both the non-parametric
+ and parametric tests, we think including both is justified. We felt it strengthens the 
+ confidence in the results when the conclusions from each the non-parametric and parametric 
  tests agree with one another. While the assumptions appear to be met for parametric
@@
  not just 2-pointers. Discussion on the importance of examining each of the three
- types of shot was added, including the 2-point shot.
+ types of shot was added, including the 2-point shot, at the end of the first 
+ paragraph in the Data section.
 
@@
 
- A paragraph was added to the discussion section detailing the practical 
+ A paragraph was added to the Discussion section(fourth paragraph) detailing the practical 
  significance of the 3\% change in 2-point shooting. In short, many NBA games 
@@
  home games, respectively, in the bubble. We did add some comments about this 
- in the conclusion explaining how this analysis would be possible if generating
- a larger sample were possible and what we'd expect to see.
+ at the end of the last paragraph in the discussion section which explains
+ how this analysis would be possible if generating a larger sample were
+ possible and what we'd expect to see.
 
@@
  factors play into creating home-court advantage and how each one may or may 
- not still exist in the bubble. For the most part, these factors disappeared 
+ not still exist in the bubble. This addition to the paper can be found in the first paragraph
+ of the Introduction section. For the most part, these factors disappeared 
  in the context of the bubble, except travel/time and miles for reasons we
@@
  We are very grateful that referee~2 brought \citet{Lopez} to our attention. 
- It was added to Literature Review and discussed in the Introduction when setting 
+ It was added to Literature Review(at the end of the third paragraph) and
+ discussed in the Introduction(first paragraph) when setting 
  up and explaining the factors that would be missing from normal home-court 
@@
  While this was not the exact coding solution used to add reference lines
- to the histograms, it was greatly appreciated and was the perfect stepping 
+ to the histograms,it was greatly appreciated and was the perfect stepping 
  stone to finding the code solution. As mentioned earlier, the code was
- added to include a reference line at the mean on each histogram we provided.
+ added to include a reference line at the mean on each histogram we provided,
+ within the Data section of the manuscript.
 
@@
  teams would be included. More details on why the regression was not required are 
- included in the manuscript.
+ included in the manuscript, in the final paragraph of the Methods section.
 
@@
  and derive a more clear well-thought out conclusion from the results. A few 
- paragraphs were added to the discussion section of the paper detailing why
+ paragraphs were added to the Discussion section(paragraphs 5 and 7) of the paper detailing why
  it is likely only away teams performance were significantly changed. The 

### 2022-02-08 Manuscript/NBABubbleAnalysis.Rmd
Update NBABubbleAnalysis.Rmd
Setting aspect ratio to 3:4

@@
       geom_hline(aes(yintercept = 0.5), color = "green")+
-      geom_text(aes(label=winpct), hjust=0, vjust=-1, size=2)
+      geom_text(aes(label=winpct), hjust=0, vjust=-1, size=2)+
+      theme(aspect.ratio=3/4)
 ```
@@
   geom_vline(data=ref_line2, aes(xintercept=PTS))+
-  geom_text(data=ref_line2, aes(x=PTS+3.5, label="mean", y=0.045))
+  geom_text(data=ref_line2, aes(x=PTS+3.5, label="mean", y=0.045))+
+  theme(aspect.ratio=3/4)
 ```
@@
   geom_vline(data=ref_line3, aes(xintercept=Percentage))+
-  geom_text(data=ref_line3, aes(x=Percentage+6, label="mean", y=0.075))
+  geom_text(data=ref_line3, aes(x=Percentage+6, label="mean", y=0.075))+
+  theme(aspect.ratio=3/4)
 ```
@@
   geom_vline(data=ref_line4, aes(xintercept=Percentage))+
-  geom_text(data=ref_line4, aes(x=Percentage+6, label="mean", y=0.075))
+  geom_text(data=ref_line4, aes(x=Percentage+6, label="mean", y=0.075))+
+  theme(aspect.ratio=3/4)
 ```

### 2022-02-08 Manuscript/manuscript.tex
Update manuscript.tex
Changing float to tbp, and width control out of tex since 3/4 aspect ratio is in R

@@
 
-\jy{use `tbp' for figure/table locations instead of `H', unless it is
-  required by the journal}
-
-\jy{remake the figure~\ref{fig:Fig1}: make aspect 3:4 (height:width;
-  increase the font size of the annonations; control the height/weight
-  from within R now that we know how the figures will appear in the
-  published version (so that you don't need to control width in latex.}
-
-\jy{explain what the vertical lines are in all the figure captions}
-
 \begin{figure}[tbp]
   \centering
-  \includegraphics[width=0.4\linewidth]{Fig1}
+  \includegraphics{Fig1}
   \caption{Winning percentage of NBA home teams in the playoffs since
@@
 
-\begin{figure}[H]
+\begin{figure}[tbp]
   \centering
-  \includegraphics[width=0.4\linewidth]{Fig2}
+  \includegraphics{Fig2}
   \caption{Histograms of home (blue) and away (red) scoring
@@
 
-\begin{figure}[H]
+\begin{figure}[tbp]
   \centering
-  \includegraphics[width=0.4\linewidth]{Fig3}
+  \includegraphics{Fig3}
   \caption{Histograms of home shooting percentages for two-
@@
 
-\begin{figure}[H]
+\begin{figure}[tbp]
   \centering
-  \includegraphics[width=0.4\linewidth]{Fig4}
+  \includegraphics{Fig4}
   \caption{Histograms of away shooting percentages for two-pointers,

### 2022-02-08 Manuscript/manuscript.tex
Update manuscript.tex
use booktabs package to add horizontal rules; replace 1 with 1.0000; fixed``Z-test'' to be in the same font as in the text

@@
 
-\jy{use booktabs package to add horizontal rules; replace 1 with
-  1.0000; make sure ``Z-test'' is in the same font as in the text.}
-
 \begin{table}[tbp]
@@
 \cmidrule(lr){4-5}\cmidrule(lr){6-7}
-  &          &                & Z-test & Wilcoxon & Z-test & Wilcoxon\\
+  &          &                & \(Z\)-test & Wilcoxon & \(Z\)-test & Wilcoxon\\
   \midrule
   Home Win & 0.482 & 0.613 & 0.0497* & 0.0400* & 0.4473 & 0.3598\\
-Home Scoring & 1.101 & 1.081 & 0.2321 & 0.2985 & 1 & 1\\
+  \midrule
+Home Scoring & 1.101 & 1.081 & 0.2321 & 0.2985 & 1.0000 & 1.0000\\
+\midrule
 Away Scoring & 1.091 & 1.040 & 0.0008* & 0.0004* & 0.0075* & 0.0035*\\
-Home 2P & 0.523 & 0.515 & 0.4335 & 0.5719 & 1 & 1\\
-Home 3P & 0.363 & 0.357 & 0.5733 & 0.8852 & 1 & 1\\
+\midrule
+Home 2P & 0.523 & 0.515 & 0.4335 & 0.5719 & 1.0000 & 1.0000\\
+\midrule
+Home 3P & 0.363 & 0.357 & 0.5733 & 0.8852 & 1.0000 & 1.0000\\
+\midrule
 Home FT & 0.793 & 0.774 & 0.0692 & 0.0496* & 0.6228 & 0.4464\\
+\midrule
 Away 2P & 0.536 & 0.504 & 0.0003* & 0.0003* & 0.0030 & 0.0023\\
-Away 3P & 0.357 & 0.346 & 0.3256 & 0.3081 & 1 & 1\\
-Away FT & 0.783 & 0.777 & 0.6601 & 0.8370 & 1 & 1\\
+\midrule
+Away 3P & 0.357 & 0.346 & 0.3256 & 0.3081 & 1.0000 & 1.0000\\
+\midrule
+Away FT & 0.783 & 0.777 & 0.6601 & 0.8370 & 1.0000 & 1.0000\\
   \bottomrule

### 2022-02-12 Manuscript/manuscript.tex
Update manuscript.tex
aded reference to Bonferroni correction
Update manuscript.tex
Added study about referee bias in NBA

@@
 factors also account for referee bias which may impact the NBA.  One
-interesting
-finding cited by \citet{Carron2005} is that the absence of crowds results
-in overall performance increases. Lastly, as discussed in the introduction, \citet{Lopez}
+interesting finding cited by \citet{Carron2005} is that the absence of crowds results
+in overall performance increases. Another study by \citet{Price} is able to find 
+some evidence of referee bias. They focus on two measures, DTOs and NTOs. DTOs, 
+discretionary turnovers, are defined as turnovers always caused by the ref blowing
+the whistle while the ball is in play. NTOs, non-discretionary turnovers, are determined 
+directly by players with no ref whistle, or the ball going out of bounds. The use
+these to test ref bias by checking how variables, like home vs. away team, affect
+DTOs relative to NTOs. They found evidence that a home bias does exist. In fact, home 
+bias increases both during the playoffs and in games with higher attendance. This is
+crucial to the NBA bubble, which consists of playoff games with no attendance. Lastly, 
+as discussed in the introduction, \citet{Lopez}
 find evidence that geographic factors like altitude may influence and strengthen home-court
@@
 \(z\)-tests and Wilcoxon tests. The p-values are all fairly similar for both
-tests giving strong confidence in the conclusions drawn. Also, reported are
-the point estimates of the two samples in each comparison.
-
-\jy{add a reference to Bonferroni correction}
+tests giving strong confidence in the conclusions drawn. Additionally, the adjusted
+p-values, calculated using a Bonferroni correction, and point estimates
+for each sample are provided.
 
@@
 shooting percentage. Likewise, NBA teams shot
-an average of about 53 two-point shots per game, in our data, leading to a three-point increase in total scoring. Even this
+an average of about 53 two-point shots per game, in our data, leading to
+a three-point increase in total scoring. Even this
 small swing in scoring for away teams can make a big difference in outcome of
@@
 are the games where the three points matter most. Suppose there are 30 seconds
-left and the away team is losing. If the away team has the ball, only down three, they can take their time finding a good shot which can
+left and the away team is losing. If the away team has the ball, only down three, 
+they can take their time finding a good shot which can
 still either be a two or three given the score and remaining time. In the context 
@@
 percentage. One could be the possible removal of officiating bias in favor of the 
-home team during the bubble. Away teams in the bubble may have benefited from more 
+home team during the bubble. As we saw in \cite{Price}, home bias is normally
+stronger in high attendance games. Thus, with no attendance at games, it's reasonable
+to say officiating bias was smaller than normal. Away teams in the bubble may have benefited from more 
 fouls called on drives to the basket. On drives to the basket there can often be a 

### 2022-02-13 Manuscript/manuscript.tex
Update manuscript.tex
Update manuscript.tex
Addition of Jones and Boudreaux sources and slight restructing of literature review

@@
 referee calls is less clear. However, the data does seem to lean slightly in
-favor of affecting visiting team performance. \citet{Harville} studied the
-effect of home-court advantage using the 1991-1992 college basketball season.
-Unlike the NBA, it is not uncommon to have a few games played at neutral sites
-during the college basketball season. This allowed them to construct two samples,
-one of home teams and one of neutral teams. They formulated their study in a
-regression predicting the expected difference in score for home teams.
-They set up their study to find if the home teams won games by more points when
-they had home-court advantage vs.~when playing at neutral court. This study
-concluded with evidence supporting home-court advantage.
+favor of affecting visiting team performance. Another study focused on scoring
+was conducted by \cite{Jones}. \cite{Jones} analyzed scoring patterns across each of the 
+4 quarters to analyze the effects of home-court advantage over the course of a game.
+He found that typically 2/3 of the benefits of home-court advantage are received in 
+the first quarter with the remaining 1/3 slowly accumulating over the 
+final 3 quarters. This implies home-court advantage is most effective when teams 
+use it to build a lead they then just have to maintain for the remainder of the game.
+If they are losing after the first quarter, there is a sharp drop in win probability.
 
@@
 advantage for teams in high altitude regions.
+
+There are a few examples of natural experiments in basketball. \citet{Harville} studied the
+effect of home-court advantage using the 1991-1992 college basketball season.
+Unlike the NBA, it is not uncommon to have a few games played at neutral sites
+during the college basketball season. This allowed them to construct two samples,
+one of home teams and one of neutral teams. They formulated their study in a
+regression predicting the expected difference in score for home teams.
+They set up their study to find if the home teams won games by more points when
+they had home-court advantage vs.~when playing at neutral court. This study
+concluded with evidence supporting home-court advantage. Also, \cite{Boudreaux} is 
+able to construct a natural experiment using the Los Angeles Lakers and Los Angeles 
+Clippers. Since these two teams share a home stadium, many factors like travel and 
+familiarity are nullified. However, the designated home team has larger crowd support
+due to attendance from their season ticket holders. When they single out the effect 
+of having a sympathetic crowd, \cite{Boudreaux} estimate crowd effects
+increase chance of winning 21--22.8 percentage pointts.
 
@@
 act as a "sixth defender" supporting their team by cheering loudly and making it
-harder on the offense. This would explain how away teams were able to improve
+harder on the offense. Also, as seen in the study \cite{Price}, fans can 
+create turnovers by pressuring officials to make calls in their favor.  
+This would explain how away teams were able to improve
 offensively in the bubble. Home teams did not have the regular help from the

### 2022-02-13 Manuscript/manuscript.tex
Update manuscript.tex

@@
 They set up their study to find if the home teams won games by more points when
-they had home-court advantage vs.~when playing at neutral court. This study
-concluded with evidence supporting home-court advantage. Also, \cite{Boudreaux} is 
-able to construct a natural experiment using the Los Angeles Lakers and Los Angeles 
-Clippers. Since these two teams share a home stadium, many factors like travel and 
-familiarity are nullified. However, the designated home team has larger crowd support
-due to attendance from their season ticket holders. When they single out the effect 
-of having a sympathetic crowd, \cite{Boudreaux} estimate crowd effects
-increase chance of winning 21--22.8 percentage pointts.
+they had home-court advantage vs.~when playing on a neutral court. This study
+concluded with evidence supporting home-court advantage. Also, \cite{Boudreaux}
+is able to construct a natural experiment using the Los Angeles Lakers and Los
+Angeles Clippers. Since these two teams share a home stadium, many factors like 
+travel and familiarity are nullified. However, the designated home team has larger
+crowd support due to attendance from their season ticket holders. When they
+single out the effect of having a sympathetic crowd, \cite{Boudreaux} estimate crowd 
+effects increase the chance of winning 21--22.8 percentage points.
 

### 2022-02-13 Manuscript/Response_To_Comments.tex
Update Response_To_Comments.tex

@@
  the high-altitude effects. This is discussed in more detail in the manuscript.
+  Also, a few other references were added to strengthen the paper.
+  First, \cite{Price} discuss empirical evidence in favor of referee bias towards
+  home teams. Second, \cite{Jones} analyzes the effects of home-court advantage
+  as component parts and how the effects of home-court advantage impact each 
+  individual quarter of play. Lastly, \cite{Boudreaux} uses a natural
+  experiment to estimate how the crowd effect of home-court advantage impacts winning.
 

### 2022-02-13 Manuscript/manuscript.tex
Update manuscript.tex

@@
 so players can easily communicate and focus. However, when on defense, fans can
-act as a "sixth defender" supporting their team by cheering loudly and making it
+act as a ``''sixth defender''`` supporting their team by cheering loudly and making it
 harder on the offense. Also, as seen in the study \cite{Price}, fans can 
@@
 
-\jy{are you interested in phd in statistics?}
-
 % The Press Summary section is NOT optional.  Write a paragraph describing the paper in a manner suitable for the press; see previous journal editions for examples.

### 2022-02-13 Manuscript/manuscript.tex
Update manuscript.tex

@@
 favor of affecting visiting team performance. Another study focused on scoring
-was conducted by \cite{Jones}. \cite{Jones} analyzed scoring patterns across each of the 
+was conducted by \citet{Jones}. \citet{Jones} analyzed scoring patterns across each of the 
 4 quarters to analyze the effects of home-court advantage over the course of a game.
@@
 they had home-court advantage vs.~when playing on a neutral court. This study
-concluded with evidence supporting home-court advantage. Also, \cite{Boudreaux}
+concluded with evidence supporting home-court advantage. Also, \citet{Boudreaux}
 is able to construct a natural experiment using the Los Angeles Lakers and Los
@@
 crowd support due to attendance from their season ticket holders. When they
-single out the effect of having a sympathetic crowd, \cite{Boudreaux} estimate crowd 
+single out the effect of having a sympathetic crowd, \citet{Boudreaux} estimate crowd 
 effects increase the chance of winning 21--22.8 percentage points.
@@
 percentage. One could be the possible removal of officiating bias in favor of the 
-home team during the bubble. As we saw in \cite{Price}, home bias is normally
+home team during the bubble. As we saw in \citet{Price}, home bias is normally
 stronger in high attendance games. Thus, with no attendance at games, it's reasonable
@@
 act as a ``''sixth defender''`` supporting their team by cheering loudly and making it
-harder on the offense. Also, as seen in the study \cite{Price}, fans can 
+harder on the offense. Also, as seen in the study \citet{Price}, fans can 
 create turnovers by pressuring officials to make calls in their favor.  

### 2022-02-13 Manuscript/manuscript.tex
Update manuscript.tex

@@
 
-\jy{add some references from google scholar search ``home team
-  advantage nba''; we have only 10 references now.}
-
 There is voluminous literature on the effects of home-court advantage.
@@
 referee calls is less clear. However, the data does seem to lean slightly in
-favor of affecting visiting team performance. Another study focused on scoring
-was conducted by \citet{Jones}. \citet{Jones} analyzed scoring patterns across each of the 
-4 quarters to analyze the effects of home-court advantage over the course of a game.
-He found that typically 2/3 of the benefits of home-court advantage are received in 
-the first quarter with the remaining 1/3 slowly accumulating over the 
-final 3 quarters. This implies home-court advantage is most effective when teams 
-use it to build a lead they then just have to maintain for the remainder of the game.
-If they are losing after the first quarter, there is a sharp drop in win probability.
+favor of affecting visiting team performance. Another study focused
+on scoring was conducted by \citet{Jones}, he analyzes scoring patterns across
+each of the 4 quarters to analyze the effects of home-court advantage over
+the course of a game. He found that typically 2/3 of the benefits of 
+home-court advantage are received in the first quarter with the 
+remaining 1/3 slowly accumulating over the final 3 quarters. This implies
+home-court advantage is most effective when teams use it to build a 
+lead early and maintain it for the remainder of the game. If they are 
+losing after the first quarter, there is a sharp drop in win probability.
 
@@
 interesting finding cited by \citet{Carron2005} is that the absence of crowds results
-in overall performance increases. Another study by \citet{Price} is able to find 
-some evidence of referee bias. They focus on two measures, DTOs and NTOs. DTOs, 
-discretionary turnovers, are defined as turnovers always caused by the ref blowing
-the whistle while the ball is in play. NTOs, non-discretionary turnovers, are determined 
-directly by players with no ref whistle, or the ball going out of bounds. The use
-these to test ref bias by checking how variables, like home vs. away team, affect
-DTOs relative to NTOs. They found evidence that a home bias does exist. In fact, home 
-bias increases both during the playoffs and in games with higher attendance. This is
-crucial to the NBA bubble, which consists of playoff games with no attendance. Lastly, 
-as discussed in the introduction, \citet{Lopez}
+in overall performance increases. Another study by \citet{Price} is able 
+to find some evidence of referee bias. They focus on two measures, DTOs and NTOs.
+DTOs, discretionary turnovers, are defined as turnovers always caused by the ref
+blowing the whistle while the ball is in play. NTOs, non-discretionary turnovers, 
+are determined directly by players with no ref whistle, or the ball going out of
+bounds. The use these to test ref bias by checking how variables, like home vs. away
+team, affect DTOs relative to NTOs. They found evidence that a home bias does exist. 
+In fact, home bias increases both during the playoffs and in games with higher attendance.
+This is crucial to the NBA bubble, which consists of playoff games with no attendance. 
+Lastly, as discussed in the introduction, \citet{Lopez}
 find evidence that geographic factors like altitude may influence and strengthen home-court
@@
 concluded with evidence supporting home-court advantage. Also, \citet{Boudreaux} 
-is able to construct a natural experiment using the Los Angeles Lakers and Los
-Angeles Clippers. Since these two teams share a home stadium, many factors like 
-travel and familiarity are nullified. However, the designated home team has larger
-crowd support due to attendance from their season ticket holders. When they
-single out the effect of having a sympathetic crowd, \citet{Boudreaux} estimate crowd 
-effects increase the chance of winning 21--22.8 percentage points.
+is able to construct a natural experiment using the Los Angeles Lakers and 
+Los Angeles Clippers. Since these two teams share a home stadium, many factors like travel 
+and familiarity are nullified. However, the designated home team has larger crowd
+support due to attendance from their season ticket holders. By comparing
+Lakers--Clippers matchups, they single out the effect of having a sympathetic
+crowd and \citet{Boudreaux} estimate crowd effects increase the chance
+of winning 21--22.8 percentage points.
 

### 2022-02-13 Manuscript/NBABubbleAnalysis.Rmd
Update NBABubbleAnalysis.Rmd

@@
 teams coming in and struggling with lower oxygen levels at a higher altitude, 
-which the home teams are accustomed to playing with \citet{Lopez}. Examples
+which the home teams are accustomed to playing with @Lopez. Examples
 of high altitude teams in the NBA would be the Utah Jazz and Denver Nuggets, 
@@
 Many NBA home-court advantage studies analyze the effects by studying shooting
-percentages. @Kotecki reported significant evidence of home-court advantage by 
-comparing field goal percentage, free throw percentage
-and points scored in home vs. away teams. He found all of these measures showed evidence that
+percentages. @Kotecki reported significant evidence of home-court
+advantage by comparing field goal
+percentage, free throw percentage and points scored in home vs.~away teams.
+He found all of these measures showed evidence that
 home-court advantage helps home teams play better. @Cao studied the effects
-of pressure on performance in the NBA. Using free throws as their measure of
-interest, they tested whether home
+of pressure on performance in the NBA. Using free throws as their measure of interest, 
+they tested whether home
 fans could distract and put pressure on opposing players to make free
 throws. However, they did not find significant evidence
-that home status has a substantial impact on missing from the free throw
-line. @Harris used two-point shots, three-point shots and free throws as measures
-of interest to study home-court advantage. Two-point shots were found to be
-the strongest predictor of home-court advantage. They suggested
-that home teams should try to shoot more two-point shots and force
-their opponent to take more two-point shot attempts. This will 
-help home teams have a greater control of the game play and help maximize the benefits 
+that home status has a substantial impact on missing from the free throw line. 
+@Harris used two-point shots, three-point shots and free
+throws as measures of interest to study home-court advantage. Two-point
+shots were found to be the strongest predictor of home-court advantage. They
+suggested that home teams should try to shoot more two-point shots and force
+their opponent to take more two-point shot attempts. This help home teams have a
+greater control of the game play and help maximize the benefits 
 home-court advantage.
@@
 in this study were scoring, violations, and turnovers. This study was conducted
-using the men’s basketball programs at two large universities. The study finds that
+using the men's basketball programs at two large universities. The study finds that
 social support, like booing, is an important contributor to home-court advantage.
@@
 referee calls is less clear. However, the data does seem to lean slightly in
-favor of affecting visiting team performance.  @Harville studied the
-effect of home-court advantage using the 1991-1992 college basketball season.
-Unlike the NBA, it is not uncommon to have a few games played at neutral sites
-during the college basketball season. This allowed them to construct two samples,
-one of home teams and one of neutral teams. They formulated their study in a
-regression predicting the expected difference in score for home teams.
-They set up their study to find if the home teams won games by more points when
-they had home-court advantage vs. when playing at neutral court. This study
-concluded with evidence supporting home-court advantage.
+favor of affecting visiting team performance. Another study focused
+on scoring was conducted by @Jones, he analyzes scoring patterns across
+each of the 4 quarters to analyze the effects of home-court advantage over
+the course of a game. He found that typically 2/3 of the benefits of 
+home-court advantage are received in the first quarter with the 
+remaining 1/3 slowly accumulating over the final 3 quarters. This implies
+home-court advantage is most effective when teams use it to build a 
+lead early and maintain it for the remainder of the game. If they are 
+losing after the first quarter, there is a sharp drop in win probability.
 
@@
 these factors would all be removed if games were played at a neutral site, even
-if one team was designated as “home team”. This study was reviewed a decade later
+if one team was designated as ``home team''. This study was reviewed a decade later
 by @Carron2005. The 2005 review goes over the new findings from studies
@@
 and professional sports. The NBA is not a league which has rules that
-may favor the home team, like batting last in the MLB, but these rule factors 
-also account for referee bias which may impact the NBA. One interesting finding cited
-by @Carron2005 is that the absence of crowds results in overall performance increases.
-Lastly, as discussed in the introduction, @Lopez find evidence that geographic
-factors like altitude may influence and strengthen home-court advantage for teams in
-high altitude regions.
+may favor the home team, like batting last in the MLB, but these rule
+factors also account for referee bias which may impact the NBA.  One
+interesting finding cited by @Carron2005 is that the absence of crowds results
+in overall performance increases. Another study by @Price is able 
+to find some evidence of referee bias. They focus on two measures, DTOs and NTOs.
+DTOs, discretionary turnovers, are defined as turnovers always caused by the ref
+blowing the whistle while the ball is in play. NTOs, non-discretionary turnovers, 
+are determined directly by players with no ref whistle, or the ball going out of
+bounds. The use these to test ref bias by checking how variables, like home vs. away
+team, affect DTOs relative to NTOs. They found evidence that a home bias does exist. 
+In fact, home bias increases both during the playoffs and in games with higher attendance.
+This is crucial to the NBA bubble, which consists of playoff games with no attendance. 
+Lastly, as discussed in the introduction, @Lopez
+find evidence that geographic factors like altitude may influence and strengthen home-court
+advantage for teams in high altitude regions.
+
+There are a few examples of natural experiments in basketball. @Harville studied the
+effect of home-court advantage using the 1991-1992 college basketball season.
+Unlike the NBA, it is not uncommon to have a few games played at neutral sites
+during the college basketball season. This allowed them to construct two samples,
+one of home teams and one of neutral teams. They formulated their study in a
+regression predicting the expected difference in score for home teams.
+They set up their study to find if the home teams won games by more points when
+they had home-court advantage vs.~when playing on a neutral court. This study
+concluded with evidence supporting home-court advantage. Also, @Boudreaux 
+is able to construct a natural experiment using the Los Angeles Lakers and 
+Los Angeles Clippers. Since these two teams share a home stadium, many factors like travel 
+and familiarity are nullified. However, the designated home team has larger crowd
+support due to attendance from their season ticket holders. By comparing
+Lakers--Clippers matchups, they single out the effect of having a sympathetic
+crowd and @Boudreaux estimate crowd effects increase the chance
+of winning 21--22.8 percentage points.
 

### 2022-02-14 Manuscript/Response_To_Comments.tex
final check

@@
 
- A paragraph was added to the Discussion section(fourth paragraph) detailing the practical 
+ A paragraph was added to the Discussion section (fourth paragraph) detailing the practical 
  significance of the 3\% change in 2-point shooting. In short, many NBA games 
@@
 We are very grateful that referee~2 brought \citet{Lopez} to our attention. 
- It was added to Literature Review(at the end of the third paragraph) and
+It was added to Literature Review (at the end of the third paragraph) and
 discussed in the Introduction(first paragraph) when setting 
@@
 the high-altitude effects. This is discussed in more detail in the manuscript.
-  Also, a few other references were added to strengthen the paper.
+
+In addition, a few other references were added to the Literature Review section.
 First, \cite{Price} discuss empirical evidence in favor of referee bias towards
@@
 
- \jy{Check the references of Lopez's paper to see if we can cite
-   anything there; this is also a place to mention the additional
-   references you found. }
-
 \begin{comment}
@@
 
-It seems like you made this using facet\_grid(shot.type ~ season) or
+It seems like you made this using \texttt{facet\_grid(shot.type ~ season)} or
 something similar. I think it would be easy to add vertical lines fairly 
-easily by using group\_by and summarise to make a data frame df.means
+easily by using \texttt{group\_by} and summarise to make a data frame df.means
 like this
@@
 and then adding
-geom\_vline(data=df.means, aes(x=mean.perc)) +
+\texttt{geom\_vline(data=df.means, aes(x=mean.perc)) +}
 somewhere in your ggplot code.
@@
  While this was not the exact coding solution used to add reference lines
- to the histograms,it was greatly appreciated and was the perfect stepping 
+ to the histograms, it was greatly appreciated and was the perfect stepping 
  stone to finding the code solution. As mentioned earlier, the code was
@@
 be done with any of the metrics you considered. For example,
-Score ~ 1 + HomeTeamOnOffenseIndicator + OffensiveTeam + DefensiveTeam
+\texttt{Score \texttildelow 1 + HomeTeamOnOffenseIndicator + OffensiveTeam + DefensiveTeam}.
 Did you consider a regression? Why would it, or would it not, be useful? 

### 2022-02-14 Manuscript/NBABubbleAnalysis.Rmd
final check

@@
 highlight <- data.frame(winpct = .482, years = 2020)
-pdf("Fig1.pdf")
-ggplot(winyear, aes(x=years, y=winpct))+
+# pdf("Fig1.pdf")
+p <- ggplot(winyear, aes(x=years, y=winpct))+
     geom_point()+
@@
     geom_hline(aes(yintercept = 0.5), color = "green")+
-      geom_text(aes(label=winpct), hjust=0, vjust=-1, size=2)+
-      theme(aspect.ratio=3/4)
+    geom_text(aes(label=winpct), hjust=0, vjust=-1, size=2)
+#    theme(aspect.ratio=3/4)
+ggsave("Fig1.pdf", plot = p, height = 4.2, width = 6)
 ```
@@
 
-pdf("Fig2.pdf")
-ggplot(combined) +
+# pdf("Fig2.pdf")
+p <- ggplot(combined) +
   geom_histogram(aes(x = PTS, y = ..density.., fill = Home.Away), bins = 10, binwidth = 5)+
@@
   geom_vline(data=ref_line2, aes(xintercept=PTS))+
-  geom_text(data=ref_line2, aes(x=PTS+3.5, label="mean", y=0.045))+
-  theme(aspect.ratio=3/4)
+    geom_text(data=ref_line2, aes(x=PTS+3.5, label="mean", y=0.045))
+ggsave("Fig2.pdf", plot = p, height=4.2, width = 6)
 ```
@@
 ref_line3 <- aggregate(data=homegraph, Percentage ~ Year + Type, FUN=mean)
-pdf("Fig3.pdf")
-ggplot(homegraph) +
+# pdf("Fig3.pdf")
+p <- ggplot(homegraph) +
   geom_histogram(aes(x = Percentage, y = ..density..), fill = "#00BFC4", bins = 10, binwidth = 5)+
@@
   geom_vline(data=ref_line3, aes(xintercept=Percentage))+
-  geom_text(data=ref_line3, aes(x=Percentage+6, label="mean", y=0.075))+
-  theme(aspect.ratio=3/4)
+  geom_text(data=ref_line3, aes(x=Percentage+6, label="mean", y=0.075))
+ggsave("Fig3.pdf", plot = p, height=4.2, width = 6)
 ```
@@
 ref_line4 <- aggregate(data=awaygraph, Percentage ~ Year + Type, FUN=mean)
-pdf("Fig4.pdf")
-ggplot(awaygraph) +
+# pdf("Fig4.pdf")
+p <- ggplot(awaygraph) +
   geom_histogram(aes(x = Percentage, y = ..density..), fill = "#F8766D", bins = 10, binwidth = 5)+
@@
   geom_vline(data=ref_line4, aes(xintercept=Percentage))+
-  geom_text(data=ref_line4, aes(x=Percentage+6, label="mean", y=0.075))+
-  theme(aspect.ratio=3/4)
+    geom_text(data=ref_line4, aes(x=Percentage+6, label="mean", y=0.075))
+ggsave("Fig4.pdf", width=6, height=4.2)
 ```
